### 2017-06-11 ms.tex
light curve figure

%%
%% Beginning of file 'sample61.tex'
%%
%% Modified 2016 September
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.1 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.1
\documentclass[twocolumn]{aastex61}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex61}
%% 
%% where the arguement options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex61}

%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{July 1, 2016}
%\revised{September 27, 2016}
%\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJL}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.1 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{\aastex\ Atmosphere of WASP-107b}
\shortauthors{Kreidberg et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{WASP-107b Atmosphere}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.1 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.1 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.1 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Laura Kreidberg}
\email{laura.kreidberg@cfa.harvard.edu}

\author{Laura Kreidberg}
\affiliation{Harvard Society of Fellows\
78 Mt. Auburn St.\\
Cambridge, MA 02138, USA}
\affiliation{Harvard-Smithsonian Center for Astrophysics\
60 Garden St.\\
Cambridge, MA 02138}
\nocollaboration

%\author{Kevin Stevenson}
%\affiliation{Space Telescope Science Institute}
%\affiliation{AAS Journals Associate Editor-in-Chief}
%\nocollaboration

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.1 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

WASP-107b is baller.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{planets and satellites: individual (WASP-107b), planets and satellites: atmospheres}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.

%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}

\section{Data Reduction}

\section{Analysis}
Light Curve Fits
Star Spots?

\section{Discussion} \label{sec:discuss}

Comparative planetology with HAT-P-11b, HAT-P-26b

\acknowledgments
Fei Dai

\vspace{5mm}
\facilities{HST(WFC3)}

%% Similar to \facility{}, there is the optional \software command to allow 
%% authors a place to specify which programs were used during the creation of 
%% the manusscript. Authors should list each code and include either a
%% citation or url to the code inside ()s when available.

\software{astropy \citep{2013A&A...558A..33A},  
          Cloudy \citep{2013RMxAA..49..137F}, 
          SExtractor \citep{1996A&AS..117..393B}
          }

%% Appendix material should be preceded with a single \appendix command.
%% There should be a \section command for each appendix. Mark appendix
%% subsections with the same markup you use in the main body of the paper.

%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
%% The equation counter will reset when it encounters the \appendix
%% command and will number appendix equations (A1), (A2), etc. The
%% Figure and Table counter will not reset.

%\appendix

%% The reference list follows the main body and any appendices.
%% Use LaTeX's thebibliography environment to mark up your reference list.
%% Note \begin{thebibliography} is followed by an empty set of
%% curly braces.  If you forget this, LaTeX will generate the error
%% "Perhaps a missing \item?".
%%
%% thebibliography produces citations in the text using \bibitem-\cite
%% cross-referencing. Each reference is preceded by a
%% \bibitem command that defines in curly braces the KEY that corresponds
%% to the KEY in the \cite commands (see the first section above).
%% Make sure that you provide a unique KEY for every \bibitem or else the
%% paper will not LaTeX. The square brackets should contain
%% the citation text that LaTeX will insert in
%% place of the \cite commands.

%% We have used macros to produce journal name abbreviations.
%% \aastex provides a number of these for the more frequently-cited journals.
%% See the Author Guide for a list of them.

%% Note that the style of the \bibitem labels (in []) is slightly
%% different from previous examples.  The natbib system solves a host
%% of citation expression problems, but it is necessary to clearly
%% delimit the year from the author name used in the citation.
%% See the natbib documentation for more details and options.

\begin{thebibliography}{}

\end{thebibliography}

%% This command is needed to show the entire author+affilation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\end{document}

% End of file `sample61.tex'.

### 2017-06-11 Figures/lc_fit.py

@@ -1,0 +1,746 @@
+import sys
+#sys.path.insert(1, '/Library/Frameworks/Python.framework/Versions/2.7/lib/python2.7/site-packages/')
+import mpfit
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import rc
+import glob
+import os
+from astropy.io import ascii
+import getopt
+import seaborn as sns
+import matplotlib
+from datetime import datetime
+import time as pythontime
+import multiprocessing as mp
+import emcee
+import batman
+import pickle
+import corner
+from scipy.stats import norm
+
+sns.set_context("talk")
+sns.set_style("white")
+sns.set_style("ticks", {"xtick.direction":"in", "ytick.direction":"in"})
+matplotlib.rcParams.update({'lines.markeredgewidth':0.3})
+matplotlib.rcParams.update({'axes.formatter.useoffset':False})
+
+def E(t, eta, f, Etot, tau, E0):
+	return eta*f/(eta*f/Etot + 1./tau) + (E0 - eta*f/(eta*f/Etot + 1./tau))*np.exp(-(eta*f/Etot +1./tau)*t)
+
+
+def computeRMS(data, maxnbins=None, binstep=1, isrmserr=False):
+	# bin data into multiple bin sizes
+	npts    = data.size
+	if maxnbins == None:
+		maxnbins = npts/10.
+	binsz   = np.arange(1, maxnbins+binstep, step=binstep)
+	nbins   = np.zeros(binsz.size)
+	rms     = np.zeros(binsz.size)
+	rmserr  = np.zeros(binsz.size)
+	for i in range(binsz.size):
+		nbins[i] = int(np.floor(data.size/binsz[i]))
+		bindata   = np.zeros(nbins[i], dtype=float)
+# bin data
+# ADDED INTEGER CONVERSION, mh 01/21/12
+		for j in range(int(nbins[i])):
+			bindata[j] = data[j*binsz[i]:(j+1)*binsz[i]].mean()
+		# get rms
+		rms[i]    = np.sqrt(np.mean(bindata**2))
+		rmserr[i] = rms[i]/np.sqrt(2.*int(nbins[i]))
+	# expected for white noise (WINN 2008, PONT 2006)
+	stderr = (data.std()/np.sqrt(binsz))*np.sqrt(nbins/(nbins - 1.))
+	if isrmserr == True:
+		return rms, stderr, binsz, rmserr
+	else:
+		return rms, stderr, binsz
+
+
[…]

### 2017-06-11 Figures/plot_lc.py

@@ -1,0 +1,77 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from lc_fit import LightCurveData, Model
+import os, glob
+import pickle
+from pylab import *
+import seaborn as sns
+
+sns.set_context("talk")
+sns.set_style("white")
+sns.set_style("ticks", {"xtick.direction":"in", "ytick.direction":"in"})
+#sns.set_palette(sns.cubehelix_palette(22, start=.5, rot=-1, light=0.4, dark = 0.6))
+#sns.set_palette(sns.husl_palette(22, h=1., s=0.4))
+sns.set_palette(sns.diverging_palette(220, 20, n=22))
+
+path =  "lsq_fits"
+files = glob.glob(os.path.join(path, "*.p"))	
+
+
+per = 5.7214742
+
+plt.figure()
+ax1 = plt.subplot(121)
+#ax2 = ax1.twinx()
+
+ax1.text(2.5, 1.002,'$\lambda$ ($\mu$m)', fontsize=10)
+
+for i, f in enumerate(files):
+	p = pickle.load(open(f, 'rb'))
+	data, model = p[0], p[1]
+	offset = -0.01*float(i)
+	ind = model.phase > 0.5
+	model.phase[ind] -= 1
+	ind = model.phase_hr > 0.5
+	model.phase_hr[ind] -= 1
+	ind = np.argsort(model.phase_hr)
+	model.phase_hr, model.lc_hr = model.phase_hr[ind], model.lc_hr[ind]
+	ax1.plot(model.phase_hr*per*24., model.lc_hr + offset, color = '0.5')
+	ax1.plot(model.phase*per*24., model.data_corr + offset, marker='.', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5)
+	#ax1.text(3., 1 + offset, '{0:d}'.format(int(model.rms))+", "+'{0:0.2f}'.format(data.wavelength), fontsize=10)
+	ax1.text(2.7, 0.992 + offset+0.002, '{0:0.2f}'.format(data.wavelength), fontsize=10)
+			
+	#ax2.plot(model.phase*per*24., np.ones_like(model.phase)*data.wavelength, linewidth=0.)
+
+	#ax2.set_ylabel("Wavelength ($\mu$m)")
+
+	ax1.set_xlim(-3.2, 3.2)
+	ax1.set_ylim(0.76, 1.01)
[…]

### 2017-06-11 starspot_properties/calc_spot_properties.py

@@ -1,0 +1,49 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+h = 6.626e-34   #J/s
+c = 3.0e8       #m/s
+kb = 1.4e-23     #J/K
+
+def blackbody(l,T): return 2*h*c**2/(l**5*(np.exp(h*c/(l*kb*T))-1))
+
+kep_response = np.genfromtxt("kepler_response.txt", skip_header = 2)
+kep_response[:,0] = kep_response[:,0]/1.e9					#gets wavelength in m
+
+wave = kep_response[:,0]
+
+Tstar = 4300
+deltaT = 300
+Tspot = Tstar - deltaT
+
+delta_wave = wave[1] - wave[0]
+
+A = 0.006 									#fractional spot area	
+
+#integrates flux over kepler bandpass
+spot_flux = np.sum(blackbody(wave, Tspot)*delta_wave)
+star_flux = np.sum(blackbody(wave, Tstar)*delta_wave)
+
+
+print "variability at Kepler wavelengths"
+print spot_flux*A/star_flux*1e2
+
+wave = np.linspace(0.5, 5, 100)*1.e-6
+spot_flux = blackbody(wave, Tspot)
+star_flux = blackbody(wave, Tstar)
+
+#conservative case? half of spots are persistent
+Fquiet = star_flux*(1. - A) + spot_flux*A
+Factive = star_flux*(1.-2*A) + spot_flux*2.*A
+
+#conservative case? half of spots are persistent
+Fquiet = star_flux 
+Factive = star_flux*(1.-A) + spot_flux*A
+
+delta = 0.021
+plt.plot(wave*1e6, (Fquiet-Factive)/Factive*delta*1e6)
+plt.xlabel("wavelength (um)")
+plt.ylabel("Transit depth correction due to unocculted spots (ppm")
+plt.show()
+
+

### 2017-06-11 starspot_properties/plot_activity.py

@@ -1,0 +1,19 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+n = np.genfromtxt("WASP107_north_20170511-20170605.txt")
+s = np.genfromtxt("WASP107_south_20170321-20170605.txt")
+
+plt.axhline(0, color='0.5')
+
+#plt.plot(n[:,0], n[:,1], '.k')
+plt.plot(s[:,0], s[:,1], '.k')
+
+plt.axvline(2457876.12557956, color = 'orange')		#Spitzer Channel 1 
+plt.axvline(2457870.404354212, color = 'red')		#Spitzer Channel 2
+plt.axvline(2457910.4, color = 'cyan')			#HST
+
+plt.ylim(-0.02, 0.02)
+
+plt.show()
+

### 2017-06-11 ms.tex
lc figure

@@
 
+\section{Observations}
+We observed a single transit of WASP-107b with HST's Wide Field Camera 3 (WFC3) instrument on UT 5-6 June 2017.
+The transit observation consisted of five HST orbits.
+At the beginning of each 96-minute orbit, we took an image of the target with the F130N filter (exposure time = 4.2 s).
+This direct image is used for wavelength calibration.
+For the remainder of the target visibility period (about 45 minutes), we obtained time series spectra with the G141 grism, which provides low-resolution spectroscopy over the wavelength range $1.1 - 1.7\,\mu$m.
+We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{https://github.com/spacetelescope/PandExoi\_HST}.
+
+As is standard for observations of bright targets, we used the spatial scanning observing mode, which smears the spectrum in the spatial direction over the course of an exposure.
+
+This mode enables longer exposures than staring mode ,
+
+improves the duty cycle of the observations  The observations had an exposure time of
+
+The scan rate was FIXME arcseconds per second, resulting in a spectrum approximately FIXME pixels in height.
+
 \section{Data Reduction}

### 2017-06-11 Figures/plot_lc.py

@@ -52,6 +52,7 @@ for i, f in enumerate(files):
 ax3 = plt.subplot(122)
 #ax4 = ax3.twinx()
 
+rms = []
 ax3.text(2.6, 0.08,' rms\n(ppm)', fontsize=10)
 #ax3.text(-3.5, 1,'$\lambda$ ($\mu$m)', fontsize=10)
 for i, f in enumerate(files):
@@ -65,7 +66,8 @@ for i, f in enumerate(files):
 	plt.axhline(offset, color ='0.5')
 	ax3.plot(model.phase*per*24., model.resid/data.flux*1e2 + offset, marker='.', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5)
 #	ax3.text(-3.5, offset-1, '{0:0.2f}'.format(data.wavelength), fontsize=10)
-	ax3.text(2.8, offset - .1, '{0:d}'.format(int(model.rms)), fontsize=10)
+	ax3.text(2.8, offset - .11, '{0:d}'.format(int(model.rms)), fontsize=10)
+	rms.append(model.rms)
 	#ax4.plot(model.phase*per*24., np.ones_like(model.phase)*data.wavelength, linewidth=0.)
 	ax3.set_xlim(-3.2, 3.4)
 	ax3.set_ylim(-4.5, .35)
@@ -73,5 +75,9 @@ for i, f in enumerate(files):
 	ax3.set_ylabel("Residuals ($\\times10^{-2}$, offset)")
 #	ax4.set_ylabel("Wavelength ($\mu$m)")
 
+
+print "mean, median rms:", np.mean(np.array(rms)), np.median(np.array(rms))
+
 plt.tight_layout()
+plt.savefig("lc.png")
 plt.show()

### 2017-07-01 ms.tex
list of figs/tabs
new figs/text

@@
 
+%Figures and Tables:
+% 1.  transit light curves
+% 2.  best fit spectrum
+% 3.  retrieval results
+% 4.  table of transit depths
+% 5.  JWST predictions for cloud models
+
 \section{Introduction} \label{sec:intro}
@@
 \section{Analysis}
-Light Curve Fits
+
+\subsection{White Light Curve Fit}
+We created a band-integrated ``white" light curve.  Typical morphology for instrument systematics, with the largest ramp
+
+The first orbit showed larger, We fit this light curve with an analytic model of the form $F(t) = S(t)\times T(t)$, where $S$ is a systematics model and $T$ is a transit model.
+
+For the transit model, we used the \texttt{batman} package \citep{kreidberg15b}. We fixed the
+
+In our initial fit, we noticed a what looked like a star spot crossing event in orbit 3. We masked this, and lo! there is the spot
+
+we fit white lc without orbit 3, used model that had extra ramp in orbit 1, fixed LD on PHOENIX model
+
+Transit
+
+\begin{equation}
+x
+\end{equation}
+
 Star Spots?
+
+\subsection{Spectroscopic Light Curve Fits}
+We explored
+
+Confirmation from KBS
+
+\section{Atmospheric Retrieval}
+We use the CHIMERA chemically-consistent transmission retrieval tool described in \cite{kreidberg15} determine the basic atmospheric properties.
+Briefly, the transmission spectrum solves the transmission geometry problem using the equations in \cite{brown11, tinetti12}.
+We parametrize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
+We have subsequently upgraded the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases generated from the pre-tabulated line-by-line cross section database ($dv/v~10^6$) described in \cite{freedman14}.
+The transmission forward model is coupled with the powerful PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.
+
+The nominal model set-up requires a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity and carbon-to-oxygen ratio for molecular composition determination, disquilibrium properties crudely parameterized with nitrogen species and carbon species quench pressures \citep{morley17}, and aerosol properties.
+We experiment with three atmospheric scenarios within this setup (Table XX):  Scenarios with and without methane, and with/without patchy cloud cover (as described and implemented in \citealt{line16}.
+In all three scenarios we fix the T-P profile "shape" but scale the irradiation temperature, which incorporates unknown variables like the albedo and heat transport efficiency.
+T-P profile shape information is not readily retrievable, nor does it strongly influence information derived from WFC3 transmission spectra \citep[e.g.][]{kreidberg15}.
+We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
+Aerosols are simply parameterized with a ``gray cloud top pressure" with the option for cloud patchiness.
+We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
 
@@
 
-Comparative planetology with HAT-P-11b, HAT-P-26b
+\subsection{Methane}
+
+\subsection{Comparative planetology with HAT-P-11b, HAT-P-26b}
+
+\subsection{Constraints on Condensate Properties}
+To model WASP-107b's spectrum with self-consistent aerosols, we use the methods described in \citep{fortney08, morley15}.
+We include models from solar to $50\times$ solar metallicity and solar C/O ratio.
+We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS, see \citealt{morley12}), varying the cloud sedimentation efficiency from 0.1 to 3.
+None of the cloudy models are sufficiently low amplitude to match the observed muted signal.
+We also model an ad hoc photochemical 'soot' layer near the top of the atmosphere, scaling results from previous photochemical models for GJ 436b \citep{line11, morley17}.
+With a sufficiently thick photochemical haze with particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.
+
+\subsection{Atmospheric Metallicity Predictions from Interior Structure Modeling}
+
+\section{Future Plans}
+Combine data from Spitzer/HST
+JWST GTO
 
@@
 Fei Dai
+Jessica Spake
+
+L.R.K. acknowledges support from the Harvard Society of Fellows and the Harvard Astronomy Department Institute for Theory and Computation.
 

### 2017-07-01 Figures/archive/plot_lc.py

@@ -1,0 +1,83 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from lc_fit import LightCurveData, Model
+import os, glob
+import pickle
+from pylab import *
+import seaborn as sns
+
+sns.set_context("talk")
+sns.set_style("white")
+sns.set_style("ticks", {"xtick.direction":"in", "ytick.direction":"in"})
+#sns.set_palette(sns.cubehelix_palette(22, start=.5, rot=-1, light=0.4, dark = 0.6))
+#sns.set_palette(sns.husl_palette(22, h=1., s=0.4))
+sns.set_palette(sns.diverging_palette(220, 20, n=22))
+
+path =  "lsq_fits"
+files = glob.glob(os.path.join(path, "*.p"))	
+
+
+per = 5.7214742
+
+plt.figure()
+ax1 = plt.subplot(121)
+#ax2 = ax1.twinx()
+
+ax1.text(2.5, 1.002,'$\lambda$ ($\mu$m)', fontsize=10)
+
+for i, f in enumerate(files):
+	p = pickle.load(open(f, 'rb'))
+	data, model = p[0], p[1]
+	offset = -0.01*float(i)
+	ind = model.phase > 0.5
+	model.phase[ind] -= 1
+	ind = model.phase_hr > 0.5
+	model.phase_hr[ind] -= 1
+	ind = np.argsort(model.phase_hr)
+	model.phase_hr, model.lc_hr = model.phase_hr[ind], model.lc_hr[ind]
+	ax1.plot(model.phase_hr*per*24., model.lc_hr + offset, color = '0.5')
+	ax1.plot(model.phase*per*24., model.data_corr + offset, marker='.', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5)
+	#ax1.text(3., 1 + offset, '{0:d}'.format(int(model.rms))+", "+'{0:0.2f}'.format(data.wavelength), fontsize=10)
+	ax1.text(2.7, 0.992 + offset+0.002, '{0:0.2f}'.format(data.wavelength), fontsize=10)
+			
+	#ax2.plot(model.phase*per*24., np.ones_like(model.phase)*data.wavelength, linewidth=0.)
+
+	#ax2.set_ylabel("Wavelength ($\mu$m)")
+
+	ax1.set_xlim(-3.2, 3.2)
+	ax1.set_ylim(0.76, 1.01)
[…]

### 2017-07-01 Figures/plot_lc.py

@@ -5,13 +5,12 @@ import os, glob
 import pickle
 from pylab import *
 import seaborn as sns
+import matplotlib.gridspec as gridspec
 
 sns.set_context("talk")
 sns.set_style("white")
 sns.set_style("ticks", {"xtick.direction":"in", "ytick.direction":"in"})
-#sns.set_palette(sns.cubehelix_palette(22, start=.5, rot=-1, light=0.4, dark = 0.6))
-#sns.set_palette(sns.husl_palette(22, h=1., s=0.4))
-sns.set_palette(sns.diverging_palette(220, 20, n=22))
+sns.set_palette(sns.diverging_palette(220, 20, n=20))
 
 path =  "lsq_fits"
 files = glob.glob(os.path.join(path, "*.p"))	
@@ -20,64 +19,99 @@ files = glob.glob(os.path.join(path, "*.p"))
 per = 5.7214742
 
 plt.figure()
-ax1 = plt.subplot(121)
-#ax2 = ax1.twinx()
+gs = gridspec.GridSpec(2, 2, height_ratios =[1,5], hspace=0.05, wspace=0.25)
 
-ax1.text(2.5, 1.002,'$\lambda$ ($\mu$m)', fontsize=10)
+ax1 = plt.subplot(gs[0,0])
+p = pickle.load(open("white_lc_fit.p", 'rb'))
+data, model = p[0], p[1]
+ind = model.phase > 0.5
+model.phase[ind] -= 1
+ind = model.phase_hr > 0.5
+model.phase_hr[ind] -= 1
+ind = np.argsort(model.phase_hr)
+model.phase_hr, model.lc_hr = model.phase_hr[ind], model.lc_hr[ind]
+ax1.plot(model.phase_hr*per*24., model.lc_hr , color = '0.5')
+ax1.plot(model.phase*per*24., model.data_corr, marker='.', color = 'w', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5)
+ax1.set_xlim(-3.2, 3.4)
+ax1.yaxis.set_major_locator(FixedLocator(np.array([0.98, 0.99, 1.0])))
+ax1.set_yticklabels(["0.98", "0.99", "1.0"])
+
+ax1.text(2.1, 0.985,'$\lambda$ ($\mu$m):\n1.1 - 1.7', fontsize=10)
+ax2 = plt.subplot(gs[0,1])
+
+ind = model.phase > 0.5
[…]

### 2017-07-01 starspot_properties/gj3470.py

@@ -1,0 +1,46 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+h = 6.626e-34   #J/s
+c = 3.0e8       #m/s
+kb = 1.4e-23     #J/K
+
+def blackbody(l,T): return 2*h*c**2/(l**5*(np.exp(h*c/(l*kb*T))-1))
+
+
+Tstar = 3600
+deltaT = 300
+Tspot = Tstar - deltaT
+
+
+A = 0.01 									#fractional spot area	
+
+
+wave = 700*10.e-9
+#integrates flux over kepler bandpass
+spot_flux = blackbody(wave, Tspot)
+star_flux = blackbody(wave, Tstar)
+
+
+print "variability amplitude"
+print spot_flux*A/star_flux*1e2
+
+wave = np.linspace(0.5, 5, 100)*1.e-6
+spot_flux = blackbody(wave, Tspot)
+star_flux = blackbody(wave, Tstar)
+
+#conservative case? half of spots are persistent
+Fquiet = star_flux*(1. - A) + spot_flux*A
+Factive = star_flux*(1.-2*A) + spot_flux*2.*A
+
+#conservative case? half of spots are persistent
+Fquiet = star_flux 
+Factive = star_flux*(1.-A) + spot_flux*A
+
+delta = 0.0057
+plt.plot(wave*1e6, (Fquiet-Factive)/Factive*delta*1e6)
+plt.xlabel("wavelength (um)")
+plt.ylabel("Transit depth correction due to unocculted spots (ppm")
+plt.show()
+
+

### 2017-07-03 ms.tex
data reduction

@@
 
-\title{WASP-107b Atmosphere}
+\title{Water, High-Altitude Condensates, and No Methane in the Atmosphere of the Warm Neptune WASP-107b}
 
@@
 
-%\author{Kevin Stevenson}
-%\affiliation{Space Telescope Science Institute}
-%\affiliation{AAS Journals Associate Editor-in-Chief}
-%\nocollaboration
+\author{Michael Line}
+\affiliation{Arizona State University}
+\nocollaboration
+
+\author{Caroline Morley}
+\affiliation{Harvard-Smithsonian Center for Astrophysics}
+\affiliation{Sagan Fellow}
+\nocollaboration
+
+\author{Kevin Stevenson}
+\affiliation{Space Telescope Science Institute}
+\nocollaboration
 
@@
 For the remainder of the target visibility period (about 45 minutes), we obtained time series spectra with the G141 grism, which provides low-resolution spectroscopy over the wavelength range $1.1 - 1.7\,\mu$m.
-We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{https://github.com/spacetelescope/PandExoi\_HST}.
-
+We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{https://github.com/spacetelescope/PandExo\_HST}.
 As is standard for observations of bright targets, we used the spatial scanning observing mode, which smears the spectrum in the spatial direction over the course of an exposure.
-
-This mode enables longer exposures than staring mode ,
-
-improves the duty cycle of the observations  The observations had an exposure time of
-
-The scan rate was FIXME arcseconds per second, resulting in a spectrum approximately FIXME pixels in height.
+The scan rate was FIXME arcseconds/sec.
 
 \section{Data Reduction}
+We reduced the data with the custom pipeline described in FIXME, which we summarize briefly here.
+For each exposure, we extracted the spectrum from each up-the-ramp sample (or ``stripe") separately using the optimal extraction algorithm of \citep{horne86}.
+The stripe spectra were then summed to create the final spectrum.
+For each stripe, the extraction box was 80 pixels high and centered on the stripe's midpoint in the spatial direction.
+To correct the change in dispersion solution over the length of the spatial scan, we interpolated each row to the wavelength scale of the row corresponding to the spectral trace.
+We also corrected for slight drift in the spectral direction (FIXME pixels) by interpolating each the first exposure.
+To subtract the background, we took the median of sky pixels that were uncontaminated by the target spectrum (rows $5-250$, columns $5-15$).
+The typical background counts were low: 40 photoelectrons/pixel, in comparison to $3\times10^4$ photoelectrons/pixel in the stellar spectrum.
 
 \section{Analysis}
+The data analysis had two parts: the band-integrated ``white" light curve fit and the spectroscopic light curve fits.
 
 \subsection{White Light Curve Fit}
-We created a band-integrated ``white" light curve.  Typical morphology for instrument systematics, with the largest ramp
+To create the raw white light curve, we summed each spectrum over the FIXME pixels in the spectral trace.
+The morphology of the white light curve is typical for WFC3 observations \cite{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend.
+The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.
 
-The first orbit showed larger, We fit this light curve with an analytic model of the form $F(t) = S(t)\times T(t)$, where $S$ is a systematics model and $T$ is a transit model.
+We fit the light curve with the analytic model of the form $F(t) = S(t)\times T(t)$, where $S$ is a systematics model and $T$ is a transit model.
+We used the same systematics model as FIXME, equation FIXME.
+To model the transit, we used the \texttt{batman} package \citep{kreidberg15b}.
+In our initial fit, we noticed an increase in flux during orbit three.
+We also found .
+We gave the data in orbit three no weight in the fit, and found that the
 
-For the transit model, we used the \texttt{batman} package \citep{kreidberg15b}. We fixed the
+In our final fit, we fixed the transit parameters on the values from \cite{dai17}.
 
-In our initial fit, we noticed a what looked like a star spot crossing event in orbit 3. We masked this, and lo! there is the spot
+fixed LD on PHOENIX model
 
-we fit white lc without orbit 3, used model that had extra ramp in orbit 1, fixed LD on PHOENIX model
+\subsection{Spectroscopic Light Curve Fits}
+We explored two different methods for fitting the spectroscopic light curves.
 
-Transit
+\subsection{Independent Analysis}
+We also performed an independent data reduction and fit using K. Stevenson's pipeline.
 
@@
 Star Spots?
-
-\subsection{Spectroscopic Light Curve Fits}
-We explored
 
@@
 \subsection{Constraints on Condensate Properties}
+The amplitude of WASP-107b's spectral features is about half that expected for an atmosphere free of condensates.
+In our retrieval analysis, we explored two simple parameterizations to model condensates: a gray opacity source and a sloped haze.
+The retrieved cloud-top pressure is FIXME at $1\,\sigma$ confidence.
+
+We also considered physically motivated, self-consistent cloud and haze models based on .
 To model WASP-107b's spectrum with self-consistent aerosols, we use the methods described in \citep{fortney08, morley15}.
@@
 
-\begin{thebibliography}{}
-
-\end{thebibliography}
+\bibliographystyle{apj}
+\bibliography{ms.bib}
 

### 2017-07-03 ms.tex
analysis

@@
 The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.
-
 We fit the light curve with the analytic model of the form $F(t) = S(t)\times T(t)$, where $S$ is a systematics model and $T$ is a transit model.
@@
 To model the transit, we used the \texttt{batman} package \citep{kreidberg15b}.
-In our initial fit, we noticed an increase in flux during orbit three.
-We also found .
-We gave the data in orbit three no weight in the fit, and found that the
+The transit model parameters are the orbital period $p$, time of FIXME $t_c$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.
 
-In our final fit, we fixed the transit parameters on the values from \cite{dai17}.
+In our initial fit, we allowed all the transit parameters to vary freely.
+We noticed two odd results: first, there was a slight increase in flux during orbit three, and second, the transit parameters were inconsistent with the precise analysis of the Kepler light curve from \cite{dai17}.
+We conjectured that the anomalies could be due to a star-spot crossing in orbit three.
+WASP-107 is an active star and star spot crossings have been observed before FIXME.
+To test this idea, we gave the data in orbit three no weight in the fit.
+In this case, the transit parameters agree well with those of \cite{dai17} and the spot-crossing feature is more pronounced (see Figure FIXME).
+The amplitude of the spot crossing feature is FIXME ppm, consistent with a spot of FIXME size and temperature (cf dai17 FIXME).
+We also found that the fitted limb darkening parameters were conistent with predictions from a PHOENIX model of FIXME properties, calculated with the FIXME package.
+FIXME: we tested interpolating models, didn't matter.
 
-fixed LD on PHOENIX model
+In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the values from \cite{dai17}.
+We also fixed the limb darkening parameters on the PHOENIX model prediction.
+The root-mean-square residuals were FIXME ppm (not including orbit three), compared to the expected shot noise of FIXME ppm.
+The achieved rms is close to the best value obtained for WFC3 data (FIXME), and we attribute the slight excess noise to either additional star spot crossings or loss of flux from the edge of the detector \citep[as has been seen for other bright targets;][]{FIXME}.
+We used the Markov chain Monte Carlo algorithm to estimate parameter uncertainties (as implemented in the the \texttt{emcee} package; FIXME).
+We scaled the per-point uncertainties by a factor of FIXME to achieve a reduced $\chi_2$ value of unity.
+The transit time was FIXME and the planet/star radius was FIXME.
 
 \subsection{Spectroscopic Light Curve Fits}
-We explored two different methods for fitting the spectroscopic light curves.
+We binned the spectrum into 20 spectrophotometric channels from FIXME to FIXME $\mu$m.
+We explored two different methods for fitting these spectroscopic light curves.
+First we used the \texttt{divide-white} technique, which assumes that the light curve systematics are nearly constant as a function of wavelength \citep{stevenson??, kreidberg14a}.
+For this approach, we fit a transit model times the systematics vector from the white light curve fit ($F(t)/T(t)$), rescaled by a factor $C_\lambda + V_\lambda t$.
+As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model.
+The remaining free parameters are $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
+As a test, we fit for a linear limb darkening parameter and found the results were consistent with the PHOENIX model predictions.
+An advantage of this approach is that it automatically removes the star spot crossing, enabling us to use orbit three with no additional correction.
+We expect that this model introduces a small chromatic effect due to the changing spot crossing feature amplitude with wavelength, but this effect is below the level of noise in our data FIXME.
+The fits reach the photon noise and have a median reduced $\chi_2$ value of FIXME.
+There are two channels with slightly higher
+
+As a test, we also fit the spectroscopic light curves with the same analytic model we used for the white light curve.
+The results were nearly identical to those obtained with the \texttt{divide-white} technique, except offset by a constant value of FIXME ppm.
+We attribute the offset to the uncorrected star-spot crossing in orbit three.
+Although the spot is not detectable in the spectroscopic light curves due to their higher photon noise, it may still affect the fit.
+report the values from the
+
+explored different binning
 
@@
 We also performed an independent data reduction and fit using K. Stevenson's pipeline.
-
-\begin{equation}
-x
-\end{equation}
-
-Star Spots?
 

### 2017-07-04 ms.tex
spec lc

--- 2017-07-03 ms.tex
@@
 
-\author{Laura Kreidberg}
+\author{Laura Kreidberg, Michael Line}
 \affiliation{Harvard Society of Fellows\
@@
 To create the raw white light curve, we summed each spectrum over the FIXME pixels in the spectral trace.
-The morphology of the white light curve is typical for WFC3 observations \cite{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend.
+The white light curve has systematic trends that are typical for WFC3 observations \cite{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend.
 The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.
-
-We fit the light curve with the analytic model of the form $F(t) = S(t)\times T(t)$, where $S$ is a systematics model and $T$ is a transit model.
+We fit the light curve with the analytic model of the form $F_\mathrm{white}(t) = S_\mathrm{white}(t)\times T_\mathrm{white}(t)$, where $S$ is a systematics model and $T$ is a transit model.
 We used the same systematics model as FIXME, equation FIXME.
 To model the transit, we used the \texttt{batman} package \citep{kreidberg15b}.
-In our initial fit, we noticed an increase in flux during orbit three.
-We also found .
-We gave the data in orbit three no weight in the fit, and found that the
+The transit model parameters are the orbital period $p$, time of FIXME $t_c$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.
 
-In our final fit, we fixed the transit parameters on the values from \cite{dai17}.
+\subsubsection{Star Spot Crossing}
+In our initial fit, we allowed all the transit parameters to vary freely.
+We noticed two odd results: first, there was a slight increase in flux during orbit three, and second, the transit parameters were inconsistent with the precise analysis of the Kepler light curve from \cite{dai17}.
+We conjectured that the anomalies could be due to a star-spot crossing event in orbit three, as WASP-107 is an active star and spot crossings have been observed before FIXME.
+To test this idea, we gave the data in orbit three no weight in the fit.
+For this case, the transit parameters agree well with those of \cite{dai17} and the spot-crossing feature is more pronounced (see Figure FIXME).
+The amplitude of the feature is FIXME ppm, consistent with a spot of FIXME size and temperature (cf dai17 FIXME).
+We conclude that the feature is indeed due to a spot crossing.
 
-fixed LD on PHOENIX model
+\subsubsection{Final Fit}
+In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the values from \cite{dai17}.
+We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star of FIXME properties, calculated with the FIXME package.
+The model values are consistent with what we obtain from fitting for the limb darkening coefficients as free parameters.
+The remaining free parameters were $t_c$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
+
+For the best fit light curve, the root-mean-square residuals were FIXME ppm (not including orbit three), compared to the expected shot noise of FIXME ppm.
+The achieved rms is close to the best value obtained for WFC3 data (FIXME), and we attribute the slight excess noise to either additional star spot crossings or loss of flux from the edge of the detector \citep[as has been seen for other bright targets;][]{FIXME}.
+We used the Markov chain Monte Carlo algorithm to estimate parameter uncertainties (as implemented in the the \texttt{emcee} package; FIXME).
+We scaled the per-point uncertainties by a factor of FIXME to achieve a reduced $\chi_2$ value of unity.
+The transit time was FIXME and the planet/star radius was FIXME.
 
 \subsection{Spectroscopic Light Curve Fits}
-We explored two different methods for fitting the spectroscopic light curves.
+We binned the spectrum into 20 spectrophotometric channels from FIXME to FIXME $\mu$m, shown in Figure FIXME.
+We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics are nearly constant as a function of wavelength \citep{stevenson??, kreidberg14a}.
+For this method, the transit model $T_\lambda(t)$ is multiplied by the systematics vector from the white light curve fit ($F_\mathrm{white}/T_\mathrm{white}$), and rescaled by a factor $C_\lambda + V_\lambda t$.
+An advantage of this approach is that it removes the star spot crossing feature, enabling us to use orbit three with no additional correction.
+There is a small chromatic effect due to the changing amplitude of the spot crossing feature with wavelength, but this effect is below the level of noise in our data FIXME.
+As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model.
+The fit has just three free parameters: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
+
+The best fit light curves have a median reduced $\chi_2$ value of FIXME and rms values consistent with the expected photon shot noise.
+There are two channels with slightly higher $\chi_2$ values: FIXME and FIXME.
+We therefore rescale the photometric uncertainties for all spectroscopic channels such that the $\chi_2$ values are unity, to ensure that we do not underestimate the uncertainty on the transit depths.
+We performed an MCMC fit to the light curves with \texttt{emcee}.
+The median transit depths and $1\,\sigma$ uncertainties are given in Table FIXME.
+FIXME: does the  median agree with the best fit?
+FIXME: convergence?
+
+We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum.
+We tested fitting the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (FIXME sigma), except offset by a constant value of FIXME ppm.
+We attribute the offset to the uncorrected star-spot crossing feature.
+Regardless of the cause, the offset does not affect our final analysis, because the planet's 1 bar radius is a free parameter in the atmospheric retrieval (see \S FIXME).
+We also explored fitting for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model predictions, and found that the `fitted limb darkening coefficients are consistent with the model.
+We opted to fix the coefficients in our final analysis because the uncertainties on the transit depths are slightly smaller (about 10\%).
+We also checked whether the uncertainty in the stellar parameters significantly affects the PHOENIX model predictions, and found that changing the stellar
+
+\subsubsection{Transit Depths Redder than FIXME $\mu$m}
+The red edge of the transmission spectrum is of interest for our analysis because methane becomes the dominant absorber over water at wavelengths $> 1.6\,\mu$m.
+Unfortunately, we find that the spectroscopic light curves between FIXME $\mu$m are too poor in quality to accurately measure the transit depth.
+The residuals show evidence for correlated noise and the reduced $\chi_2$ values increase to FIXME.
 
@@
 We also performed an independent data reduction and fit using K. Stevenson's pipeline.
-
-\begin{equation}
-x
-\end{equation}
-
-Star Spots?
-
-Confirmation from KBS
 

### 2017-07-05 ms.tex
fixing FIXMEs

@@
 \documentclass[twocolumn]{aastex61}
+\usepackage{natbib}
 
@@
 %% this information may be modified by the editorial offices.
-\shorttitle{\aastex\ Atmosphere of WASP-107b}
+\shorttitle{\aastex\ Atmosphere of WASP-107\lowercase{b}}
 \shortauthors{Kreidberg et al.}
@@
 
-\title{Water, High-Altitude Condensates, and No Methane in the Atmosphere of the Warm Neptune WASP-107b}
+\title{Water, High-Altitude Condensates, and No Methane in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}}
 
@@
 As is standard for observations of bright targets, we used the spatial scanning observing mode, which smears the spectrum in the spatial direction over the course of an exposure.
-The scan rate was FIXME arcseconds/sec.
+The scan rate was 0.12 arcseconds/sec.
 
 \section{Data Reduction}
-We reduced the data with the custom pipeline described in FIXME, which we summarize briefly here.
+We reduced the data with the custom pipeline described in \cite{kreidberg14a}, which we summarize briefly here.
 For each exposure, we extracted the spectrum from each up-the-ramp sample (or ``stripe") separately using the optimal extraction algorithm of \citep{horne86}.
@@
 To correct the change in dispersion solution over the length of the spatial scan, we interpolated each row to the wavelength scale of the row corresponding to the spectral trace.
-We also corrected for slight drift in the spectral direction (FIXME pixels) by interpolating each the first exposure.
+We also corrected for slight drift in the spectral direction ($<0.1$ pixels) by interpolating each the first exposure.
 To subtract the background, we took the median of sky pixels that were uncontaminated by the target spectrum (rows $5-250$, columns $5-15$).
 The typical background counts were low: 40 photoelectrons/pixel, in comparison to $3\times10^4$ photoelectrons/pixel in the stellar spectrum.
+%dispersion solution: 46 Angstroms/pixel. max drift: 3 angstroms (from diagnostics.txt)
 
@@
 \subsection{White Light Curve Fit}
-To create the raw white light curve, we summed each spectrum over the FIXME pixels in the spectral trace.
+To create the raw white light curve, we summed each spectrum over the 181 pixels in the spectral trace.
 The white light curve has systematic trends that are typical for WFC3 observations \cite{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend.
 The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.
-We fit the light curve with the analytic model of the form $F_\mathrm{white}(t) = S_\mathrm{white}(t)\times T_\mathrm{white}(t)$, where $S$ is a systematics model and $T$ is a transit model.
-We used the same systematics model as FIXME, equation FIXME.
+We fit the light curve with the analytic model of the form $F_\mathrm{white}(t) = S_\mathrm{white}(t)\times T_\mathrm{white}(t)$, where $S_\mathrm{white}$ is a systematics model and $T_\mathrm{white}$ is a transit model.
+We used the same systematics model as \cite{kreidberg15b}, Equation 1.
 To model the transit, we used the \texttt{batman} package \citep{kreidberg15b}.
-The transit model parameters are the orbital period $p$, time of FIXME $t_c$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.
+The transit model parameters are the orbital period $p$, time of inferior conjunction $t_0$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.
+
+\begin{figure*}
+\includegraphics[width = \textwidth]{Figures/lc.pdf}
+\caption{FIXME}
+\label{fig:lc}
+\end{figure*}
 
 \subsubsection{Star Spot Crossing}
-In our initial fit, we allowed all the transit parameters to vary freely.
-We noticed two odd results: first, there was a slight increase in flux during orbit three, and second, the transit parameters were inconsistent with the precise analysis of the Kepler light curve from \cite{dai17}.
+In our initial analysis, we noticed a bump in the light curve during orbit three that could not be fit with our model.
+We attribute this feature to a star spot crossing event, as WASP-107 is an active star and spot crossings have been observed before \citep{dai17, mocnik17}.
+In our subsequent analysis, we gave the data in orbit three no weight in the fit.
+The amplitude of the spot crossing feature is FIXME ppm, consistent with a spot of FIXME size and temperature.
+
+%In our initial fit, we allowed all the transit parameters to vary freely.
+We noticed two odd results: first, there was a slight increase in flux during orbit three, and second, the transit parameters were inconsistent with the analysis of the precise Kepler data from \cite{dai17}.
 We conjectured that the anomalies could be due to a star-spot crossing event in orbit three, as WASP-107 is an active star and spot crossings have been observed before FIXME.
@@
 \subsubsection{Final Fit}
-In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the values from \cite{dai17}.
+In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \cite{dai17}.
 We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star of FIXME properties, calculated with the FIXME package.
-The model values are consistent with what we obtain from fitting for the limb darkening coefficients as free parameters.
-The remaining free parameters were $t_c$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
+The remaining free parameters were $t_0$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
+%The model values are consistent with what we obtain from fitting for the limb darkening coefficients as free parameters.
 
-For the best fit light curve, the root-mean-square residuals were FIXME ppm (not including orbit three), compared to the expected shot noise of FIXME ppm.
+For the best fit light curve, the root-mean-square residuals were FIXME ppm (not including orbit three), somewhat larger than the expected shot noise of FIXME ppm.
 The achieved rms is close to the best value obtained for WFC3 data (FIXME), and we attribute the slight excess noise to either additional star spot crossings or loss of flux from the edge of the detector \citep[as has been seen for other bright targets;][]{FIXME}.
@@
 
-\software{astropy \citep{2013A&A...558A..33A},  
-          Cloudy \citep{2013RMxAA..49..137F}, 
-          SExtractor \citep{1996A&AS..117..393B}
-          }
+%\software{astropy \citep{2013A&A...558A..33A},  
+%          Cloudy \citep{2013RMxAA..49..137F}, 
+%          SExtractor \citep{1996A&AS..117..393B}
+%          }
 

### 2017-07-05 Figures/plot_lc.py

@@ -83,12 +83,10 @@ ax4 = plt.subplot(gs[1,1])
 #ax4 = ax3.twinx()
 
 rms = []
-sns.set_palette(sns.diverging_palette(220, 20, n=20))
+#sns.set_palette(sns.diverging_palette(220, 20, n=20))
+c = sns.diverging_palette(220, 20, n=20)
 ax4.text(2.8, 800,' rms', fontsize=10)
-#ax3.text(-3.5, 1,'$\lambda$ ($\mu$m)', fontsize=10)
 for i, f in enumerate(files):
-	sns.set_palette(sns.diverging_palette(220, 20, n=20))
-	#plt.plot(model.phase*per*24., np.ones_like(model.resid), color='0.5') 
 	p = pickle.load(open(f, 'rb'))
 	data, model = p[0], p[1]
 	ind = model.phase > 0.5
@@ -96,22 +94,20 @@ for i, f in enumerate(files):
 	
 	offset = -0.0022*float(i)*1e6
 	plt.axhline(offset, color ='0.5')
-	plt.plot(model.phase*per*24., model.resid/data.flux*1e6 + offset, marker='.', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5)
-#	ax4.text(-3.5, offset-1, '{0:0.2f}'.format(data.wavelength), fontsize=10)
+	plt.plot(model.phase*per*24., model.resid/data.flux*1e6 + offset, marker='.', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5, color = c[i])
 	ax4.text(2.8, offset - 1400, '{0:d}'.format(int(model.rms)), fontsize=10)
 	rms.append(model.rms)
 	ax4.plot(model.phase*per*24., np.ones_like(model.phase)*data.wavelength, linewidth=0.)
 	ax4.set_xlim(-3.2, 3.4)
-	#ax4.set_ylim(-4.5, .35)
 	ax4.set_ylim(-45000, 3500)
 	ax4.yaxis.set_major_locator(FixedLocator(np.array([-40000, -30000, -20000, -10000, 0])))
 	plt.gca().set_yticklabels(["0", "1e4", "2e4", "2e4", "3e4"])
 	ax4.set_xlabel("Time from central transit (hours)")
 	ax4.set_ylabel("Residuals (ppm, offset)") 
-	#ax4.set_ylabel("Wavelength ($\mu$m)")
 
 
[…]

### 2017-07-05 ms.tex
edits

@@
 \documentclass[twocolumn]{aastex61}
-\usepackage{natbib}
 
@@
 We used the same systematics model as \cite{kreidberg15b}, Equation 1.
-To model the transit, we used the \texttt{batman} package \citep{kreidberg15b}.
+To model the transit, we used the \texttt{batman} package \citep{kreidberg15a}.
 The transit model parameters are the orbital period $p$, time of inferior conjunction $t_0$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.
@@
 \subsubsection{Final Fit}
-In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \cite{dai17}.
+In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \citep{dai17}.
 We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star of FIXME properties, calculated with the FIXME package.
@@
 \section{Atmospheric Retrieval}
-We use the CHIMERA chemically-consistent transmission retrieval tool described in \cite{kreidberg15} determine the basic atmospheric properties.
-Briefly, the transmission spectrum solves the transmission geometry problem using the equations in \cite{brown11, tinetti12}.
+We use the CHIMERA chemically-consistent transmission retrieval tool described in \cite{kreidberg15b} to determine the basic atmospheric properties.
+Briefly, the transmission spectrum solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
 We parametrize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
@@
 In all three scenarios we fix the T-P profile "shape" but scale the irradiation temperature, which incorporates unknown variables like the albedo and heat transport efficiency.
-T-P profile shape information is not readily retrievable, nor does it strongly influence information derived from WFC3 transmission spectra \citep[e.g.][]{kreidberg15}.
+T-P profile shape information is not readily retrievable, nor does it strongly influence information derived from WFC3 transmission spectra \citep[e.g.][]{kreidberg15b}.
 We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
@@
 
-\bibliographystyle{apj}
+\bibliographystyle{aasjournal}
 \bibliography{ms.bib}

### 2017-07-24 ms.tex
new text/figs

@@
 
-\title{Water, High-Altitude Condensates, and No Methane in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}}
+\title{Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}}
 
@@
 
+We note that some directly imaged self-luminous planets in this
+temperature regime are also unexpectedly depleted in methane (e.g.,
+Skemer et al. 2014), when compared to similar temperature brown
+dwarfs.   The leading hypothesis of this depletion (Zahnle & Marley
+2015) is due to the CH4-CO quench point occurring in the CO stability
+field due to the low gravity of self-luminous planets when compared to
+brown dwarfs.  In order this mechanism to work for neptune massed
+planet, the internal temperature would have to be near ~500K to get
+the deep temperatures high enough to thermochemically deplete methane
+at the quench point. Given the age of the star, it is unlikely that
+such high internal temperatures are still due to heat of formation. It
+is possible that tidal heating could increase the internal temperature
+(Agundez et al. 2014; Morley et al. 2017) but that would require a
+negligible eccentricity, of which is currently uncertain for WASP107b.
+
 \subsection{Methane}
@@
 \subsection{Atmospheric Metallicity Predictions from Interior Structure Modeling}
+ We determine the bulk metallicity of WASP-107 b using the structure evolution modeling of Thorngren et al.
+(2016).
+These consider a thermally inert heavy-element core with a convective envelope of additively mixed H/He (Saumon et al.
+1995) and heavy-element impurities.
+The heavy-elements were a 50-50 rock-ice mix from ANEOS (Thompson 1990).
+Planets were evolved in time using the atmospheric models of Fortney (2007).
+We explored possible envelope metallicities and core masses using a metropolis-hastings MCMC with M = .12 +/- .03, R = .94 +/- .02, and the age as alternatively .2-1.0 Gyr or 1.0-13.8 Gyr.
+Priors were $\propto 1$ in envelope metal mass fraction and core mass.
+The MCMC burned in for 1000 steps and then collected 4,000,000 samples, thinned down to 100,000.
+The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was about 18 (the value for water), using the approach of Fortney (2013).
 

### 2017-07-24 Analysis/get_COratio.py

@@ -1,0 +1,66 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pickle
+from pylab import hist
+import scipy.stats as st
+
+
+def get_significance(alpha):
+	z = st.norm.ppf(1.-alpha)
+	print "should I be dividing alpha by 2?"
+	return z
+
+"""alpha = np.linspace(0.001, 0.999, 100)
+plt.plot(alpha, st.norm.ppf(alpha))
+plt.show()"""
+
+solarctoo = np.log10(0.54)
+
+plt.subplot(221)
+p = pickle.load(open("Global_Cloud_samples.pic", "rb"))
+ctoo = p[:,2]
+hist(ctoo)
+plt.axvline(np.log10(0.54), color = 'y')
+plt.title("CC")
+alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
+print "CC sig", get_significance(alpha)
+
+plt.subplot(222)
+p = pickle.load(open("Global_Cloud_noCH4_samples.pic", "rb"))
+ctoo = p[:,2]
+hist(ctoo)
+plt.axvline(np.log10(0.54), color = 'y')
+alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
+plt.title("CC no CH4")
+print "CC no CH4 sig", get_significance(alpha)
+
+plt.subplot(223)
+p = pickle.load(open("MCMC_CC_QUENCH.pic", "rb"))
+p = p[:,:-1]
+ctoo = p[:,2]
+#plt.clf()
+#plt.plot(ctoo)
+#plt.show()
+hist(ctoo)
+plt.axvline(np.log10(0.54), color = 'y')
+alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
+plt.title("CC quench")
+print "CC quench", get_significance(alpha)
+
+plt.subplot(224)
+p = pickle.load(open("MCMC_CC_Limit_Metallicity.pic", "rb"))
+p = p[:,:-1]
+ctoo = p[:,2]
+hist(ctoo)
+alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
+plt.axvline(np.log10(0.54), color = 'y')
+plt.title("CC Z")
+print "CC Z", get_significance(alpha)
+plt.show()
+
+p = pickle.load(open("MCMC_Free.pic", "rb"))
+p = p[:,:-1]
+x = p[:,4]
+hist(x)
+plt.title("FREE")
+plt.show()

### 2017-07-24 Figures/fig2_spectrum.py

@@ -1,0 +1,46 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pickle
+import seaborn as sns
+
+sns.set_context("notebook", font_scale = 1.2)
+sns.set_style("white")
+sns.set_style("ticks", {"xtick.direction":"in", "ytick.direction":"in"})
+
+
+#pickle = hi_res_wlgrid, hi_res_-2sigma, hi_res_-1sigma, hi_res_median, hi_res_+1sigma, hi_res+2sigma, data_wl_grid, raw_data, data_err, median_binned_model
+
+d= np.genfromtxt("w107_spectrum_06-22-17.txt")
+plt.errorbar(d[:,0], d[:,1]*100., d[:,2]*100., linestyle = 'none', zorder = 100, marker = 'o', color = 'w', markeredgecolor = 'k', ecolor ='k', markeredgewidth = 1.3)
+
+
+
+
+p = pickle.load(open("best_fit.pic", "rb"))
+w = p[0]
+med = p[1]
+plt.fill_between(w, med*99.8, med*100.2, color = "navy", alpha = '0.3')
+plt.plot(w, med*100., color = 'navy', label = "Best fit model")
+
+
+offset = np.median(med)
+plt.plot(w, ((med-offset)*4 + offset)*100., color = '0.5', alpha = 0.5, label = "1X solar, cloud-free", zorder = -20) 
+plt.legend(loc= 'upper right')
+
+plt.ylabel("Transit depth (%)")
+plt.ylim(2.03, 2.1)
+plt.xlabel("Wavelength (microns)")
+
+x2 = plt.gca().twinx()
+scaleheight = 1000./1.e6
+plt.plot(w,med/scaleheight - np.min(med/scaleheight), linewidth=0.)
+
+plt.ylabel("Scale heights")
+plt.xlim(1, 1.8)
+
+
+
+plt.tight_layout()
+plt.savefig("spectrum.pdf")
+plt.show()
+

### 2017-07-24 Analysis/get_COratio.py
fancy fig

@@ -55,7 +55,7 @@ hist(ctoo)
 alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
 plt.axvline(np.log10(0.54), color = 'y')
 plt.title("CC Z")
-print "CC Z", get_significance(alpha)
+print "CC Z", get_significance(alpha), alpha
 plt.show()
 
 p = pickle.load(open("MCMC_Free.pic", "rb"))

### 2017-07-24 Figures/fig2_spectrum.py

@@ -14,26 +14,26 @@ d= np.genfromtxt("w107_spectrum_06-22-17.txt")
 plt.errorbar(d[:,0], d[:,1]*100., d[:,2]*100., linestyle = 'none', zorder = 100, marker = 'o', color = 'w', markeredgecolor = 'k', ecolor ='k', markeredgewidth = 1.3)
 
 
-
-
-p = pickle.load(open("best_fit.pic", "rb"))
+p = pickle.load(open("WASP107b_Global_Cloud_limit_Metallicity_spectra.pic", "rb"))
 w = p[0]
-med = p[1]
-plt.fill_between(w, med*99.8, med*100.2, color = "navy", alpha = '0.3')
-plt.plot(w, med*100., color = 'navy', label = "Best fit model")
+
+plt.fill_between(w, p[1], p[5], color = "navy", alpha = '0.2')
+plt.fill_between(w, p[2], p[4], color = "navy", alpha = '0.3')
+plt.plot(w, p[3], color = 'navy', label = "Best fit model")
 
 
-offset = np.median(med)
-plt.plot(w, ((med-offset)*4 + offset)*100., color = '0.5', alpha = 0.5, label = "1X solar, cloud-free", zorder = -20) 
-plt.legend(loc= 'upper right')
+"""d = pickle.load(open("Model_Solar.pic", "rb"))
+x, y = d[0], d[1]*100. - 0.7
+plt.plot(x, y, color = '0.5', alpha = 0.5, label = "1X solar, cloud-free", zorder = -20) 
+plt.legend(loc= 'upper right')"""
 
 plt.ylabel("Transit depth (%)")
 plt.ylim(2.03, 2.1)
 plt.xlabel("Wavelength (microns)")
 
 x2 = plt.gca().twinx()
-scaleheight = 1000./1.e6
-plt.plot(w,med/scaleheight - np.min(med/scaleheight), linewidth=0.)
+scaleheight = 460./1.e6*100.
+plt.plot(w,p[3]/scaleheight - np.min(p[3]/scaleheight), linewidth=0.)
 
 plt.ylabel("Scale heights")
 plt.xlim(1, 1.8)

### 2017-08-05 ms.tex
corrected FIXMEs

@@
-%%
-%% Beginning of file 'sample61.tex'
-%%
-%% Modified 2016 September
-%%
-%% This is a sample manuscript marked up using the
-%% AASTeX v6.1 LaTeX 2e macros.
-%%
-%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
-%% (Copyright 2000-2015).  See the classfile for details.
-
-%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
-%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
-%% All of these external packages should already be present in the modern TeX 
-%% distributions.  If not they can also be obtained at www.ctan.org.
-
-%% The first piece of markup in an AASTeX v6.x document is the \documentclass
-%% command. LaTeX will ignore any data that comes before this command. The 
-%% documentclass can take an optional argument to modify the output style.
-%% The command below calls the preprint style  which will produce a tightly 
-%% typeset, one-column, single-spaced document.  It is the default and thus
-%% does not need to be explicitly stated.
-%%
-%%
-%% using aastex version 6.1
 \documentclass[twocolumn]{aastex61}
 
-%% The default is a single spaced, 10 point font, single spaced article.
-%% There are 5 other style options available via an optional argument. They
-%% can be envoked like this:
-%%
-%% \documentclass[argument]{aastex61}
-%% 
-%% where the arguement options are:
-%%
-%%  twocolumn   : two text columns, 10 point font, single spaced article.
-%%                This is the most compact and represent the final published
-%%                derived PDF copy of the accepted manuscript from the publisher
-%%  manuscript  : one text column, 12 point font, double spaced article.
-%%  preprint    : one text column, 12 point font, single spaced article.  
-%%  preprint2   : two text columns, 12 point font, single spaced article.
-%%  modern      : a stylish, single text column, 12 point font, article with
-%% 		  wider left and right margins. This uses the Daniel
-%% 		  Foreman-Mackey and David Hogg design.
-%%
-%% Note that you can submit to the AAS Journals in any of these 6 styles.
-%%
-%% There are other optional arguments one can envoke to allow other stylistic
-%% actions. The available options are:
-%%
-%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
-%%  tighten      : Makes baselineskip slightly smaller, only works with 
-%%                 the twocolumn substyle.
-%%  times        : uses times font instead of the default
-%%  linenumbers  : turn on lineno package.
-%%  trackchanges : required to see the revision mark up and print its output
-%%  longauthor   : Do not use the more compressed footnote style (default) for 
-%%                 the author/collaboration/affiliations. Instead print all
-%%                 affiliation information after each name. Creates a much
-%%                 long author list but may be desirable for short author papers
-%%
-%% these can be used in any combination, e.g.
-%%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex61}
-
-%% AASTeX v6.* now includes \hyperref support. While we have built in specific
-%% defaults into the classfile you can manually override them with the
-%% \hypersetup command. For example,
-%%
-%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-%%
-%% will change the color of the internal links to red, the links to the
-%% bibliography to green, the file links to cyan, and the external links to
-%% magenta. Additional information on \hyperref options can be found here:
-%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
-
-%% If you want to create your own macros, you can do so
-%% using \newcommand. Your macros should appear before
-%% the \begin{document} command.
-%%
 \newcommand{\vdag}{(v)^\dagger}
@@
 %\accepted{\today}
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the arguement.
 \submitjournal{ApJL}
 
-%% Mark up commands to limit the number of authors on the front page.
-%% Note that in AASTeX v6.1 a \collaboration call (see below) counts as
-%% an author in this case.
-%
-%\AuthorCollaborationLimit=3
-%
-%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
-%% collaboration" on the front page of this example manuscript.
-%%
-%% Note that all of the author will be shown in the published article.
-%% This feature is meant to be used prior to acceptance to make the
-%% front end of a long author article more manageable. Please do not use
-%% this functionality for manuscripts with less than 20 authors. Conversely,
-%% please do use this when the number of authors exceeds 40.
-%%
-%% Use \allauthors at the manuscript end to show the full author list.
-%% This command should only be used with \AuthorCollaborationLimit is used.
-
-%% The following command can be used to set the latex table counters.  It
-%% is needed in this document because it uses a mix of latex tabular and
-%% AASTeX deluxetables.  In general it should not be needed.
-%\setcounter{table}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% The following section outlines numerous optional output that
-%% can be displayed in the front matter or as running meta-data.
-%%
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-\shorttitle{\aastex\ Atmosphere of WASP-107\lowercase{b}}
+\shorttitle{Atmosphere Characterization of WASP-107\lowercase{b}}
 \shortauthors{Kreidberg et al.}
-%%
-%% You can add a light gray and diagonal water-mark to the first page 
-%% with this command:
-% \watermark{text}
-%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
-%% long you can control the water-mark size with:
-%  \setwatermarkfontsize{dimension}
-%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
 
@@
 
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire. In v6.1 you can include a footnote in the title.
-
-%% A significant change from earlier AASTEX versions is in the structure for 
-%% calling author and affilations. The change was necessary to implement 
-%% autoindexing of affilations which prior was a manual process that could 
-%% easily be tedious in large author manuscripts.
-%%
-%% The \author command is the same as before except it now takes an optional
-%% arguement which is the 16 digit ORCID. The syntax is:
-%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
-%%
-%% This will hyperlink the author name to the author's ORCID page. Note that
-%% during compilation, LaTeX will do some limited checking of the format of
-%% the ID to make sure it is valid.
-%%
-%% Use \affiliation for affiliation information. The old \affil is now aliased
-%% to \affiliation. AASTeX v6.1 will automatically index these in the header.
-%% When a duplicate is found its index will be the same as its previous entry.
-%%
-%% Note that \altaffilmark and \altaffiltext have been removed and thus 
-%% can not be used to document secondary affiliations. If they are used latex
-%% will issue a specific error message and quit. Please use multiple 
-%% \affiliation calls for to document more than one affiliation.
-%%
-%% The new \altaffiliation can be used to indicate some secondary information
-%% such as fellowships. This command produces a non-numeric footnote that is
-%% set away from the numeric \affiliation footnotes.  NOTE that if an
-%% \altaffiliation command is used it must come BEFORE the \affiliation call,
-%% right after the \author command, in order to place the footnotes in
-%% the proper location.
-%%
-%% Use \email to set provide email addresses. Each \email will appear on its
-%% own line so you can put multiple email address in one \email call. A new
-%% \correspondingauthor command is available in V6.1 to identify the
-%% corresponding author of the manuscript. It is the author's responsibility
-%% to make sure this name is also in the author list.
-%%
-%% While authors can be grouped inside the same \author and \affiliation
-%% commands it is better to have a single author for each. This allows for
-%% one to exploit all the new benefits and should make book-keeping easier.
-%%
-%% If done correctly the peer review system will be able to
-%% automatically put the author and affiliation information from the manuscript
-%% and save the corresponding author the trouble of entering it by hand.
-
 \correspondingauthor{Laura Kreidberg}
@@
 
-\author{Laura Kreidberg, Michael Line}
+\author{Laura Kreidberg}
 \affiliation{Harvard Society of Fellows\
@@
 \author{Caroline Morley}
+\altaffiliation{Sagan Fellow}
 \affiliation{Harvard-Smithsonian Center for Astrophysics}
-\affiliation{Sagan Fellow}
 \nocollaboration
@@
 
-%% Note that the \and command from previous versions of AASTeX is now
-%% depreciated in this version as it is no longer necessary. AASTeX 
-%% automatically takes care of all commas and "and"s between authors names.
-
-%% AASTeX 6.1 has the new \collaboration and \nocollaboration commands to
-%% provide the collaboration status of a group of authors. These commands 
-%% can be used either before or after the list of corresponding authors. The
-%% argument for \collaboration is the collaboration identifier. Authors are
-%% encouraged to surround collaboration identifiers with ()s. The 
-%% \nocollaboration command takes no argument and exists to indicate that
-%% the nearby authors are not part of surrounding collaborations.
-
-%% Mark off the abstract in the ``abstract'' environment. 
 \begin{abstract}
-
-WASP-107b is baller.
-
+We report a measurement of the near-infrared transmission spectrum of the warm Neptune exoplanet WASP-107b measured with the Hubble Space Telescope Wide Field Camera 3.
+We detecta water absorption feature (FIXME confidence).
+The feature is small amplitude, clouds.
+We combine an interior structure model of the planet with an atmospheric retrieval .
+We find an upper limit on the atmospheric metallicity of FIXME solar, a cloud deck .
 \end{abstract}
@@
 We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{https://github.com/spacetelescope/PandExo\_HST}.
-As is standard for observations of bright targets, we used the spatial scanning observing mode, which smears the spectrum in the spatial direction over the course of an exposure.
+As is standard for observations of bright targets, we used the spatial scanning observing mode, which slews the telescope in the spatial direction over the course of an exposure.
 The scan rate was 0.12 arcseconds/sec.
@@
 We reduced the data with the custom pipeline described in \cite{kreidberg14a}, which we summarize briefly here.
-For each exposure, we extracted the spectrum from each up-the-ramp sample (or ``stripe") separately using the optimal extraction algorithm of \citep{horne86}.
+For each exposure, we extracted the spectrum from each up-the-ramp sample (or ``stripe") separately using the optimal extraction algorithm of \cite{horne86}.
 The stripe spectra were then summed to create the final spectrum.
@@
 To create the raw white light curve, we summed each spectrum over the 181 pixels in the spectral trace.
-The white light curve has systematic trends that are typical for WFC3 observations \cite{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend.
+The white light curve has systematic trends that are typical for WFC3 observations \citep{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend.
 The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.
@@
 In our subsequent analysis, we gave the data in orbit three no weight in the fit.
-The amplitude of the spot crossing feature is FIXME ppm, consistent with a spot of FIXME size and temperature.
-
-%In our initial fit, we allowed all the transit parameters to vary freely.
-We noticed two odd results: first, there was a slight increase in flux during orbit three, and second, the transit parameters were inconsistent with the analysis of the precise Kepler data from \cite{dai17}.
-We conjectured that the anomalies could be due to a star-spot crossing event in orbit three, as WASP-107 is an active star and spot crossings have been observed before FIXME.
-To test this idea, we gave the data in orbit three no weight in the fit.
-For this case, the transit parameters agree well with those of \cite{dai17} and the spot-crossing feature is more pronounced (see Figure FIXME).
-The amplitude of the feature is FIXME ppm, consistent with a spot of FIXME size and temperature (cf dai17 FIXME).
-We conclude that the feature is indeed due to a spot crossing.
+The amplitude of the spot crossing feature is 300 ppm.
 
@@
 In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \citep{dai17}.
-We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star of FIXME properties, calculated with the FIXME package.
+These values are consistent with our best fit when we allowed them to vary freely (and masked the star spot).
+We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star with effective temperature 4300 K, calculated with the \texttt{limb-darkening} package from \cite{espinoza15}.
 The remaining free parameters were $t_0$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
@@
 
-For the best fit light curve, the root-mean-square residuals were FIXME ppm (not including orbit three), somewhat larger than the expected shot noise of FIXME ppm.
-The achieved rms is close to the best value obtained for WFC3 data (FIXME), and we attribute the slight excess noise to either additional star spot crossings or loss of flux from the edge of the detector \citep[as has been seen for other bright targets;][]{FIXME}.
-We used the Markov chain Monte Carlo algorithm to estimate parameter uncertainties (as implemented in the the \texttt{emcee} package; FIXME).
-We scaled the per-point uncertainties by a factor of FIXME to achieve a reduced $\chi_2$ value of unity.
-The transit time was FIXME and the planet/star radius was FIXME.
+For the best fit white light curve, the root-mean-square (rms) residuals were 93 ppm (excluding the star spot crossing), which is somewhat larger than the expected shot noise of 50 ppm.
+We attribute the excess noise to loss of flux off the edge of the detector, which can occur if there is variation in the position or length of the spatial scan.
+There is no evidence for correlated noise in the residuals, so we increased the per-point uncertainties by a factor of 1.7 to achieve a reduced $\chi_2$ value of unity.
+We used the Markov chain Monte Carlo algorithm to estimate parameter uncertainties \citep{foremanmackey13}.
+The transit time was $t_0 = 2457910.45407\pm6\mathrm{e}{-5}$ and the planet/star radius was $r_p/r_s = 0.14399\pm0.00017$.
 
@@
 
+\begin{figure*}
+\includegraphics[width = \textwidth]{Figures/spectrum.pdf}
+\caption{FIXME}
+\label{fig:spectrum}
+\end{figure*}
+
 \section{Atmospheric Retrieval}
@@
 
-We note that some directly imaged self-luminous planets in this
-temperature regime are also unexpectedly depleted in methane (e.g.,
-Skemer et al. 2014), when compared to similar temperature brown
-dwarfs.   The leading hypothesis of this depletion (Zahnle & Marley
-2015) is due to the CH4-CO quench point occurring in the CO stability
-field due to the low gravity of self-luminous planets when compared to
-brown dwarfs.  In order this mechanism to work for neptune massed
-planet, the internal temperature would have to be near ~500K to get
-the deep temperatures high enough to thermochemically deplete methane
-at the quench point. Given the age of the star, it is unlikely that
-such high internal temperatures are still due to heat of formation. It
-is possible that tidal heating could increase the internal temperature
-(Agundez et al. 2014; Morley et al. 2017) but that would require a
-negligible eccentricity, of which is currently uncertain for WASP107b.
+We note that some directly imaged self-luminous planets in this temperature regime are also unexpectedly depleted in methane (e.g., Skemer et al.
+2014), when compared to similar temperature brown dwarfs.
+The leading hypothesis of this depletion \citep{zahnle15} is due to the CH4-CO quench point occurring in the CO stability field due to the low gravity of self-luminous planets when compared to brown dwarfs.
+In order this mechanism to work for neptune massed planet, the internal temperature would have to be near ~500K to get the deep temperatures high enough to thermochemically deplete methane at the quench point.
+Given the age of the star, it is unlikely that such high internal temperatures are still due to heat of formation.
+It is possible that tidal heating could increase the internal temperature (Agundez et al.
+2014; Morley et al.
+2017) but that would require a negligible eccentricity, of which is currently uncertain for WASP107b.
 

### 2017-08-05 Analysis/get_COratio.py

@@ -56,6 +56,14 @@ alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
 plt.axvline(np.log10(0.54), color = 'y')
 plt.title("CC Z")
 print "CC Z", get_significance(alpha), alpha
+
+print "3 sigma upper limit on metallicity from CC Z case:"
+ind = np.argsort(ctoo)
+ctoo = ctoo[ind]
+upperlim = int(np.ceil(0.95*len(ctoo)))
+print 10.**ctoo[upperlim]
+
+
 plt.show()
 
 p = pickle.load(open("MCMC_Free.pic", "rb"))

### 2017-08-05 Figures/fig2_spectrum.py

@@ -2,17 +2,23 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pickle
 import seaborn as sns
+from astropy.convolution import convolve
+from astropy.convolution import Gaussian1DKernel
 
 sns.set_context("notebook", font_scale = 1.2)
 sns.set_style("white")
 sns.set_style("ticks", {"xtick.direction":"in", "ytick.direction":"in"})
 
+g = Gaussian1DKernel(stddev=10)
 
+plt.figure(figsize=(8,4))
 #pickle = hi_res_wlgrid, hi_res_-2sigma, hi_res_-1sigma, hi_res_median, hi_res_+1sigma, hi_res+2sigma, data_wl_grid, raw_data, data_err, median_binned_model
 
 d= np.genfromtxt("w107_spectrum_06-22-17.txt")
 plt.errorbar(d[:,0], d[:,1]*100., d[:,2]*100., linestyle = 'none', zorder = 100, marker = 'o', color = 'w', markeredgecolor = 'k', ecolor ='k', markeredgewidth = 1.3)
 
+#d = np.genfromtxt("wasp107b_10xsolar_0.1mbarcloud.dat")
+#plt.plot(d[:,0]*1e6, convolve(d[:,1], g, boundary = 'extend')-0.08)
 
 p = pickle.load(open("WASP107b_Global_Cloud_limit_Metallicity_spectra.pic", "rb"))
 w = p[0]
@@ -20,7 +26,6 @@ w = p[0]
 plt.fill_between(w, p[1], p[5], color = "navy", alpha = '0.2')
 plt.fill_between(w, p[2], p[4], color = "navy", alpha = '0.3')
 plt.plot(w, p[3], color = 'navy', label = "Best fit model")
-
 
 """d = pickle.load(open("Model_Solar.pic", "rb"))
 x, y = d[0], d[1]*100. - 0.7
@@ -36,7 +41,7 @@ scaleheight = 460./1.e6*100.
 plt.plot(w,p[3]/scaleheight - np.min(p[3]/scaleheight), linewidth=0.)
 
 plt.ylabel("Scale heights")
-plt.xlim(1, 1.8)
+plt.xlim(1.05, 1.7)
 
 
 

### 2017-08-05 Figures/plot_lc.py
white lc rms

@@ -46,7 +46,8 @@ x = np.linspace(0.1, 1.1)
 ax2.fill_between(x, -1000, 1000, color = '0.7', zorder=-10)
 plt.axhline(0., color ='0.5')
 plt.plot(model.phase*per*24., model.resid/data.flux*1e6, marker='.', color = 'w', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5)
-ax2.text(1.8, 210., 'rms (ppm):\n' + '{0:d}'.format(int(model.rms*1e6)), fontsize=10)
+#ax2.text(1.8, 210., 'rms (ppm):\n' + '{0:d}'.format(int(model.rms*1e6)), fontsize=10)
+ax2.text(1.7, 300., 'rms: 50 ppm', fontsize=10)
 ax2.plot(model.phase*per*24., np.ones_like(model.phase)*data.wavelength, linewidth=0.)
 ax2.set_ylim(-500, 700)
 ax2.set_xlim(-3.2, 3.4)

### 2017-08-05 ms.tex
table

@@
 We attribute the excess noise to loss of flux off the edge of the detector, which can occur if there is variation in the position or length of the spatial scan.
-There is no evidence for correlated noise in the residuals, so we increased the per-point uncertainties by a factor of 1.7 to achieve a reduced $\chi_2$ value of unity.
-We used the Markov chain Monte Carlo algorithm to estimate parameter uncertainties \citep{foremanmackey13}.
+There is no evidence for correlated noise in the residuals, so to account for the excess noise we simply increased the per-point uncertainties by a factor of 1.7 to achieve a reduced $\chi_2$ value of unity.
+We then used the Markov chain Monte Carlo algorithm to estimate parameter uncertainties \citep{foremanmackey13}.
 The transit time was $t_0 = 2457910.45407\pm6\mathrm{e}{-5}$ and the planet/star radius was $r_p/r_s = 0.14399\pm0.00017$.
@@
 \subsection{Spectroscopic Light Curve Fits}
-We binned the spectrum into 20 spectrophotometric channels from FIXME to FIXME $\mu$m, shown in Figure FIXME.
+We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.65 $\mu$m, shown in Figure\,\ref{fig:lc}.
 We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics are nearly constant as a function of wavelength \citep{stevenson??, kreidberg14a}.
@@
 An advantage of this approach is that it removes the star spot crossing feature, enabling us to use orbit three with no additional correction.
-There is a small chromatic effect due to the changing amplitude of the spot crossing feature with wavelength, but this effect is below the level of noise in our data FIXME.
+There is a small chromatic effect due to the changing amplitude of the spot crossing feature with wavelength, but this effect is not detectable in our data.
 As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model.
@@
 
-The best fit light curves have a median reduced $\chi_2$ value of FIXME and rms values consistent with the expected photon shot noise.
-There are two channels with slightly higher $\chi_2$ values: FIXME and FIXME.
-We therefore rescale the photometric uncertainties for all spectroscopic channels such that the $\chi_2$ values are unity, to ensure that we do not underestimate the uncertainty on the transit depths.
+The best fit light curves have a median reduced $\chi_2$ value of 1.16.
+To ensure that we do not underestimate the uncertainty on the transit depths, we rescale the photometric uncertainties for all spectroscopic channels such that the $\chi_2$ values are unity.
 We performed an MCMC fit to the light curves with \texttt{emcee}.
 The median transit depths and $1\,\sigma$ uncertainties are given in Table FIXME.
-FIXME: does the  median agree with the best fit?
-FIXME: convergence?
 
@@
 
-\subsection{Independent Analysis}
+\subsubsection{Independent Analysis}
 We also performed an independent data reduction and fit using K. Stevenson's pipeline.
@@
 \end{figure*}
+
+\begin{deluxetable}{c c c}
+\tablecaption{WASP-107b transmission spectrum \label{transit_depths}}
+\tablehead{\colhead{Wavelength ($\mu$m)}  & \colhead{Transit depth} & \colhead{Uncertainty}}
+\startdata
+1.133 & 0.020641 & 5.9e-05 \\
+1.158 & 0.020733 & 5.5e-05 \\
+1.184 & 0.020505 & 5.6e-05 \\
+1.209 & 0.020498 & 5.4e-05 \\
+1.235 & 0.020455 & 5.9e-05 \\
+1.260 & 0.020492 & 5.0e-05 \\
+1.285 & 0.020620 & 6.2e-05 \\
+1.311 & 0.020739 & 5.0e-05 \\
+1.336 & 0.020660 & 5.7e-05 \\
+1.362 & 0.020858 & 4.8e-05 \\
+1.387 & 0.020794 & 4.8e-05 \\
+1.413 & 0.020888 & 5.2e-05 \\
+1.438 & 0.020821 & 6.2e-05 \\
+1.464 & 0.020691 & 5.1e-05 \\
+1.489 & 0.020682 & 6.9e-05 \\
+1.515 & 0.020679 & 6.7e-05 \\
+1.540 & 0.020509 & 6.0e-05 \\
+1.565 & 0.020480 & 6.4e-05 \\
+1.591 & 0.020500 & 5.6e-05 \\
+1.616 & 0.020514 & 6.5e-05 \\
+\enddata
+\end{deluxetable}
 

### 2017-08-05 Figures/format_table.py

@@ -1,0 +1,10 @@
+import numpy as np
+
+d = np.genfromtxt("w107_spectrum_06-22-17.txt")
+
+n = len(d[:,0])
+
+print np.mean(d[:,1])
+for i in range(n):
+#	print '{0:0.3f}'.format(d[i,0]), '&', '{0:0.0f}'.format((d[i,1] - np.mean(d[:,1]))*1.e6), '&', '{0:0.0f}'.format(d[i,2]*1e6), '\\\\' 
+	print '{0:0.3f}'.format(d[i,0]), '&', '{0:0.6f}'.format(d[i,1]), '&', '{0:0.1e}'.format(d[i,2]), '\\\\' 

### 2017-08-05 Figures/plot_lc.py

@@ -7,10 +7,11 @@ from pylab import *
 import seaborn as sns
 import matplotlib.gridspec as gridspec
 
-sns.set_context("talk")
+sns.set_context("talk", rc={"lines.linewidth":1.5})
 sns.set_style("white")
 sns.set_style("ticks", {"xtick.direction":"in", "ytick.direction":"in"})
-sns.set_palette(sns.diverging_palette(220, 20, n=20))
+#sns.set_palette(sns.diverging_palette(220, 20, n=20))
+sns.set_palette(sns.diverging_palette(40, 20, n=20, s=90))
 
 path =  "lsq_fits"
 files = glob.glob(os.path.join(path, "*.p"))	
@@ -33,6 +34,7 @@ model.phase_hr, model.lc_hr = model.phase_hr[ind], model.lc_hr[ind]
 ax1.plot(model.phase_hr*per*24., model.lc_hr , color = '0.5')
 ax1.plot(model.phase*per*24., model.data_corr, marker='.', color = 'w', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5)
 ax1.set_xlim(-3.2, 3.4)
+ax1.set_ylim(0.973, 1.005)
 ax1.yaxis.set_major_locator(FixedLocator(np.array([0.98, 0.99, 1.0])))
 ax1.set_yticklabels(["0.98", "0.99", "1.0"])
 
@@ -47,7 +49,7 @@ ax2.fill_between(x, -1000, 1000, color = '0.7', zorder=-10)
 plt.axhline(0., color ='0.5')
 plt.plot(model.phase*per*24., model.resid/data.flux*1e6, marker='.', color = 'w', ms = 8, linestyle = 'None', markeredgecolor='k', markeredgewidth=0.5)
 #ax2.text(1.8, 210., 'rms (ppm):\n' + '{0:d}'.format(int(model.rms*1e6)), fontsize=10)
-ax2.text(1.7, 300., 'rms: 50 ppm', fontsize=10)
+ax2.text(1.7, 300., 'rms: 92 ppm', fontsize=10)
 ax2.plot(model.phase*per*24., np.ones_like(model.phase)*data.wavelength, linewidth=0.)
 ax2.set_ylim(-500, 700)
 ax2.set_xlim(-3.2, 3.4)
@@ -85,7 +87,8 @@ ax4 = plt.subplot(gs[1,1])
 
 rms = []
[…]

### 2017-08-06 ms.tex
FIXMEs

@@
 
-\subsection{White Light Curve Fit}
+\subsection{White Light Curve}
 To create the raw white light curve, we summed each spectrum over the 181 pixels in the spectral trace.
@@
 \subsubsection{Star Spot Crossing}
-In our initial analysis, we noticed a bump in the light curve during orbit three that could not be fit with our model.
+In our initial analysis, we noticed a bump in the light curve during orbit three that was not fit well with our model.
 We attribute this feature to a star spot crossing event, as WASP-107 is an active star and spot crossings have been observed before \citep{dai17, mocnik17}.
 In our subsequent analysis, we gave the data in orbit three no weight in the fit.
-The amplitude of the spot crossing feature is 300 ppm.
+The amplitude of the spot crossing feature is 300 ppm, as illustrated in Figure\,\ref{fig:lc}.
 
@@
 We attribute the excess noise to loss of flux off the edge of the detector, which can occur if there is variation in the position or length of the spatial scan.
-There is no evidence for correlated noise in the residuals, so to account for the excess noise we simply increased the per-point uncertainties by a factor of 1.7 to achieve a reduced $\chi_2$ value of unity.
-We then used the Markov chain Monte Carlo algorithm to estimate parameter uncertainties \citep{foremanmackey13}.
+There is no evidence for correlated noise in the residuals, so to account for the excess noise we simply increased the per-point uncertainties by a factor of 1.7 to achieve a reduced $\chi^2$ value of unity.
+We then used the Markov chain Monte Carlo (MCMC) algorithm to estimate parameter uncertainties \citep{foremanmackey13}.
 The transit time was $t_0 = 2457910.45407\pm6\mathrm{e}{-5}$ and the planet/star radius was $r_p/r_s = 0.14399\pm0.00017$.
+FIXME: converged?
 
@@
 We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.65 $\mu$m, shown in Figure\,\ref{fig:lc}.
-We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics are nearly constant as a function of wavelength \citep{stevenson??, kreidberg14a}.
+We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics are nearly constant as a function of wavelength \citep{stevenson14c, kreidberg14a}.
 For this method, the transit model $T_\lambda(t)$ is multiplied by the systematics vector from the white light curve fit ($F_\mathrm{white}/T_\mathrm{white}$), and rescaled by a factor $C_\lambda + V_\lambda t$.
@@
 As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model.
-The fit has just three free parameters: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
+The fit had just three free parameters: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
 
-The best fit light curves have a median reduced $\chi_2$ value of 1.16.
-To ensure that we do not underestimate the uncertainty on the transit depths, we rescale the photometric uncertainties for all spectroscopic channels such that the $\chi_2$ values are unity.
+The best fit light curves have a median reduced $\chi^2$ value of 1.16.
+To ensure that we did not underestimate the uncertainty on the transit depths, we rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2$ values are unity.
 We performed an MCMC fit to the light curves with \texttt{emcee}.
-The median transit depths and $1\,\sigma$ uncertainties are given in Table FIXME.
+The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:transit_depths}.
+FIXME: converged?
 
 We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum.
-We tested fitting the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (FIXME sigma), except offset by a constant value of FIXME ppm.
+For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (differing by just $0.3\sigma$ on average), except with a constant offset of 350 ppm.
 We attribute the offset to the uncorrected star-spot crossing feature.
-Regardless of the cause, the offset does not affect our final analysis, because the planet's 1 bar radius is a free parameter in the atmospheric retrieval (see \S FIXME).
-We also explored fitting for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model predictions, and found that the `fitted limb darkening coefficients are consistent with the model.
-We opted to fix the coefficients in our final analysis because the uncertainties on the transit depths are slightly smaller (about 10\%).
-We also checked whether the uncertainty in the stellar parameters significantly affects the PHOENIX model predictions, and found that changing the stellar
+Regardless of the cause, the offset does not affect our final analysis, because the planet's one-bar radius is a free parameter in the atmospheric retrieval (see \S\,\ref{sec:retrieval}).
+We also tried fitting for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model predictions, and found that the fitted limb darkening coefficients are consistent with the model.
+We opted to fix the coefficients in our final analysis because it improves the precision on the transit depths by about 10\%.
+We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.
 
-\subsubsection{Transit Depths Redder than FIXME $\mu$m}
-The red edge of the transmission spectrum is of interest for our analysis because methane becomes the dominant absorber over water at wavelengths $> 1.6\,\mu$m.
-Unfortunately, we find that the spectroscopic light curves between FIXME $\mu$m are too poor in quality to accurately measure the transit depth.
-The residuals show evidence for correlated noise and the reduced $\chi_2$ values increase to FIXME.
+\subsubsection{Transit Depths Redder than 1.62 $\mu$m}
+The red edge of the transmission spectrum is of interest for our analysis because methane is expected to be the dominant absorber over water at wavelengths greater than $1.6\,\mu$m.
+Unfortunately, we find that the our reddest spectroscopic light curves ($>1.62\mu$m) are too poor in quality to robustly measure the transit depth.
+The residuals exhibit correlated noise and the fits have reduced $\chi^2_\nu>2$.
+The transit depths are also sensitive to which method we use to fit for instrument systematics.
+We therefore opt not to report transit depths redder than the 1.62 $\mu$m channel.
 
@@
 \begin{deluxetable}{c c c}
-\tablecaption{WASP-107b transmission spectrum \label{transit_depths}}
+\tablecaption{WASP-107b transmission spectrum \label{tab:transit_depths}}
 \tablehead{\colhead{Wavelength ($\mu$m)}  & \colhead{Transit depth} & \colhead{Uncertainty}}
@@
 \section{Atmospheric Retrieval}
+\label{sec:retrieval}
 We use the CHIMERA chemically-consistent transmission retrieval tool described in \cite{kreidberg15b} to determine the basic atmospheric properties.
@@
 \section{Future Plans}
+WASP-107b is slated to be observed with numerous facilities.
+Transits were recently observed with the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels.
+Spitzer eclipses are planned (Program FIXME; PI L.
+Kreidberg).
+In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations.
+
 Combine data from Spitzer/HST

### 2017-08-06 ms.tex
mcmc convergence

@@
 We then used the Markov chain Monte Carlo (MCMC) algorithm to estimate parameter uncertainties \citep{foremanmackey13}.
+The chain had 50 walkers which each ran for 10e4 steps with the first 10\% discarded as burn-in.
+We tested for convergence by dividing the chain in two halves and confirming that they gave consistent results.
 The transit time was $t_0 = 2457910.45407\pm6\mathrm{e}{-5}$ and the planet/star radius was $r_p/r_s = 0.14399\pm0.00017$.
-FIXME: converged?
 
@@
 We performed an MCMC fit to the light curves with \texttt{emcee}.
+For each light curve we ran a fit with 50 walkers and 1000 steps per walker, and tested for convergence as we did for the white light curve.
 The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:transit_depths}.
-FIXME: converged?
 

### 2017-08-11 ms.tex
minor edits

@@
 We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{https://github.com/spacetelescope/PandExo\_HST}.
-As is standard for observations of bright targets, we used the spatial scanning observing mode, which slews the telescope in the spatial direction over the course of an exposure.
+As is standard for WFC3 observations of bright targets, we used the spatial scanning observing mode, which slews the telescope in the spatial direction over the course of an exposure.
 The scan rate was 0.12 arcseconds/sec.
@@
 To correct the change in dispersion solution over the length of the spatial scan, we interpolated each row to the wavelength scale of the row corresponding to the spectral trace.
-We also corrected for slight drift in the spectral direction ($<0.1$ pixels) by interpolating each the first exposure.
+We then corrected for slight drift in the spectral direction ($<0.1$ pixels) by interpolating each spectrum to the wavelength scale of the first exposure.
 To subtract the background, we took the median of sky pixels that were uncontaminated by the target spectrum (rows $5-250$, columns $5-15$).
@@
 In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \citep{dai17}.
-These values are consistent with our best fit when we allowed them to vary freely (and masked the star spot).
+These values are consistent with our best fit when we allowed them to vary freely.
 We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star with effective temperature 4300 K, calculated with the \texttt{limb-darkening} package from \cite{espinoza15}.
@@
 We then used the Markov chain Monte Carlo (MCMC) algorithm to estimate parameter uncertainties \citep{foremanmackey13}.
-The chain had 50 walkers which each ran for 10e4 steps with the first 10\% discarded as burn-in.
+The chain had 50 walkers which each ran for $10^4$ steps with the first 10\% discarded as burn-in.
 We tested for convergence by dividing the chain in two halves and confirming that they gave consistent results.
-The transit time was $t_0 = 2457910.45407\pm6\mathrm{e}{-5}$ and the planet/star radius was $r_p/r_s = 0.14399\pm0.00017$.
+The transit time was $t_0 = 2457910.45407\pm6\mathrm{e}{-5}$ BJD$_\mathrm{TDB}$ and the planet/star radius was $r_p/r_s = 0.14399\pm0.00017$.
 
@@
 We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.65 $\mu$m, shown in Figure\,\ref{fig:lc}.
-We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics are nearly constant as a function of wavelength \citep{stevenson14c, kreidberg14a}.
+We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics have the same morphology at all wavelengths \citep{stevenson14c, kreidberg14a}.
 For this method, the transit model $T_\lambda(t)$ is multiplied by the systematics vector from the white light curve fit ($F_\mathrm{white}/T_\mathrm{white}$), and rescaled by a factor $C_\lambda + V_\lambda t$.
-An advantage of this approach is that it removes the star spot crossing feature, enabling us to use orbit three with no additional correction.
-There is a small chromatic effect due to the changing amplitude of the spot crossing feature with wavelength, but this effect is not detectable in our data.
+One advantage of this approach is that it removes the star spot crossing feature, enabling us to use orbit three with no additional correction.
+The amplitude of the feature has no detectable wavelength dependence at the level of precision of our data.
 As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model.
-The fit had just three free parameters: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
+The final spectroscopic light curve fits had just three free parameters: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
 
@@
 We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum.
-For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (differing by just $0.3\sigma$ on average), except with a constant offset of 350 ppm.
-We attribute the offset to the uncorrected star-spot crossing feature.
-Regardless of the cause, the offset does not affect our final analysis, because the planet's one-bar radius is a free parameter in the atmospheric retrieval (see \S\,\ref{sec:retrieval}).
-We also tried fitting for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model predictions, and found that the fitted limb darkening coefficients are consistent with the model.
-We opted to fix the coefficients in our final analysis because it improves the precision on the transit depths by about 10\%.
+For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (differing by just $0.3\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
+This offset does not affect our final analysis because the planet's one-bar radius is a free parameter in the atmospheric retrieval (see \S\,\ref{sec:retrieval}).
+We also fit for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values.
+We found that the fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths by about 10\%.
 We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.

### 2017-08-11 ms.tex
interior structure

@@
 
-\section{Analysis}
+\section{Light Curve Analysis}
 The data analysis had two parts: the band-integrated ``white" light curve fit and the spectroscopic light curve fits.
@@
 
-\section{Atmospheric Retrieval}
+\section{Composition of the Atmosphere}
+In this section we discuss constraints on the composition of WASP-107b's atmosphere based on interior structure modeling and atmospheric retrieval of the measured transmission spectrum.
+
+\subsection{Atmospheric Metallicity from Interior Structure Modeling}
+Given WASP-107b's unusually low density, we sought to estimate the range of atmospheric metallicities that are consistent with the observed mass and radius.
+We estimated the bulk metallicity of WASP-107b using the structure evolution modeling of \cite{thorngren16}.
+These models assume a thermally inert heavy-element core with a convective envelope of additively mixed H/He \citep{saumon95} and heavy-element impurities.
+The heavy elements were a 50-50 rock-ice mix from ANEOS \citep{thompson90}.
+We evolved the planets in time using the atmospheric models of \cite{fortney07}.
+The results are sensitive to assumptions about the stellar age, which is uncertain \citep[either $0.6\pm0.2$ to $8.3\pm4.3$ Gyr depending on model assumptions;][]{mocnik17}.
+We therefore ran two models, with uniform age priors of either $0.2-1.0$ or $1.0-13.8$ Gyr.
+We used the published mass and radius estimates \citep[$0.12\pm0.01\,M_\mathrm{J}$, $0.94\pm0.02$;][]{anderson17}.
+A more precise mass estimate is in preparation (B.
+Benneke, priv.\,comm.); however, we find that our results are not signficantly affected when we vary the assumed mass by $\pm3\,\sigma$ from the published value because the dominant uncertainties are the stellar age and core mass.
+
+Based on these assumptions, we fit for envelope metallicity and core mass using an MCMC with uniform priors on both parameters.
+The MCMC burned in for $10^3$ steps and then collected $4\times10^6$ samples, which we thinned to $10^5$.
+The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was about 18 (the value for water), using the approach of \cite{fortney13}.
+Figure\,\ref{fig:metal_prior} shows the results.
+We find that the inferred atmospheric metallicity and core mass are inversely proportional, as expected (the more massive the core, the lighter the envelope must be to maintain the observed radius).
+Larger core masses and envelope metallicities are allowed for the younger stellar age (when planets age, they cool and contract).
+For the younger (older) stellar age, we find an upper limit on the core mass of FIXME and an upper limit on atmospheric metallicity of FIXME.
+
+\begin{figure}
+\includegraphics[width = 0.5\textwidth]{Figures/metalByMass.png}
+\caption{FIXME}
+\label{fig:metal_prior}
+\end{figure}
+
+\subsection{Retrieval}
 \label{sec:retrieval}
-We use the CHIMERA chemically-consistent transmission retrieval tool described in \cite{kreidberg15b} to determine the basic atmospheric properties.
-Briefly, the transmission spectrum solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
-We parametrize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
-We have subsequently upgraded the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases generated from the pre-tabulated line-by-line cross section database ($dv/v~10^6$) described in \cite{freedman14}.
+We used the CHIMERA chemically-consistent retrieval tool described in \cite{kreidberg15b} to determine the planet's basic atmospheric properties from the transmission spectrum.
+Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
+We parameterize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
+We recently upgraded the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases generated from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
 The transmission forward model is coupled with the powerful PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.
@@
 The nominal model set-up requires a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity and carbon-to-oxygen ratio for molecular composition determination, disquilibrium properties crudely parameterized with nitrogen species and carbon species quench pressures \citep{morley17}, and aerosol properties.
-We experiment with three atmospheric scenarios within this setup (Table XX):  Scenarios with and without methane, and with/without patchy cloud cover (as described and implemented in \citealt{line16}.
-In all three scenarios we fix the T-P profile "shape" but scale the irradiation temperature, which incorporates unknown variables like the albedo and heat transport efficiency.
+We experiment with three atmospheric scenarios within this setup: scenarios with and without methane, and with/without patchy cloud cover (as described and implemented in \citealt{line16}.
+In all three scenarios we fix the T-P profile ``shape" but scale the irradiation temperature, to allow for the unknown albedo and heat transport efficiency.
 T-P profile shape information is not readily retrievable, nor does it strongly influence information derived from WFC3 transmission spectra \citep[e.g.][]{kreidberg15b}.
-We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
-Aerosols are simply parameterized with a ``gray cloud top pressure" with the option for cloud patchiness.
-We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
+Aerosols are parameterized with a ``gray cloud top pressure" with the option for cloud patchiness.
+%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
+%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
 
@@
 
-We note that some directly imaged self-luminous planets in this temperature regime are also unexpectedly depleted in methane (e.g., Skemer et al.
-2014), when compared to similar temperature brown dwarfs.
-The leading hypothesis of this depletion \citep{zahnle15} is due to the CH4-CO quench point occurring in the CO stability field due to the low gravity of self-luminous planets when compared to brown dwarfs.
+We note that some directly imaged self-luminous planets in this temperature regime are also unexpectedly depleted in methane when compared to similar temperature brown dwarfs \citep{skemer14}.
+The leading hypothesis of this depletion \citep{zahnle14} is due to the CH4-CO quench point occurring in the CO stability field due to the low gravity of self-luminous planets when compared to brown dwarfs.
 In order this mechanism to work for neptune massed planet, the internal temperature would have to be near ~500K to get the deep temperatures high enough to thermochemically deplete methane at the quench point.
@@
 With a sufficiently thick photochemical haze with particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.
-
-\subsection{Atmospheric Metallicity Predictions from Interior Structure Modeling}
- We determine the bulk metallicity of WASP-107 b using the structure evolution modeling of Thorngren et al.
-(2016).
-These consider a thermally inert heavy-element core with a convective envelope of additively mixed H/He (Saumon et al.
-1995) and heavy-element impurities.
-The heavy-elements were a 50-50 rock-ice mix from ANEOS (Thompson 1990).
-Planets were evolved in time using the atmospheric models of Fortney (2007).
-We explored possible envelope metallicities and core masses using a metropolis-hastings MCMC with M = .12 +/- .03, R = .94 +/- .02, and the age as alternatively .2-1.0 Gyr or 1.0-13.8 Gyr.
-Priors were $\propto 1$ in envelope metal mass fraction and core mass.
-The MCMC burned in for 1000 steps and then collected 4,000,000 samples, thinned down to 100,000.
-The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was about 18 (the value for water), using the approach of Fortney (2013).
 

### 2017-08-12 ms.tex
captions

@@
 \affiliation{Arizona State University}
+\nocollaboration
+
+\author{Daniel Thorngren}
+\affiliation{University of California Santa Cruz}
 \nocollaboration
@@
 \includegraphics[width = \textwidth]{Figures/lc.pdf}
-\caption{FIXME}
+\caption{Broadband and spectrophotometric transit light curves for WASP-107b measured with \textit{HST}/WFC3 compared to best fit models (left) and residuals from the best fit (right).
+Annotations indicate the central wavelength and root-mean-square (rms) residuals for each light curve.
+A possible star-spot crossing feature is shaded in gray in the upper right; our systematic error correction removes this feature from the spectroscopic light curves.}
 \label{fig:lc}
@@
 We attribute the excess noise to loss of flux off the edge of the detector, which can occur if there is variation in the position or length of the spatial scan.
-There is no evidence for correlated noise in the residuals, so to account for the excess noise we simply increased the per-point uncertainties by a factor of 1.7 to achieve a reduced $\chi^2$ value of unity.
+There is no evidence for correlated noise in the residuals, so to account for the excess noise we simply increased the per-point uncertainties by a factor of 1.7 to achieve a $\chi^2_\nu$ value of unity.
 We then used the Markov chain Monte Carlo (MCMC) algorithm to estimate parameter uncertainties \citep{foremanmackey13}.
@@
 
-The best fit light curves have a median reduced $\chi^2$ value of 1.16.
-To ensure that we did not underestimate the uncertainty on the transit depths, we rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2$ values are unity.
+The best fit light curves have a median $\chi^2_\nu$ value of 1.16.
+To ensure that we did not underestimate the uncertainty on the transit depths, we rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2_\nu$ values are unity.
 We performed an MCMC fit to the light curves with \texttt{emcee}.
@@
 Unfortunately, we find that the our reddest spectroscopic light curves ($>1.62\mu$m) are too poor in quality to robustly measure the transit depth.
-The residuals exhibit correlated noise and the fits have reduced $\chi^2_\nu>2$.
+The residuals exhibit correlated noise and the fits have $\chi^2_\nu>2$.
 The transit depths are also sensitive to which method we use to fit for instrument systematics.
@@
 \includegraphics[width = \textwidth]{Figures/spectrum.pdf}
-\caption{FIXME}
+\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals).
+The features at $1.15$ and $1.4$ $\mu$m are due to water absorption.
+The right-hand axis indicates the normalized atmospheric scale height assuming a $100\times$ solar composition.}
 \label{fig:spectrum}
@@
 A more precise mass estimate is in preparation (B.
-Benneke, priv.\,comm.); however, we find that our results are not signficantly affected when we vary the assumed mass by $\pm3\,\sigma$ from the published value because the dominant uncertainties are the stellar age and core mass.
+Benneke, priv.\,comm.); however, we find that our results are not significantly affected when we vary the assumed mass by $\pm3\,\sigma$ from the published value because the dominant uncertainties are the stellar age and core mass.
 
@@
 We find that the inferred atmospheric metallicity and core mass are inversely proportional, as expected (the more massive the core, the lighter the envelope must be to maintain the observed radius).
-Larger core masses and envelope metallicities are allowed for the younger stellar age (when planets age, they cool and contract).
+Larger core masses and envelope metallicities are allowed for the younger stellar age since planets cool and contract as they age.
 For the younger (older) stellar age, we find an upper limit on the core mass of FIXME and an upper limit on atmospheric metallicity of FIXME.
@@
 \includegraphics[width = 0.5\textwidth]{Figures/metalByMass.png}
-\caption{FIXME}
+\caption{Estimated envelope metallicity and core mass based on interior structure modeling of the mass and radius of WASP-107b. FIXME ages.}
 \label{fig:metal_prior}
@@
 \label{sec:retrieval}
-We used the CHIMERA chemically-consistent retrieval tool described in \cite{kreidberg15b} to determine the planet's basic atmospheric properties from the transmission spectrum.
+We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval tool described in \cite{kreidberg15b}.
 Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
 We parameterize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
-We recently upgraded the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases generated from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
+We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases generated from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
 The transmission forward model is coupled with the powerful PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.

### 2017-08-13 ms.tex
intro

@@
 \begin{abstract}
-We report a measurement of the near-infrared transmission spectrum of the warm Neptune exoplanet WASP-107b measured with the Hubble Space Telescope Wide Field Camera 3.
+%We report a measurement of the near-infrared transmission spectrum of the warm Neptune exoplanet WASP-107b measured with the Hubble Space Telescope Wide Field Camera 3.
 We detecta water absorption feature (FIXME confidence).
@@
 \section{Introduction} \label{sec:intro}
+The composition of a planet's atmosphere depends on where and how the planet formed.
+By measuring atmospheric metallicity and molecular abundance ratios, we can shed light on important aspects of the formation process such as location within the disk and the relative accretion rates of gas versus solids \citep[][]{oberg11, fortney13, madhusudhan14, alidib16, mordasini16, espinoza17}.
+
+The warm Neptune WASP-107b is an intriguing target for atmosphere characterization for several reasons.
+It is an intermediate size between ice giants and gas giants, with a mass similar to Neptune's and a radius close to Jupiter's \citep[$0.12\,M_\mathrm{Jup}$, $0.94\,R_\mathrm{Jup}$;][]{anderson17}.
+Studying the atmospheres of planets in this transition region will provide additional clues in the much-debated mystery of what stunts the growth of Neptune-size planets \citep[e.g][]{pollack96, dawson16, frelikh17}.
+
+WASP-107b also has a relatively low equilibrium temperature (780 K) compared to most other exoplanets that are amenable to atmosphere characterization.
+This results in a distinct atmospheric chemistry from other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.
+Spectral features from both water and methane are accessible with current observing facilities, raising the possibility of a spectroscopic estimate of the carbon-to-oxygen ratio (C/O).
+Previous measurements of C/O have been challenging because they rely on photometry that covers absorption features from multiple molecules \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.
+FIXME: maybe cut this.
+
+In addition, WASP-107b is one of the best targets discovered to date for atmosphere characterization.
+Thanks to its large atmospheric scale height and small, bright host star, the expected signal-to-noise for the transmission spectrum is comparable to the best studied benchmarks in the field (e.g.
+HD\,209458b, HD\,189733b).
+The planet is also a feasible target for thermal emission measurements, with an expected secondary eclipse depth of $\sim0.1$\% in \emph{Spitzer}'s $4.5\,\mu$m channel.
+In this paper we report the first atmosphere characterization of WASP-107b: a near-infrared transmission spectrum measured with the \emph{Hubble Space Telescope} (\emph{HST}).
 
@@
 \subsection{Atmospheric Metallicity from Interior Structure Modeling}
-Given WASP-107b's unusually low density, we sought to estimate the range of atmospheric metallicities that are consistent with the observed mass and radius.
+Given WASP-107b's unusually low density, we quantiatively explored the range of atmospheric metallicities that are consistent with the observed mass and radius.
 We estimated the bulk metallicity of WASP-107b using the structure evolution modeling of \cite{thorngren16}.

### 2017-08-13 temp.py

@@ -1,0 +1,19 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+#constants (all SI units)
+h = 6.626e-34		#Planck constant
+k = 1.38e-23		#Boltzmann constant
+c = 3.0e8		#speed of light
+
+
+#return blackbody spectral radiance as a function of wavelength lambda (in m) and temperature T (Kelvin)
+def bb(l, T):
+	return 2*h*c**2/(l**5*(np.exp(h*c/(l*k*T)-1)))
+
+l = 4.5e-6
+depth = 0.021
+Tp, Ts = 780, 4430
+print bb(l, Tp)/bb(l, Ts)*depth*1e6
+

### 2017-08-13 temp.py
intro

@@ -1,19 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-
-#constants (all SI units)
-h = 6.626e-34		#Planck constant
-k = 1.38e-23		#Boltzmann constant
-c = 3.0e8		#speed of light
-
-
-#return blackbody spectral radiance as a function of wavelength lambda (in m) and temperature T (Kelvin)
-def bb(l, T):
-	return 2*h*c**2/(l**5*(np.exp(h*c/(l*k*T)-1)))
-
-l = 4.5e-6
-depth = 0.021
-Tp, Ts = 780, 4430
-print bb(l, Tp)/bb(l, Ts)*depth*1e6
-

### 2017-08-13 ms.tex
ctoo

@@
 We parameterize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
-We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases generated from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
+We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
 The transmission forward model is coupled with the powerful PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.
 
-The nominal model set-up requires a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity and carbon-to-oxygen ratio for molecular composition determination, disquilibrium properties crudely parameterized with nitrogen species and carbon species quench pressures \citep{morley17}, and aerosol properties.
-We experiment with three atmospheric scenarios within this setup: scenarios with and without methane, and with/without patchy cloud cover (as described and implemented in \citealt{line16}.
-In all three scenarios we fix the T-P profile ``shape" but scale the irradiation temperature, to allow for the unknown albedo and heat transport efficiency.
-T-P profile shape information is not readily retrievable, nor does it strongly influence information derived from WFC3 transmission spectra \citep[e.g.][]{kreidberg15b}.
-Aerosols are parameterized with a ``gray cloud top pressure" with the option for cloud patchiness.
+The nominal model set-up requires a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity and carbon-to-oxygen ratio for molecular composition determination, disequilibrium properties parameterized with nitrogen and carbon species quench pressures \citep{morley17}, and aerosol properties.
+We fix the T-P profile morphology but scale the irradiation temperature to allow for the unknown albedo and heat transport efficiency.
+Aerosols are parameterized with a gray cloud top pressure with the option for cloud patchiness (as described by \citealt{line16}).
+
+Our nominal model included a gray cloud, prior
+
+We experiment with several scenarios within this framework, including models with/without clouds, models with/without quenching, including methane opacity, and a uniform prior in atmospheric metallicity from (FIXME to $30\times$ solar based on the interior structure modeling).
+We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary freely.
+
+We find strong evidence for water absorption (FIXME sigma) and clouds (FIXME sigma).
+
 %We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.

### 2017-08-13 Analysis/get_COratio.py

@@ -7,7 +7,7 @@ import scipy.stats as st
 
 def get_significance(alpha):
 	z = st.norm.ppf(1.-alpha)
-	print "should I be dividing alpha by 2?"
+#	print "should I be dividing alpha by 2?"		YESS! check with significance test of alpha = 0.003
 	return z
 
 """alpha = np.linspace(0.001, 0.999, 100)
@@ -23,7 +23,7 @@ hist(ctoo)
 plt.axvline(np.log10(0.54), color = 'y')
 plt.title("CC")
 alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
-print "CC sig", get_significance(alpha)
+print "CC sig", get_significance(alpha/2.)
 
 plt.subplot(222)
 p = pickle.load(open("Global_Cloud_noCH4_samples.pic", "rb"))
@@ -32,7 +32,7 @@ hist(ctoo)
 plt.axvline(np.log10(0.54), color = 'y')
 alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
 plt.title("CC no CH4")
-print "CC no CH4 sig", get_significance(alpha)
+print "CC no CH4 sig", get_significance(alpha/2.)
 
 plt.subplot(223)
 p = pickle.load(open("MCMC_CC_QUENCH.pic", "rb"))
@@ -45,7 +45,7 @@ hist(ctoo)
 plt.axvline(np.log10(0.54), color = 'y')
 alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
 plt.title("CC quench")
-print "CC quench", get_significance(alpha)
+print "CC quench", get_significance(alpha/2.)
 
 plt.subplot(224)
 p = pickle.load(open("MCMC_CC_Limit_Metallicity.pic", "rb"))
@@ -55,7 +55,7 @@ hist(ctoo)
 alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
 plt.axvline(np.log10(0.54), color = 'y')
 plt.title("CC Z")
-print "CC Z", get_significance(alpha), alpha
+print "CC Z", get_significance(alpha/2.), alpha
 
 print "3 sigma upper limit on metallicity from CC Z case:"
 ind = np.argsort(ctoo)
[…]

### 2017-08-13 ms.tex
pairs

@@
 \subsection{Atmospheric Metallicity from Interior Structure Modeling}
-Given WASP-107b's unusually low density, we quantiatively explored the range of atmospheric metallicities that are consistent with the observed mass and radius.
-We estimated the bulk metallicity of WASP-107b using the structure evolution modeling of \cite{thorngren16}.
+Given WASP-107b's unusually low density, we quantitatively explored the range of atmospheric metallicities that are consistent with the observed mass and radius using the structure evolution modeling of \cite{thorngren16}.
 These models assume a thermally inert heavy-element core with a convective envelope of additively mixed H/He \citep{saumon95} and heavy-element impurities.
@@
 The MCMC burned in for $10^3$ steps and then collected $4\times10^6$ samples, which we thinned to $10^5$.
-The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was about 18 (the value for water), using the approach of \cite{fortney13}.
+The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was 18 (the value for water), using the approach of \cite{fortney13}.
 Figure\,\ref{fig:metal_prior} shows the results.
@@
 We parameterize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
-We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} for the aforementioned gases from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
-The transmission forward model is coupled with the powerful PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.
+We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
+The transmission forward model is coupled with the PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.
 
-The nominal model set-up requires a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity and carbon-to-oxygen ratio for molecular composition determination, disequilibrium properties parameterized with nitrogen and carbon species quench pressures \citep{morley17}, and aerosol properties.
+Our nominal model includes a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity, the carbon-to-oxygen ratio, and a gray cloud-top pressure.
 We fix the T-P profile morphology but scale the irradiation temperature to allow for the unknown albedo and heat transport efficiency.
-Aerosols are parameterized with a gray cloud top pressure with the option for cloud patchiness (as described by \citealt{line16}).
+We put a uniform prior on the atmospheric metallicity of
 
-Our nominal model included a gray cloud, prior
+and the quench pressures for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
+
+for cloud patchiness (as described by \citealt{line16}).
 

### 2017-08-13 Figures/mycorner.py

@@ -1,0 +1,654 @@
+# -*- coding: utf-8 -*-
+
+from __future__ import print_function, absolute_import
+
+import logging
+import numpy as np
+import matplotlib.pyplot as pl
+from matplotlib.ticker import MaxNLocator, NullLocator
+from matplotlib.colors import LinearSegmentedColormap, colorConverter
+from matplotlib.ticker import ScalarFormatter
+
+try:
+    from scipy.ndimage import gaussian_filter
+except ImportError:
+    gaussian_filter = None
+
+__all__ = ["corner", "hist2d", "quantile"]
+
+
+def corner(xs, bins=20, range=None, weights=None, color="k",
+           smooth=None, smooth1d=None,
+           labels=None, label_kwargs=None,
+           show_titles=False, title_fmt=".2f", title_kwargs=None,
+           truths=None, truth_color="#4682b4",
+           scale_hist=False, quantiles=None, verbose=False, fig=None,
+           max_n_ticks=5, top_ticks=False, use_math_text=False, reverse=False,
+           hist_kwargs=None, **hist2d_kwargs):
+    """
+    Make a *sick* corner plot showing the projections of a data set in a
+    multi-dimensional space. kwargs are passed to hist2d() or used for
+    `matplotlib` styling.
+
+    Parameters
+    ----------
+    xs : array_like[nsamples, ndim]
+        The samples. This should be a 1- or 2-dimensional array. For a 1-D
+        array this results in a simple histogram. For a 2-D array, the zeroth
+        axis is the list of samples and the next axis are the dimensions of
+        the space.
+
+    bins : int or array_like[ndim,]
+        The number of bins to use in histograms, either as a fixed value for
+        all dimensions, or as a list of integers for each dimension.
+
+    weights : array_like[nsamples,]
+        The weight of each sample. If `None` (default), samples are given
+        equal weight.
+
+    color : str
+        A ``matplotlib`` style color for all histograms.
+
+    smooth, smooth1d : float
+       The standard deviation for Gaussian kernel passed to
+       `scipy.ndimage.gaussian_filter` to smooth the 2-D and 1-D histograms
+       respectively. If `None` (default), no smoothing is applied.
+
+    labels : iterable (ndim,)
+        A list of names for the dimensions. If a ``xs`` is a
+        ``pandas.DataFrame``, labels will default to column names.
+
+    label_kwargs : dict
+        Any extra keyword arguments to send to the `set_xlabel` and
+        `set_ylabel` methods.
+
+    show_titles : bool
[…]

### 2017-08-13 Figures/pairs.py

@@ -1,0 +1,25 @@
+import mycorner as corner
+import matplotlib.pyplot as plt
+import numpy as np
+import pickle
+from pylab import hist
+
+
+p = pickle.load(open("MCMC_CC_Limit_Metallicity.pic", "rb"))
+p = p[:,:-1]
+
+d = p[:,0:4]
+ind = (d[:,1]<0.8)&(d[:,3]<-0.2)#&(d[:,0]<1000)&(d[:,2]<-0.25)
+d = d[ind]
+
+ctoo = np.log10(0.54)
+print ctoo
+print 1.0*sum(d[:,2]>ctoo)/(1.*len(d[:,2]))
+
+labels = ["T_irr", "Z", "C/O", "P_cloud"]
+
+plt.figure(figsize = (6,6))
+#fig = corner.corner(d, labels = labels, plot_contours = False, no_fill_contours = True, plot_datapoints = False, range = [[500,1000], 0.997, 0.997, 0.997])
+fig = corner.corner(d, labels = labels, smooth = 0.7)#, levels = [0.16, 0.5, 0.84]) 
+
+plt.savefig("pairs.png")

### 2017-08-13 ms.tex
retrieval fig

@@
 We used the published mass and radius estimates \citep[$0.12\pm0.01\,M_\mathrm{J}$, $0.94\pm0.02$;][]{anderson17}.
-A more precise mass estimate is in preparation (B.
+%A more precise mass estimate is in preparation (B.
 Benneke, priv.\,comm.); however, we find that our results are not significantly affected when we vary the assumed mass by $\pm3\,\sigma$ from the published value because the dominant uncertainties are the stellar age and core mass.
@@
 %We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
+
+\begin{figure}
+\includegraphics[width = 0.5\textwidth]{Figures/retrieval.pdf}
+\caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure.
+Dashed vertical lines show the median and $\pm1\,\sigma$ confidence interval.
+Solar metallicity and C/O values are indicated with solid vertical lines.}
+\label{fig:retrieval}
+\end{figure}
 

### 2017-08-13 Figures/histograms.py

@@ -1,0 +1,99 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pickle
+from pylab import hist
+import matplotlib as mpl
+
+def quantiles(d, q):
+        ind = np.argsort(d)
+        d = d[ind]
+        return_quantile = np.zeros_like(q)
+        n = len(d)
+        for i in range(0, len(q)):
+                return_quantile[i] = d[int(q[i]*float(n))]
+        return return_quantile
+
+mpl.rcParams['font.sans-serif'] = 'Arial'
+mpl.rcParams['font.size'] = 18
+mpl.rcParams['lines.linewidth'] = 1.4
+
+p = pickle.load(open("MCMC_CC_Limit_Metallicity.pic", "rb"))
+p = p[:,:-1]
+
+d = p[:,0:4]
+#ind = (d[:,1]<0.8)&(d[:,3]<-0.2)#&(d[:,0]<1000)&(d[:,2]<-0.25)
+#d = d[ind]
+nbins = 30
+q = [0.16, 0.5, 0.84]
+
+plt.subplot(221)
+hist(d[:,1], histtype ='stepfilled', color = 'k', facecolor='#b7c9e2', bins = nbins, alpha = 0.8, normed = True)
+
+plt.gca().set_yticklabels([])
+plt.xlabel("[Fe/H]")
+plt.ylabel("Frequency")
+plt.xlim(-2, np.max(d[:,1]))
+
+plt.gca().text(0.08, 0.95, "a", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
+
+lo, med, hi = quantiles(d[:,1], q)
+plt.axvline(lo, color = '0.1', linestyle = 'dashed')
+plt.axvline(med, color = '0.1', linestyle = 'dashed')
+plt.axvline(hi, color = '0.1', linestyle = 'dashed')
+plt.axvline(0., color = '#c44240', linestyle = 'solid')
+plt.gca().text(0.52, 0.7, "solar", transform=plt.gca().transAxes, fontsize=18, va='top', rotation = 90)
+
+##############################
+
+plt.subplot(222)
+hist(d[:,2], histtype ='stepfilled', color = 'k', facecolor='#b7c9e2', bins = nbins, alpha = 0.8, normed = True)
+
+plt.gca().get_yaxis().set_visible(False)		
+plt.xlabel("log(C/O)")
[…]

### 2017-08-13 Figures/mycorner.py

@@ -258,6 +258,9 @@ def corner(xs, bins=20, range=None, weights=None, color="k",
             x0 = np.array(list(zip(b[:-1], b[1:]))).flatten()
             y0 = np.array(list(zip(n, n))).flatten()
             ax.plot(x0, y0, **hist_kwargs)
+	    ax.get_xaxis().set_visible(False)		#LK
+	    ax.get_yaxis().set_visible(False)		#LK
+	    #ax.set_frame_on(False)
 
         if truths is not None and truths[i] is not None:
             ax.axvline(truths[i], color=truth_color)
@@ -335,6 +338,7 @@ def corner(xs, bins=20, range=None, weights=None, color="k",
                 ScalarFormatter(useMathText=use_math_text))
 
         for j, y in enumerate(xs):
+	    if i!=j: ax.set_frame_on(False)	#LK
             if np.shape(xs)[0] == 1:
                 ax = axes
             else:
@@ -623,6 +627,7 @@ def hist2d(x, y, bins=20, range=None, weights=None, levels=None, smooth=None,
         ax.plot(x, y, "o", zorder=-1, rasterized=True, **data_kwargs)
 	ax.get_xaxis().set_visible(False)		#LK
 	ax.get_yaxis().set_visible(False)		#LK
+        ax.set_frame_on(False)
 
     # Plot the base fill to hide the densest data points.
     if (plot_contours or plot_density) and not no_fill_contours:

### 2017-08-13 ms.tex
abstract

@@
 \begin{abstract}
-%We report a measurement of the near-infrared transmission spectrum of the warm Neptune exoplanet WASP-107b measured with the Hubble Space Telescope Wide Field Camera 3.
-We detecta water absorption feature (FIXME confidence).
-The feature is small amplitude, clouds.
-We combine an interior structure model of the planet with an atmospheric retrieval .
-We find an upper limit on the atmospheric metallicity of FIXME solar, a cloud deck .
+We report a Hubble Space Telescope measurement of the near-infrared transmission spectrum of the warm Neptune exoplanet WASP-107b.
+We analyzed the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the planet's mass and radius.
+We detect water absorption at $6.5\,\sigma$ confidence and set an upper limit on the atmospheric metallicity of $30\times$ solar.
+The methane abundance is depleted relative to expectations for a solar abundance pattern, suggesting a low C/O ratio or high internal heat flux.
+The amplitude of features in the transmission spectrum is smaller than expected for a cloudless atmosphere, crossing just one scale height, and requires an opaque condensate layer at high altitudes (0.1 - 3 mbar).
+We calculated physically motivated cloud and haze models and find that it is challenging for these models to reproduce the high enough in the atmosphere.
+
+Additional observations are planned for this system.
 \end{abstract}
@@
 \subsection{Atmospheric Metallicity from Interior Structure Modeling}
+\label{sec:interior}
 Given WASP-107b's unusually low density, we quantitatively explored the range of atmospheric metallicities that are consistent with the observed mass and radius using the structure evolution modeling of \cite{thorngren16}.
@@
 We fix the T-P profile morphology but scale the irradiation temperature to allow for the unknown albedo and heat transport efficiency.
-We put a uniform prior on the atmospheric metallicity of
+We put a uniform prior on the atmospheric metallicity of $0.01 - 30\times$ solar based on the upper limit from $\S$\,\ref{sec:interior}.
+
+The retrieval results are shown in Figure\,\ref{fig:retrieval}.
+The metallicity is fairly degenerate with the cloud-top pressure \citep{griffith14}, leading to a large allowed range for both values.
+Low metallicities are
+
+Low C/O values are favored, with a $1\,\sigma$ confidence interval of $0.01 - 0.1$. Solar C/O (0.54) is disfavored at FIXME confidence.
+
+cloud at low pressures 
+solar C/O is almost ruled out 
+metal -0.622107202821 0.569340080292 1.19470825739
+C/O 0.013703698053 0.0291553314874 0.101164675722
+T 525.868471118 650.602023714 818.754084294
+P -4.00123780684 -3.33972349689 -2.44574267349
 
 and the quench pressures for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}. 
-
 for cloud patchiness (as described by \citealt{line16}).
@@
 
-%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
-%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
+\section{Discussion}
+\subsubsection{Methane Abundance}
 
-\begin{figure}
-\includegraphics[width = 0.5\textwidth]{Figures/retrieval.pdf}
-\caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure.
-Dashed vertical lines show the median and $\pm1\,\sigma$ confidence interval.
-Solar metallicity and C/O values are indicated with solid vertical lines.}
-\label{fig:retrieval}
-\end{figure}
-
-\section{Discussion} \label{sec:discuss}
-
-We note that some directly imaged self-luminous planets in this temperature regime are also unexpectedly depleted in methane when compared to similar temperature brown dwarfs \citep{skemer14}.
-The leading hypothesis of this depletion \citep{zahnle14} is due to the CH4-CO quench point occurring in the CO stability field due to the low gravity of self-luminous planets when compared to brown dwarfs.
-In order this mechanism to work for neptune massed planet, the internal temperature would have to be near ~500K to get the deep temperatures high enough to thermochemically deplete methane at the quench point.
-Given the age of the star, it is unlikely that such high internal temperatures are still due to heat of formation.
-It is possible that tidal heating could increase the internal temperature (Agundez et al.
-2014; Morley et al.
-2017) but that would require a negligible eccentricity, of which is currently uncertain for WASP107b.
-
-\subsection{Methane}
-
-\subsection{Comparative planetology with HAT-P-11b, HAT-P-26b}
-
-\subsection{Constraints on Condensate Properties}
-The amplitude of WASP-107b's spectral features is about half that expected for an atmosphere free of condensates.
-In our retrieval analysis, we explored two simple parameterizations to model condensates: a gray opacity source and a sloped haze.
-The retrieved cloud-top pressure is FIXME at $1\,\sigma$ confidence.
-
+\subsubsection{Condensate Properties}
 We also considered physically motivated, self-consistent cloud and haze models based on .
@@
 
-\section{Future Plans}
+%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
+%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
+
+\begin{figure}
+\includegraphics[width = 0.5\textwidth]{Figures/retrieval.pdf}
+\caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure in bars.
+Dashed vertical lines show the median and $\pm1\,\sigma$ confidence interval.
+Solar metallicity and C/O values are indicated with solid vertical lines.}  \label{fig:retrieval}
+\end{figure}
+
+\section{Discussion} \label{sec:discuss}
+Our data suggest a sub-solar methane abundance at FIXME $\sigma$ confidence.
+This result was surprising because methane is expected to be the dominant carbon-bearing molecule at WASP-107b's temperature \citep{moses13}, but super-solar metal
+
+Some directly imaged planets in this temperature regime are also unexpectedly depleted in methane when compared to similar temperature brown dwarfs \citep{skemer14}.
+
+The leading hypothesis of this depletion \citep{zahnle14} is due to the CH4-CO quench point occurring in the CO stability field due to the low gravity of self-luminous planets when compared to brown dwarfs.
+
+In order this mechanism to work for neptune massed planet, the internal temperature would have to be near ~500K to get the deep temperatures high enough to thermochemically deplete methane at the quench point.
+Given the age of the star, it is unlikely that such high internal temperatures are still due to heat of formation.
+It is possible that tidal heating could increase the internal temperature (Agundez et al.
+2014; Morley et al.
+2017) but that would require a negligible eccentricity, of which is currently uncertain for WASP107b.
+
+\subsection{Comparative planetology with HAT-P-11b, HAT-P-26b}
+
+\section{Conclusions and Future Plans}
 WASP-107b is slated to be observed with numerous facilities.
@@
 \acknowledgments
-Fei Dai
-Jessica Spake
-
-L.R.K. acknowledges support from the Harvard Society of Fellows and the Harvard Astronomy Department Institute for Theory and Computation.
+We thank Fei Dai, Jessica Spake, and Ian Crossfield for productive conversations.
+L.R.K.
+acknowledges support from the Harvard Society of Fellows and the Harvard Astronomy Department Institute for Theory and Computation.
 

### 2017-08-13 Analysis/get_COratio.py

@@ -78,7 +78,8 @@ plt.subplot(132)
 plt.title("CH4")
 x = p[:,4]-6
 hist(x)
-CH4 = np.log10(4.4e-4)
+CH4 = np.log10(4.4e-3)
+CH4 = np.log10(0.00132)
 alpha = 1.0*sum(x>CH4)/len(x)*1.0
 print "CH4 sig", get_significance(alpha/2)
 print "test", get_significance(0.003/2)

### 2017-08-13 Figures/retrieval.py

@@ -37,6 +37,7 @@ plt.xlim(-2, np.max(d[:,1]))
 plt.gca().text(0.08, 0.95, "a", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
 
 lo, med, hi = quantiles(d[:,1], q)
+print "metal", lo, med, hi
 plt.axvline(lo, color = '0.1', linestyle = 'dashed')
 plt.axvline(med, color = '0.1', linestyle = 'dashed')
 plt.axvline(hi, color = '0.1', linestyle = 'dashed')
@@ -53,6 +54,7 @@ plt.xlabel("log(C/O)")
 plt.gca().text(0.08, 0.95, "b", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
 
 lo, med, hi = quantiles(d[:,2], q)
+print "C/O", 10**lo, 10**med, 10**hi
 plt.axvline(lo, color = '0.1', linestyle = 'dashed')
 plt.axvline(med, color = '0.1', linestyle = 'dashed')
 plt.axvline(hi, color = '0.1', linestyle = 'dashed')
@@ -74,6 +76,7 @@ plt.gca().set_yticklabels([])
 plt.gca().text(0.08, 0.95, "c", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
 
 lo, med, hi = quantiles(d[:,0], q)
+print "T", lo, med, hi
 plt.axvline(lo, color = '0.1', linestyle = 'dashed')
 plt.axvline(med, color = '0.1', linestyle = 'dashed')
 plt.axvline(hi, color = '0.1', linestyle = 'dashed')
@@ -90,6 +93,7 @@ plt.gca().get_yaxis().set_visible(False)		#LK
 plt.gca().text(0.08, 0.95, "c", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
 
 lo, med, hi = quantiles(d[:,3], q)
+print "P", lo, med, hi
 plt.axvline(lo, color = '0.1', linestyle = 'dashed')
 plt.axvline(med, color = '0.1', linestyle = 'dashed')
 plt.axvline(hi, color = '0.1', linestyle = 'dashed')

### 2017-08-14 ms.tex
retrieval

@@
 \begin{abstract}
-We report a Hubble Space Telescope measurement of the near-infrared transmission spectrum of the warm Neptune exoplanet WASP-107b.
+The transiting planet WASP-107b is an intriguing target for atmosphere characterization, with a large atmospheric scale height and a small, bright host star.
+
+w107 is a good target for atmospheric characterization, raising the exciting possibility of characterizing its formation history
+
+.
+With a mass similar to Neptune's and a radius close to Jupiter's, it occupies the transition region between ice giants and gas giants.
+The planet is also one of the strongest targets for atmospheric characterization yet discovered, thanks to its small, bright host star and large scale height.
+We report the first atmosphere characterization of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.
 We analyzed the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the planet's mass and radius.
 We detect water absorption at $6.5\,\sigma$ confidence and set an upper limit on the atmospheric metallicity of $30\times$ solar.
-The methane abundance is depleted relative to expectations for a solar abundance pattern, suggesting a low C/O ratio or high internal heat flux.
+We find that the methane abundance is depleted relative to expectations for a solar abundance pattern (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.
 The amplitude of features in the transmission spectrum is smaller than expected for a cloudless atmosphere, crossing just one scale height, and requires an opaque condensate layer at high altitudes (0.1 - 3 mbar).
-We calculated physically motivated cloud and haze models and find that it is challenging for these models to reproduce the high enough in the atmosphere.
-
-Additional observations are planned for this system.
+We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at this height.
+Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
+We can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities.
 \end{abstract}
@@
 Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
-We parameterize atmospheric composition with metallicity and carbon-to-oxygen ratio (C/O) under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
+We parameterize atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
 We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
@@
 
-Our nominal model includes a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity, the carbon-to-oxygen ratio, and a gray cloud-top pressure.
+Our nominal model includes a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity, the C/O, and a gray cloud-top pressure.
 We fix the T-P profile morphology but scale the irradiation temperature to allow for the unknown albedo and heat transport efficiency.
@@
 The retrieval results are shown in Figure\,\ref{fig:retrieval}.
-The metallicity is fairly degenerate with the cloud-top pressure \citep{griffith14}, leading to a large allowed range for both values.
-Low metallicities are
+The metallicity distribution spans the full range allowed by our priors, with some preference for larger values.
+The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence.
+These are not particularly precise constraints because the metallicity and cloud height are degenerate \citep{griffith14}.
+We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.
+The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).
 
-Low C/O values are favored, with a $1\,\sigma$ confidence interval of $0.01 - 0.1$. Solar C/O (0.54) is disfavored at FIXME confidence.
+We explored several alternate retrieval scenarios, including the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressures for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
+We did not find significant evidence for either additional complexity in the model, and the main retrieval results were unchanged.
 
-cloud at low pressures 
-solar C/O is almost ruled out 
-metal -0.622107202821 0.569340080292 1.19470825739
-C/O 0.013703698053 0.0291553314874 0.101164675722
-T 525.868471118 650.602023714 818.754084294
-P -4.00123780684 -3.33972349689 -2.44574267349
+\subsubsection{Methane Abundance}
+We ran two additional retrievals to explore whether methane depletion is the underlying cause of the low C/O measurement.
+In one test, we assumed chemical equilibrium but excluded methane opacity.
+In this test, the C/O value
 
-and the quench pressures for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}. 
-for cloud patchiness (as described by \citealt{line16}).
-
-We experiment with several scenarios within this framework, including models with/without clouds, models with/without quenching, including methane opacity, and a uniform prior in atmospheric metallicity from (FIXME to $30\times$ solar based on the interior structure modeling).
-We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary freely.
-
-We find strong evidence for water absorption (FIXME sigma) and clouds (FIXME sigma).
+We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium. These results $3\,\sigma$
 
 \section{Discussion}
-\subsubsection{Methane Abundance}
 
@@
 2017) but that would require a negligible eccentricity, of which is currently uncertain for WASP107b.
+
+low C/O ratio - planetesimal enrichment (cite Nestor)
+metallicity 30x solar - how does this compare to mass/atmospheric metallicity 
+clouds
 

### 2017-08-14 ms.tex
just conclusion left

@@
 
-We explored several alternate retrieval scenarios, including the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressures for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
+We explored several alternate retrieval scenarios, including the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
 We did not find significant evidence for either additional complexity in the model, and the main retrieval results were unchanged.
-
-\subsubsection{Methane Abundance}
-We ran two additional retrievals to explore whether methane depletion is the underlying cause of the low C/O measurement.
-In one test, we assumed chemical equilibrium but excluded methane opacity.
-In this test, the C/O value
-
-We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium. These results $3\,\sigma$
-
-\section{Discussion}
-
-\subsubsection{Condensate Properties}
-We also considered physically motivated, self-consistent cloud and haze models based on .
-To model WASP-107b's spectrum with self-consistent aerosols, we use the methods described in \citep{fortney08, morley15}.
-We include models from solar to $50\times$ solar metallicity and solar C/O ratio.
-We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS, see \citealt{morley12}), varying the cloud sedimentation efficiency from 0.1 to 3.
-None of the cloudy models are sufficiently low amplitude to match the observed muted signal.
-We also model an ad hoc photochemical 'soot' layer near the top of the atmosphere, scaling results from previous photochemical models for GJ 436b \citep{line11, morley17}.
-With a sufficiently thick photochemical haze with particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.
-
-%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
-%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
 
@@
 
-\section{Discussion} \label{sec:discuss}
-Our data suggest a sub-solar methane abundance at FIXME $\sigma$ confidence.
-This result was surprising because methane is expected to be the dominant carbon-bearing molecule at WASP-107b's temperature \citep{moses13}, but super-solar metal
+\subsubsection{Methane Abundance}
+We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O.
+In one test, we assumed chemical equilibrium but excluded methane opacity.
+This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$), resolving the tension with the solar value.
+
+We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium.
+The $3\,\sigma$ upper limit on the methane abundance is $1.2\times10^{-3}$.
+We compared this to the expected methane abundance for the temperature and pressure sensed at the terminator.
+Assuming  solar C/O and $3\,\times$ solar metallicity (the median retrieved value for the chemical equilbrium), the methane abundance is predicted to be $1.3\times10^{-3}$, inconsistent with our estimated value.
+Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.
+
+\subsection{Condensate Properties}
+In addition to the atmospheric retrieval, we also performed forward modeling of physically motivated, self-consistent clouds and hazes using the methods described in \cite{fortney08, morley15}.
+We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS, see \citealt{morley12}), for a range of metallicites from $1-50\times$ solar.
+We vary the cloud sedimentation efficiency from 0.1 to 3 (highly lofted, small particulate clouds).
+None of the models produce sufficiently low amplitude spectral features to match the observed spectrum.
+
+We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere, scaling results from previous photochemical models for GJ 436b \citep{line11, morley17}.
+With the most efficient haze production and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.
+However, these photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis, and we found in the previous section that methane is depleted in the atmosphere of WASP-107b.
+%Sulfur-based hazes are another possibility \citep{gao17}.
+Further modeling of the condensates is needed for a satisfactory explanation of the data.
+
+%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
+%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
+
+\section{Discussion and Conclusions} \label{sec:discuss}
+We combined our high precision transmission spectrum of WASP-107b with interior structure modeling to assess the planet's atmospheric composition.
+
+Upper limit on metallicity.
+The planet's radius is so large that we put an upper limit on the metallicity (additional metals would compress it).
+metallicity 30x solar - how does this compare to mass/atmospheric metallicity
+
+Methane depletion
+low C/O ratio - planetesimal enrichment (cite Nestor)
+internal heat, directly imaged planets
+
+Thick, high-altitude condensate layer
 
@@
 
-low C/O ratio - planetesimal enrichment (cite Nestor)
-metallicity 30x solar - how does this compare to mass/atmospheric metallicity 
-clouds
-
-\subsection{Comparative planetology with HAT-P-11b, HAT-P-26b}
-
-\section{Conclusions and Future Plans}
 WASP-107b is slated to be observed with numerous facilities.

### 2017-08-14 Analysis/get_COratio.py

@@ -4,15 +4,19 @@ import pickle
 from pylab import hist
 import scipy.stats as st
 
+def quantiles(d, q):
+        ind = np.argsort(d)
+        d = d[ind]
+        return_quantile = np.zeros_like(q)
+        n = len(d)
+        for i in range(0, len(q)):
+                return_quantile[i] = d[int(q[i]*float(n))]
+        return return_quantile
 
 def get_significance(alpha):
 	z = st.norm.ppf(1.-alpha)
 #	print "should I be dividing alpha by 2?"		YESS! check with significance test of alpha = 0.003
 	return z
-
-"""alpha = np.linspace(0.001, 0.999, 100)
-plt.plot(alpha, st.norm.ppf(alpha))
-plt.show()"""
 
 solarctoo = np.log10(0.54)
 
@@ -33,6 +37,9 @@ plt.axvline(np.log10(0.54), color = 'y')
 alpha = 1.0*sum(ctoo>solarctoo)/len(ctoo)*1.0
 plt.title("CC no CH4")
 print "CC no CH4 sig", get_significance(alpha/2.)
+
+q = [0.16, 0.5, 0.84]
+print 10.**quantiles(ctoo, q)
 
 plt.subplot(223)
 p = pickle.load(open("MCMC_CC_QUENCH.pic", "rb"))
@@ -82,7 +89,12 @@ CH4 = np.log10(4.4e-3)
 CH4 = np.log10(0.00132)
 alpha = 1.0*sum(x>CH4)/len(x)*1.0
 print "CH4 sig", get_significance(alpha/2)
-print "test", get_significance(0.003/2)
+print "alpha", alpha
+q = [0.997]
+print quantiles(x, q)
+
+
+#print "test", get_significance(0.003/2)
 
 
 plt.subplot(133)

### 2017-08-14 ms.tex
retreival edits

@@
 Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
-We parameterize atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
+We parameterized atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
 We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.
@@
 
-Our nominal model includes a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity, the C/O, and a gray cloud-top pressure.
-We fix the T-P profile morphology but scale the irradiation temperature to allow for the unknown albedo and heat transport efficiency.
+Our nominal model included a temperature-pressure profile (parameterized via the \citealt{guillot10} relations), the atmospheric metallicity, the C/O, a gray cloud-top pressure, and the planet's 10-bar radius.
+We fixed the T-P profile morphology but scaled the irradiation temperature to allow for the unknown albedo and heat transport efficiency.
 We put a uniform prior on the atmospheric metallicity of $0.01 - 30\times$ solar based on the upper limit from $\S$\,\ref{sec:interior}.
 
-The retrieval results are shown in Figure\,\ref{fig:retrieval}.
+The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).
+There is one $3.4\sigma$ outlier data point (1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.
+We tested the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
+These models did not significantly improve the fit quality, and the main retrieval results were unchanged.
+We also tested the effect of varying the planet's surface gravity within the published FIXME $\sigma$ range, and found that our results were again minimally changed.
+
+The nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.
 The metallicity distribution spans the full range allowed by our priors, with some preference for larger values.
 The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence.
-These are not particularly precise constraints because the metallicity and cloud height are degenerate \citep{griffith14}.
+These are not particularly precise constraints because the metallicity and cloud height are somewhat degenerate \citep{griffith14}.
+The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).
 We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.
-The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).
-
-We explored several alternate retrieval scenarios, including the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
-We did not find significant evidence for either additional complexity in the model, and the main retrieval results were unchanged.
 
@@
 We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O.
+Even though water is the dominant absorber in the WFC3 bandpass, there is enough methane opacity to change the shape of the spectral features at a level we can detect.
 In one test, we assumed chemical equilibrium but excluded methane opacity.
@@
 We compared this to the expected methane abundance for the temperature and pressure sensed at the terminator.
-Assuming  solar C/O and $3\,\times$ solar metallicity (the median retrieved value for the chemical equilbrium), the methane abundance is predicted to be $1.3\times10^{-3}$, inconsistent with our estimated value.
+Assuming  solar C/O and $3\,\times$ solar metallicity (the median value for the chemical equilbrium retrieval), the methane abundance is predicted to be $1.3\times10^{-3}$, inconsistent with our estimated value.
 Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.

### 2017-08-14 Analysis/get_COratio.py

@@ -69,6 +69,8 @@ ind = np.argsort(ctoo)
 ctoo = ctoo[ind]
 upperlim = int(np.ceil(0.95*len(ctoo)))
 print 10.**ctoo[upperlim]
+q = [0.997]
+print "3 sig Z", 10**quantiles(ctoo, q)
 
 
 #plt.show()

### 2017-08-14 Figures/fig2_spectrum.py

@@ -21,6 +21,7 @@ plt.errorbar(d[:,0], d[:,1]*100., d[:,2]*100., linestyle = 'none', zorder = 100,
 #plt.plot(d[:,0]*1e6, convolve(d[:,1], g, boundary = 'extend')-0.08)
 
 p = pickle.load(open("WASP107b_Global_Cloud_limit_Metallicity_spectra.pic", "rb"))
+print "outliers", (p[9] - p[7])/p[8]
 w = p[0]
 
 plt.fill_between(w, p[1], p[5], color = "navy", alpha = '0.2')

### 2017-08-14 ms.tex
conclusions

@@
 The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).
-There is one $3.4\sigma$ outlier data point (1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.
+There is one $3.4\sigma$ outlier data point (at 1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.
 We tested the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
@@
 We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O.
-Even though water is the dominant absorber in the WFC3 bandpass, there is enough methane opacity to change the shape of the spectral features at a level we can detect.
+Even though water is the dominant absorber in the WFC3 bandpass, there is enough methane opacity to change the shape of the spectral features at a detectable level for high precision data.
 In one test, we assumed chemical equilibrium but excluded methane opacity.
@@
 In addition to the atmospheric retrieval, we also performed forward modeling of physically motivated, self-consistent clouds and hazes using the methods described in \cite{fortney08, morley15}.
-We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS, see \citealt{morley12}), for a range of metallicites from $1-50\times$ solar.
+We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS; see \citealt{morley12}), for a range of metallicites from $1-50\times$ solar.
 We vary the cloud sedimentation efficiency from 0.1 to 3 (highly lofted, small particulate clouds).
 None of the models produce sufficiently low amplitude spectral features to match the observed spectrum.
+FIXME: what did we assume about surface gravity?
 
@@
 %We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
+\section{Conclusions} \label{sec:discuss}
+We presented an analysis of the atmospheric composition of WASP-107b based on its measured near-infrared transmission spectrum and interior structure modeling of the planet's mass and radius.
+Key results from this analysis include:
 
-\section{Discussion and Conclusions} \label{sec:discuss}
-We combined our high precision transmission spectrum of WASP-107b with interior structure modeling to assess the planet's atmospheric composition.
+\begin{itemize}
+\item{The upper limit on atmospheric metallicity is $30\times$ solar (assuming the star is young) or FIXME solar for an older system.
+Interior structure modeling shows that higher metallicity envelopes would decrease the radius below the observed value.
+Based on the Solar System trend of increasing planet metallicity with decreasing planet mass, WASP-107b is expected to have a metallicity of $30\times$ solar, just at the edge of our allowed range \citep{kreidberg14b}.
+Further characterization of the stellar age is needed to determine whether WASP-107b is consistent with expectations, or discrepant with the Solar System trend, which would suggest an intriguing difference in its formation history.}
+\item{The methane abundance is depleted relative to expectations for a solar abundance pattern.
+This could be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
+Another possibility is that the planet has a hot interior.}
+\item{Optically thick condensates are present at high altitudes ($0.1 - 3$ mbar).
+Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observed .
+Clouds form too deep in the atmosphere }
+\end{itemize}
 
-Upper limit on metallicity.
-The planet's radius is so large that we put an upper limit on the metallicity (additional metals would compress it).
-metallicity 30x solar - how does this compare to mass/atmospheric metallicity
-
-Methane depletion
-low C/O ratio - planetesimal enrichment (cite Nestor)
-internal heat, directly imaged planets
-
-Thick, high-altitude condensate layer
-
-Some directly imaged planets in this temperature regime are also unexpectedly depleted in methane when compared to similar temperature brown dwarfs \citep{skemer14}.
-
+%Some directly imaged planets in this temperature regime are also unexpectedly depleted in methane when compared to similar temperature brown dwarfs \citep{skemer14}.
 The leading hypothesis of this depletion \citep{zahnle14} is due to the CH4-CO quench point occurring in the CO stability field due to the low gravity of self-luminous planets when compared to brown dwarfs.
 
-In order this mechanism to work for neptune massed planet, the internal temperature would have to be near ~500K to get the deep temperatures high enough to thermochemically deplete methane at the quench point.
+%In order this mechanism to work for neptune massed planet, the internal temperature would have to be near ~500K to get the deep temperatures high enough to thermochemically deplete methane at the quench point.
 Given the age of the star, it is unlikely that such high internal temperatures are still due to heat of formation.

### 2017-08-14 ms.tex
conclu

@@
 The features at $1.15$ and $1.4$ $\mu$m are due to water absorption.
-The right-hand axis indicates the normalized atmospheric scale height assuming a $100\times$ solar composition.}
+The right-hand axis indicates the normalized atmospheric scale height assuming a $10\times$ solar composition.}
 \label{fig:spectrum}
@@
 \section{Conclusions} \label{sec:discuss}
-We presented an analysis of the atmospheric composition of WASP-107b based on its measured near-infrared transmission spectrum and interior structure modeling of the planet's mass and radius.
+We analyzed the atmospheric composition of WASP-107b based on its measured near-infrared transmission spectrum and interior structure models of the planet's mass and radius.
 Key results from this analysis include:
@@
 \item{The upper limit on atmospheric metallicity is $30\times$ solar (assuming the star is young) or FIXME solar for an older system.
-Interior structure modeling shows that higher metallicity envelopes would decrease the radius below the observed value.
+Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.
 Based on the Solar System trend of increasing planet metallicity with decreasing planet mass, WASP-107b is expected to have a metallicity of $30\times$ solar, just at the edge of our allowed range \citep{kreidberg14b}.
-Further characterization of the stellar age is needed to determine whether WASP-107b is consistent with expectations, or discrepant with the Solar System trend, which would suggest an intriguing difference in its formation history.}
+Further characterization of the stellar age is needed to determine whether WASP-107b is consistent with expectations, or discrepant with the Solar System trend, which would suggest an intriguing difference in its formation history.
+Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{morley17}.
+Together, these exoplanets suggest a greater diversity of compositions than exists in the Solar System.}
 \item{The methane abundance is depleted relative to expectations for a solar abundance pattern.
-This could be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
-Another possibility is that the planet has a hot interior.}
+This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
+Another possibility is that the planet has a hot interior ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}.
+Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}.}
 \item{Optically thick condensates are present at high altitudes ($0.1 - 3$ mbar).
 Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observed .
-Clouds form too deep in the atmosphere }
+Clouds form too deep in the atmosphere .
+\citep{crossfield17}} GJ 1214b
 \end{itemize}
-
-%Some directly imaged planets in this temperature regime are also unexpectedly depleted in methane when compared to similar temperature brown dwarfs \citep{skemer14}.
-The leading hypothesis of this depletion \citep{zahnle14} is due to the CH4-CO quench point occurring in the CO stability field due to the low gravity of self-luminous planets when compared to brown dwarfs.
-
-%In order this mechanism to work for neptune massed planet, the internal temperature would have to be near ~500K to get the deep temperatures high enough to thermochemically deplete methane at the quench point.
-Given the age of the star, it is unlikely that such high internal temperatures are still due to heat of formation.
-It is possible that tidal heating could increase the internal temperature (Agundez et al.
-2014; Morley et al.
-2017) but that would require a negligible eccentricity, of which is currently uncertain for WASP107b.
 
@@
 In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations.
-
-Combine data from Spitzer/HST
-JWST GTO
 

### 2017-08-14 Analysis/get_COratio.py

@@ -95,13 +95,18 @@ print "alpha", alpha
 q = [0.997]
 print quantiles(x, q)
 
-
 #print "test", get_significance(0.003/2)
-
 
 plt.subplot(133)
 plt.title("NH3")
 x = p[:,5]-6
 hist(x)
 
+#plt.show()
+
+plt.clf()
+ctoo = 10**(p[:,4]-6)/10**(p[:,3]-6)
+hist(np.log10(ctoo))
+plt.axvline(np.log10(0.54), color = 'y')
+print sum(ctoo>0.54), len(ctoo)
 plt.show()

### 2017-08-14 Figures/fig2_spectrum.py

@@ -21,12 +21,17 @@ plt.errorbar(d[:,0], d[:,1]*100., d[:,2]*100., linestyle = 'none', zorder = 100,
 #plt.plot(d[:,0]*1e6, convolve(d[:,1], g, boundary = 'extend')-0.08)
 
 p = pickle.load(open("WASP107b_Global_Cloud_limit_Metallicity_spectra.pic", "rb"))
-print "outliers", (p[9] - p[7])/p[8]
+#print "outliers", (p[9] - p[7])/p[8]
 w = p[0]
 
 plt.fill_between(w, p[1], p[5], color = "navy", alpha = '0.2')
 plt.fill_between(w, p[2], p[4], color = "navy", alpha = '0.3')
-plt.plot(w, p[3], color = 'navy', label = "Best fit model")
+#plt.plot(w, p[3], color = 'navy', label = "Best fit model")
+
+# best fit
+pb = pickle.load(open("Model_best_fit.pic", "rb"))
+plt.plot(pb[0], pb[1]*100., color = 'navy', label = 'Best fit model')
+print 'outliers', (p[7] - pb[3]*100.)/p[8]
 
 """d = pickle.load(open("Model_Solar.pic", "rb"))
 x, y = d[0], d[1]*100. - 0.7

### 2017-08-14 ms.tex
conclu

@@
 Figure\,\ref{fig:metal_prior} shows the results.
-We find that the inferred atmospheric metallicity and core mass are inversely proportional, as expected (the more massive the core, the lighter the envelope must be to maintain the observed radius).
-Larger core masses and envelope metallicities are allowed for the younger stellar age since planets cool and contract as they age.
-For the younger (older) stellar age, we find an upper limit on the core mass of FIXME and an upper limit on atmospheric metallicity of FIXME.
+We find an upper limit on the metallicity of $30\times$ solar for the young stellar age range.
+Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.
+For the older age, the upper limit is even lower (FIXME), because planets cool and contract as they age \citep{fortney08}.
+The inferred metallicity and core mass are inversely proportional, as expected (the more massive the core, the lighter the envelope must be to maintain the observed radius).
+The upper limit on the core mass is FIXME.
 
@@
 However, these photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis, and we found in the previous section that methane is depleted in the atmosphere of WASP-107b.
-%Sulfur-based hazes are another possibility \citep{gao17}.
 Further modeling of the condensates is needed for a satisfactory explanation of the data.
@@
 %We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
-\section{Conclusions} \label{sec:discuss}
-We analyzed the atmospheric composition of WASP-107b based on its measured near-infrared transmission spectrum and interior structure models of the planet's mass and radius.
+\section{Discussion and Conclusions} \label{sec:discuss}
+We analyzed the atmospheric composition of WASP-107b based on retrieval of its near-infrared transmission spectrum and interior structure models of the planet's mass and radius.
 Key results from this analysis include:
@@
 \begin{itemize}
-\item{The upper limit on atmospheric metallicity is $30\times$ solar (assuming the star is young) or FIXME solar for an older system.
-Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.
-Based on the Solar System trend of increasing planet metallicity with decreasing planet mass, WASP-107b is expected to have a metallicity of $30\times$ solar, just at the edge of our allowed range \citep{kreidberg14b}.
-Further characterization of the stellar age is needed to determine whether WASP-107b is consistent with expectations, or discrepant with the Solar System trend, which would suggest an intriguing difference in its formation history.
+\item{\emph{The upper limit on atmospheric metallicity is $30\times$ solar.} Our estimated upper limit is at the edge of consistency with the Solar System metal enrichment pattern, with predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.
+If the star is old (FIXME Gyr), the upper limit on metallicity would be even lower, implying that WASP-107b has a different formation history from the Solar System planets.
+Further characterization of the stellar age is needed to explore this scenario.
 Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{morley17}.
-Together, these exoplanets suggest a greater diversity of compositions than exists in the Solar System.}
+Together, these exoplanets hint at a greater diversity of compositions than exists in the Solar System.}
 \item{The methane abundance is depleted relative to expectations for a solar abundance pattern.
@@
 Another possibility is that the planet has a hot interior ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}.
-Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}.}
+Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}.
+Further observations of the transmission spectrum over a broader wavelength range will refine the C/O estimate and help differentiate between these two scenarios.}
 \item{Optically thick condensates are present at high altitudes ($0.1 - 3$ mbar).
-Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observed .
-Clouds form too deep in the atmosphere .
+Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observations.
+The atmospheric scale height is so large (800 km) that clouds form too deep in the atmosphere to loft to the needed height.
+Hydrocarbon hazes that form at high altitudes can match the data, but these require methane photolysis as a precursor, and methane is depleted in the atmosphere.
+Other haze precursors such as $H_2S$ are a possibility and would be worth exploring in future analyses \citep{gao17}.
 \citep{crossfield17}} GJ 1214b

### 2017-08-14 ms.tex
draft 1

@@
 \begin{abstract}
-The transiting planet WASP-107b is an intriguing target for atmosphere characterization, with a large atmospheric scale height and a small, bright host star.
-
-w107 is a good target for atmospheric characterization, raising the exciting possibility of characterizing its formation history
-
-.
-With a mass similar to Neptune's and a radius close to Jupiter's, it occupies the transition region between ice giants and gas giants.
-The planet is also one of the strongest targets for atmospheric characterization yet discovered, thanks to its small, bright host star and large scale height.
-We report the first atmosphere characterization of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.
-We analyzed the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the planet's mass and radius.
+The Neptune-mass exoplanet WASP-107b is an exciting target for atmosphere characterization.
+It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.
+We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.
+We analyzed the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius.
 We detect water absorption at $6.5\,\sigma$ confidence and set an upper limit on the atmospheric metallicity of $30\times$ solar.
 We find that the methane abundance is depleted relative to expectations for a solar abundance pattern (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.
-The amplitude of features in the transmission spectrum is smaller than expected for a cloudless atmosphere, crossing just one scale height, and requires an opaque condensate layer at high altitudes (0.1 - 3 mbar).
+The amplitude of features in the transmission spectrum is smaller than expected for a cloudless atmosphere, crossing less than one scale height, and requires a thick condensate layer at high altitudes (0.1 - 3 mbar).
 We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at this height.
@@
 This results in a distinct atmospheric chemistry from other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.
-Spectral features from both water and methane are accessible with current observing facilities, raising the possibility of a spectroscopic estimate of the carbon-to-oxygen ratio (C/O).
+Spectral features from both water and methane are accessible with current observing facilities, enabling a spectroscopic estimate of the carbon-to-oxygen ratio (C/O).
 Previous measurements of C/O have been challenging because they rely on photometry that covers absorption features from multiple molecules \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.
-FIXME: maybe cut this.
 
@@
 For the remainder of the target visibility period (about 45 minutes), we obtained time series spectra with the G141 grism, which provides low-resolution spectroscopy over the wavelength range $1.1 - 1.7\,\mu$m.
-We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{https://github.com/spacetelescope/PandExo\_HST}.
+We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{\url{https://github.com/spacetelescope/PandExo\_HST}}.
 As is standard for WFC3 observations of bright targets, we used the spatial scanning observing mode, which slews the telescope in the spatial direction over the course of an exposure.
@@
 We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.
+We also performed an independent data reduction and fit using K.
+Stevenson's pipeline and found consistent results (well within $1\,\sigma$).
 
@@
 We therefore opt not to report transit depths redder than the 1.62 $\mu$m channel.
-
-\subsubsection{Independent Analysis}
-We also performed an independent data reduction and fit using K. Stevenson's pipeline.
 
@@
 \includegraphics[width = 0.5\textwidth]{Figures/metalByMass.png}
-\caption{Estimated envelope metallicity and core mass based on interior structure modeling of the mass and radius of WASP-107b. FIXME ages.}
+\caption{Estimated envelope metallicity and core mass based on interior structure modeling of the mass and radius of WASP-107b. FIXME get new figure showing two ages.}
 \label{fig:metal_prior}
@@
 We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.
+We tested the detection significance for water by removing water opacity from the nominal model.
+The Bayesian evidence favors the inclusion of water at $6.5\,\sigma$ confidence.
 
@@
 None of the models produce sufficiently low amplitude spectral features to match the observed spectrum.
-FIXME: what did we assume about surface gravity?
+The results are sensitive to surface gravity, but even when we increase the gravity by FIXME $\sigma$ the clouds still form too low.
 
@@
 \begin{itemize}
-\item{\emph{The upper limit on atmospheric metallicity is $30\times$ solar.} Our estimated upper limit is at the edge of consistency with the Solar System metal enrichment pattern, with predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.
-If the star is old (FIXME Gyr), the upper limit on metallicity would be even lower, implying that WASP-107b has a different formation history from the Solar System planets.
+\item{\emph{The upper limit on atmospheric metallicity is $30\times$ solar.} This limit is at the edge of consistency with the Solar System metal enrichment pattern, which predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.
+If the star is old ($>1$ Gyr), the upper limit on metallicity would be even lower (FIXME), implying that WASP-107b has a different formation history from the Solar System planets.
 Further characterization of the stellar age is needed to explore this scenario.
 Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{morley17}.
-Together, these exoplanets hint at a greater diversity of compositions than exists in the Solar System.}
-\item{The methane abundance is depleted relative to expectations for a solar abundance pattern.
-This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
+The difference between WASP-107b and GJ 436b suggests a greater diversity of compositions for exoplanets than is present in the Solar System.}
+\item{\emph{The methane abundance is depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
 Another possibility is that the planet has a hot interior ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}.
@@
 Further observations of the transmission spectrum over a broader wavelength range will refine the C/O estimate and help differentiate between these two scenarios.}
-\item{Optically thick condensates are present at high altitudes ($0.1 - 3$ mbar).
-Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observations.
+\item{\emph{Optically thick condensates are present at high altitudes} ($0.1 - 3$ mbar).
+Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observed transmission spectrum.
 The atmospheric scale height is so large (800 km) that clouds form too deep in the atmosphere to loft to the needed height.
-Hydrocarbon hazes that form at high altitudes can match the data, but these require methane photolysis as a precursor, and methane is depleted in the atmosphere.
-Other haze precursors such as $H_2S$ are a possibility and would be worth exploring in future analyses \citep{gao17}.
-\citep{crossfield17}} GJ 1214b
+Hydrocarbon hazes that form at high altitudes can match the data, but these require methane photolysis as a precursor, and we found that methane is depleted.
+Other haze precursors such as H$_2$S are a possibility and would be worth exploring in future analyses \citep{gao17}.
+Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the coolest planets.}
 \end{itemize}
 
-WASP-107b is slated to be observed with numerous facilities.
-Transits were recently observed with the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels.
-Spitzer eclipses are planned (Program FIXME; PI L.
+These results are a first look at the atmosphere of WASP-107b.
+The planet is already being observed at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (HST Program GO 14916, PI J.
+Spake, Spitzer Program 13052, PI M.
+Werner; Spitzer Program 13167, PI L.
 Kreidberg).
-In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations.
+In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations for the NIRISS and NIRSpec instruments\footnote{\url{https://jwst-docs.stsci.edu}}.
+This spate of observing programs is sure to add to the already rich and complex picture of WASP-107b's atmosphere presented here.
 
@@
 
-\vspace{5mm}
-\facilities{HST(WFC3)}
-
-%% Similar to \facility{}, there is the optional \software command to allow 
-%% authors a place to specify which programs were used during the creation of 
-%% the manusscript. Authors should list each code and include either a
-%% citation or url to the code inside ()s when available.
-
-%\software{astropy \citep{2013A&A...558A..33A},  
-%          Cloudy \citep{2013RMxAA..49..137F}, 
-%          SExtractor \citep{1996A&AS..117..393B}
-%          }
-
-%% Appendix material should be preceded with a single \appendix command.
-%% There should be a \section command for each appendix. Mark appendix
-%% subsections with the same markup you use in the main body of the paper.
-
-%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
-%% The equation counter will reset when it encounters the \appendix
-%% command and will number appendix equations (A1), (A2), etc. The
-%% Figure and Table counter will not reset.
-
-%\appendix
-
-%% The reference list follows the main body and any appendices.
-%% Use LaTeX's thebibliography environment to mark up your reference list.
-%% Note \begin{thebibliography} is followed by an empty set of
-%% curly braces.  If you forget this, LaTeX will generate the error
-%% "Perhaps a missing \item?".
-%%
-%% thebibliography produces citations in the text using \bibitem-\cite
-%% cross-referencing. Each reference is preceded by a
-%% \bibitem command that defines in curly braces the KEY that corresponds
-%% to the KEY in the \cite commands (see the first section above).
-%% Make sure that you provide a unique KEY for every \bibitem or else the
-%% paper will not LaTeX. The square brackets should contain
-%% the citation text that LaTeX will insert in
-%% place of the \cite commands.
-
-%% We have used macros to produce journal name abbreviations.
-%% \aastex provides a number of these for the more frequently-cited journals.
-%% See the Author Guide for a list of them.
-
-%% Note that the style of the \bibitem labels (in []) is slightly
-%% different from previous examples.  The natbib system solves a host
-%% of citation expression problems, but it is necessary to clearly
-%% delimit the year from the author name used in the citation.
-%% See the natbib documentation for more details and options.
-
 \bibliographystyle{aasjournal}
@@
 
-%% This command is needed to show the entire author+affilation list when
-%% the collaboration and author truncation commands are used.  It has to
-%% go at the end of the manuscript.
-%\allauthors
-
-%% Include this line if you are using the \added, \replaced, \deleted
-%% commands to see a summary list of all changes at the end of the article.
-%\listofchanges
-
 \end{document}
-
-% End of file `sample61.tex'.

### 2017-08-14 ms.tex
minor edits
edits

@@
 Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
-We can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities.
+The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities.
 \end{abstract}
@@
 The planet is also a feasible target for thermal emission measurements, with an expected secondary eclipse depth of $\sim0.1$\% in \emph{Spitzer}'s $4.5\,\mu$m channel.
-In this paper we report the first atmosphere characterization of WASP-107b: a near-infrared transmission spectrum measured with the \emph{Hubble Space Telescope} (\emph{HST}).
+In this paper we report the first atmosphere characterization of WASP-107b: a near-infrared transmission spectrum measured with the \emph{Hubble Space Telescope} (\emph{HST}; Program GO 14915, PI L.
+Kreidberg).
 
 \section{Observations}
-We observed a single transit of WASP-107b with HST's Wide Field Camera 3 (WFC3) instrument on UT 5-6 June 2017.
-The transit observation consisted of five HST orbits.
+We observed a single transit of WASP-107b with \emph{HST}'s Wide Field Camera 3 (WFC3) instrument on UT 5-6 June 2017.
+The transit observation consisted of five \emph{HST} orbits.
 At the beginning of each 96-minute orbit, we took an image of the target with the F130N filter (exposure time = 4.2 s).
@@
 \item{\emph{Optically thick condensates are present at high altitudes} ($0.1 - 3$ mbar).
-Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observed transmission spectrum.
+The amplitude of the water absorption feature in the transmissin spectrum is roughly half that expected for a clear atmosphere.
+Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the data.
 The atmospheric scale height is so large (800 km) that clouds form too deep in the atmosphere to loft to the needed height.
@@
 These results are a first look at the atmosphere of WASP-107b.
-The planet is already being observed at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (HST Program GO 14916, PI J.
-Spake, Spitzer Program 13052, PI M.
-Werner; Spitzer Program 13167, PI L.
+The planet is already being observed at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (\emph{HST} Program GO 14916, PI J.
+Spake, \emph{Spitzer} Program 13052, PI M.
+Werner; \emph{Spitzer} Program 13167, PI L.
 Kreidberg).

### 2017-08-15 ms.tex
mike edits

@@
 We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.
-We analyzed the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius.
-We detect water absorption at $6.5\,\sigma$ confidence and set an upper limit on the atmospheric metallicity of $30\times$ solar.
+We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius.
+We detect water at $6.5\,\sigma$ confidence and set an upper limit on the atmospheric metallicity of $30\times$ solar.
 We find that the methane abundance is depleted relative to expectations for a solar abundance pattern (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.
-The amplitude of features in the transmission spectrum is smaller than expected for a cloudless atmosphere, crossing less than one scale height, and requires a thick condensate layer at high altitudes (0.1 - 3 mbar).
+The water feature amplitude is smaller than expected for a cloudless atmosphere, crossing less than one scale height, and requires a thick condensate layer at high altitudes (0.1 - 3 mbar).
 We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at this height.
@@
 For each stripe, the extraction box was 80 pixels high and centered on the stripe's midpoint in the spatial direction.
-To correct the change in dispersion solution over the length of the spatial scan, we interpolated each row to the wavelength scale of the row corresponding to the spectral trace.
-We then corrected for slight drift in the spectral direction ($<0.1$ pixels) by interpolating each spectrum to the wavelength scale of the first exposure.
+We corrected for the changing dispersion solution over the length of the spatial scan and small drifts over time ($<0.1$ pixel).
 To subtract the background, we took the median of sky pixels that were uncontaminated by the target spectrum (rows $5-250$, columns $5-15$).
-The typical background counts were low: 40 photoelectrons/pixel, in comparison to $3\times10^4$ photoelectrons/pixel in the stellar spectrum.
 %dispersion solution: 46 Angstroms/pixel. max drift: 3 angstroms (from diagnostics.txt)
@@
 Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
-We parameterized atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon96} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
+We parameterized atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon94} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.
 We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} from the pre-tabulated line-by-line cross section database described in \cite{freedman14}.

### 2017-08-15 ms.tex
edits

@@
 We find that the methane abundance is depleted relative to expectations for a solar abundance pattern (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.
-The water feature amplitude is smaller than expected for a cloudless atmosphere, crossing less than one scale height, and requires a thick condensate layer at high altitudes (0.1 - 3 mbar).
-We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at this height.
+The water features are smaller than expected for a cloudless atmosphere, crossing less than one scale height, and requires a thick condensate layer at high altitudes (0.1 - 3 mbar).
+We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.
 Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
@@
 The composition of a planet's atmosphere depends on where and how the planet formed.
-By measuring atmospheric metallicity and molecular abundance ratios, we can shed light on important aspects of the formation process such as location within the disk and the relative accretion rates of gas versus solids \citep[][]{oberg11, fortney13, madhusudhan14, alidib16, mordasini16, espinoza17}.
+By measuring the inventory of atmospheric elemental abundances, we can shed light on important aspects of the formation process such as location within the disk and the relative accretion rates of gas versus solids \citep[][]{oberg11, fortney13, madhusudhan14, alidib16, mordasini16, espinoza17}.
 
 The warm Neptune WASP-107b is an intriguing target for atmosphere characterization for several reasons.
-It is an intermediate size between ice giants and gas giants, with a mass similar to Neptune's and a radius close to Jupiter's \citep[$0.12\,M_\mathrm{Jup}$, $0.94\,R_\mathrm{Jup}$;][]{anderson17}.
+It has an intermediate size between ice giants and gas giants, with a mass similar to Neptune's and a radius close to Jupiter's \citep[$0.12\,M_\mathrm{Jup}$, $0.94\,R_\mathrm{Jup}$;][]{anderson17}.
 Studying the atmospheres of planets in this transition region will provide additional clues in the much-debated mystery of what stunts the growth of Neptune-size planets \citep[e.g][]{pollack96, dawson16, frelikh17}.
@@
 Spectral features from both water and methane are accessible with current observing facilities, enabling a spectroscopic estimate of the carbon-to-oxygen ratio (C/O).
-Previous measurements of C/O have been challenging because they rely on photometry that covers absorption features from multiple molecules \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.
+Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.
 
@@
 For each stripe, the extraction box was 80 pixels high and centered on the stripe's midpoint in the spatial direction.
-We corrected for the changing dispersion solution over the length of the spatial scan and small drifts over time ($<0.1$ pixel).
+We corrected the spectra for the changing dispersion solution over the length of the spatial scan and small drifts over time ($<0.1$ pixel).
 To subtract the background, we took the median of sky pixels that were uncontaminated by the target spectrum (rows $5-250$, columns $5-15$).
@@
 For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (differing by just $0.3\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
-This offset does not affect our final analysis because the planet's one-bar radius is a free parameter in the atmospheric retrieval (see \S\,\ref{sec:retrieval}).
+This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval (see \S\,\ref{sec:retrieval}).
 We also fit for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values.
@@
 \label{sec:retrieval}
-We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval tool described in \cite{kreidberg15b}.
+We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval \citep{line13a, kreidberg15b}.
 Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
@@
 
-The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).
-There is one $3.4\sigma$ outlier data point (at 1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.
+%There is one $3.4\sigma$ outlier data point (at 1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.
 We tested the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
@@
 The nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.
+The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).
 The metallicity distribution spans the full range allowed by our priors, with some preference for larger values.
 The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence.
-These are not particularly precise constraints because the metallicity and cloud height are somewhat degenerate \citep{griffith14}.
 The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).
@@
 
-We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere, scaling results from previous photochemical models for GJ 436b \citep{line11, morley17}.
-With the most efficient haze production and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.
+We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere.
+We predict the abundance of hydrocarbon haze precursors from previous photochemical models for GJ 436b \citep{line11, morley17}.
+With the most efficient haze production (100\% of precursors form haze) and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.
 However, these photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis, and we found in the previous section that methane is depleted in the atmosphere of WASP-107b.
+Even if the precursors were present, it is unclear whether the 100\% haze production efficiency is realistic.
 Further modeling of the condensates is needed for a satisfactory explanation of the data.
@@
 Further characterization of the stellar age is needed to explore this scenario.
-Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{morley17}.
+Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{moses13, morley17}.
 The difference between WASP-107b and GJ 436b suggests a greater diversity of compositions for exoplanets than is present in the Solar System.}
 \item{\emph{The methane abundance is depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
-Another possibility is that the planet has a hot interior ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}.
+Another possibility is that the planet has a hot interior effective temperature ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}.
 Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}.
@@
 The atmospheric scale height is so large (800 km) that clouds form too deep in the atmosphere to loft to the needed height.
-Hydrocarbon hazes that form at high altitudes can match the data, but these require methane photolysis as a precursor, and we found that methane is depleted.
-Other haze precursors such as H$_2$S are a possibility and would be worth exploring in future analyses \citep{gao17}.
+Efficient hydrocarbon haze production at high altitudes can match the data, but these require methane photolysis as a precursor, and we found that methane is depleted.
+Other haze precursors such as H$_2$S are a possibility and would be worth exploring in future analyses \citep{zahnle09, gao17}.
 Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the coolest planets.}

### 2017-08-16 Analysis/metallicity_v_radius.py
stuff

@@ -1,0 +1,18 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+r = [0.94, 0.38, 0.56]
+
+z = [24.3, 100., 4.8] 
+lo = [24.3, 0., 4.]
+hi = [0., 1000., 21.5]
+
+
+plt.errorbar(r, z, yerr = [lo, hi], fmt = '.k')
+plt.plot(0.94, 12.4, '.k')
+plt.yscale('log')
+
+plt.ylim(0.1, 1000.)
+plt.ylabel("Metallicity (x solar)")
+plt.xlabel("Radius (Rj)")
+plt.show()

### 2017-08-16 ms.tex
todo list

@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{Figures/metalByMass.png}
-\caption{Estimated envelope metallicity and core mass based on interior structure modeling of the mass and radius of WASP-107b. FIXME get new figure showing two ages.}
+\includegraphics[width = 0.5\textwidth]{Figures/metalByAge.pdf}
+\caption{Estimated envelope metallicity for a given core mass based on interior structure modeling.
+The red shading corresponds to the $3\sigma$ confidence interval for a young planet ($0.2-1$ Gyr); the blue is for an older age ($1-14$ Gyr).}
 \label{fig:metal_prior}

### 2017-08-17 ms.tex
v3

@@
 
-\author{Michael Line}
+\author{Michael R. Line}
 \affiliation{Arizona State University}
@@
 
-\author{Caroline Morley}
+\author{Caroline V. Morley}
 \altaffiliation{Sagan Fellow}
-\affiliation{Harvard-Smithsonian Center for Astrophysics}
+%\affiliation{Harvard-Smithsonian Center for Astrophysics}
+\affiliation{Harvard-Smithsonian Center for Astrophysics\
+60 Garden St.\\
+Cambridge, MA 02138}
 \nocollaboration
 
-\author{Kevin Stevenson}
+\author{Kevin B. Stevenson}
 \affiliation{Space Telescope Science Institute}
@@
 We find that the methane abundance is depleted relative to expectations for a solar abundance pattern (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.
-The water features are smaller than expected for a cloudless atmosphere, crossing less than one scale height, and requires a thick condensate layer at high altitudes (0.1 - 3 mbar).
+The water features are smaller than expected for a cloudless atmosphere, crossing less than one scale height, and require a thick condensate layer at high altitudes (0.1 - 3 mbar).
 We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.
@@
 
-WASP-107b also has a relatively low equilibrium temperature (780 K) compared to most other exoplanets that are amenable to atmosphere characterization.
+WASP-107b also has a relatively low equilibrium temperature compared to most other exoplanets that are amenable to atmosphere characterization (780 K, assuming zero albedo).
 This results in a distinct atmospheric chemistry from other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.
@@
 Thanks to its large atmospheric scale height and small, bright host star, the expected signal-to-noise for the transmission spectrum is comparable to the best studied benchmarks in the field (e.g.
-HD\,209458b, HD\,189733b).
-The planet is also a feasible target for thermal emission measurements, with an expected secondary eclipse depth of $\sim0.1$\% in \emph{Spitzer}'s $4.5\,\mu$m channel.
+HD\,209458b).
 In this paper we report the first atmosphere characterization of WASP-107b: a near-infrared transmission spectrum measured with the \emph{Hubble Space Telescope} (\emph{HST}; Program GO 14915, PI L.
@@
 
-\section{Observations}
+\section{Observations and Data Reduction}
 We observed a single transit of WASP-107b with \emph{HST}'s Wide Field Camera 3 (WFC3) instrument on UT 5-6 June 2017.
@@
 
-\section{Data Reduction}
 We reduced the data with the custom pipeline described in \cite{kreidberg14a}, which we summarize briefly here.
@@
 To create the raw white light curve, we summed each spectrum over the 181 pixels in the spectral trace.
-The white light curve has systematic trends that are typical for WFC3 observations \citep{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend.
+The white light curve has systematic trends that are typical for WFC3 observations \citep{zhou17}: the flux increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend.
 The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.
 We fit the light curve with the analytic model of the form $F_\mathrm{white}(t) = S_\mathrm{white}(t)\times T_\mathrm{white}(t)$, where $S_\mathrm{white}$ is a systematics model and $T_\mathrm{white}$ is a transit model.
-We used the same systematics model as \cite{kreidberg15b}, Equation 1.
-To model the transit, we used the \texttt{batman} package \citep{kreidberg15a}.
-The transit model parameters are the orbital period $p$, time of inferior conjunction $t_0$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.
+We used the same systematics model as \cite{kreidberg15b}.
+We modelled the transit with the \texttt{batman} package \citep{kreidberg15a}.
+The model parameters are the orbital period $p$, time of inferior conjunction $t_0$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.
 
@@
 \includegraphics[width = \textwidth]{Figures/lc.pdf}
-\caption{Broadband and spectrophotometric transit light curves for WASP-107b measured with \textit{HST}/WFC3 compared to best fit models (left) and residuals from the best fit (right).
-Annotations indicate the central wavelength and root-mean-square (rms) residuals for each light curve.
-A possible star-spot crossing feature is shaded in gray in the upper right; our systematic error correction removes this feature from the spectroscopic light curves.}
+\caption{Left: Broadband and spectrophotometric transit light curves for WASP-107b compared to best fit models.
+Right: residuals from the best fit.
+Annotations indicate the central wavelength and root-mean-square (rms) residuals.
+A star-spot crossing feature is shaded in gray in the upper right; our systematic error correction removes this feature from the spectroscopic light curves.}
 \label{fig:lc}
@@
 In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \citep{dai17}.
-These values are consistent with our best fit when we allowed them to vary freely.
 We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star with effective temperature 4300 K, calculated with the \texttt{limb-darkening} package from \cite{espinoza15}.
-The remaining free parameters were $t_0$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
-%The model values are consistent with what we obtain from fitting for the limb darkening coefficients as free parameters.
+We checked that the values we fixed were consistent with our estimates when we allowed them to vary freely.
+We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.
+The remaining free parameters in the fit were $t_0$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
 
@@
 The best fit light curves have a median $\chi^2_\nu$ value of 1.16.
-To ensure that we did not underestimate the uncertainty on the transit depths, we rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2_\nu$ values are unity.
+We conservatively rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2_\nu$ values are unity.
 We performed an MCMC fit to the light curves with \texttt{emcee}.
@@
 We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum.
-For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (differing by just $0.3\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
+For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths.
+%(differing by just $0.3\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
 This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval (see \S\,\ref{sec:retrieval}).
@@
 We found that the fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths by about 10\%.
-We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.
 We also performed an independent data reduction and fit using K.
@@
 
-\subsubsection{Transit Depths Redder than 1.62 $\mu$m}
-The red edge of the transmission spectrum is of interest for our analysis because methane is expected to be the dominant absorber over water at wavelengths greater than $1.6\,\mu$m.
+%\subsubsection{Transit Depths Redder than 1.62 $\mu$m}
+%The red edge of the transmission spectrum is of interest for our analysis because methane is expected to be the dominant absorber over water at wavelengths greater than $1.6\,\mu$m.
 Unfortunately, we find that the our reddest spectroscopic light curves ($>1.62\mu$m) are too poor in quality to robustly measure the transit depth.
@@
 The features at $1.15$ and $1.4$ $\mu$m are due to water absorption.
-The right-hand axis indicates the normalized atmospheric scale height assuming a $10\times$ solar composition.}
+The right-hand axis indicates the normalized atmospheric scale height assuming a $10\times$ solar composition.
+The water feature crosses less than one scale height, indicating that condensates are present at high altitude.}
 \label{fig:spectrum}
@@
 These models assume a thermally inert heavy-element core with a convective envelope of additively mixed H/He \citep{saumon95} and heavy-element impurities.
-The heavy elements were a 50-50 rock-ice mix from ANEOS \citep{thompson90}.
+The heavy elements were a 50-50 rock-ice mix.
 We evolved the planets in time using the atmospheric models of \cite{fortney07}.
@@
 Based on these assumptions, we fit for envelope metallicity and core mass using an MCMC with uniform priors on both parameters.
-The MCMC burned in for $10^3$ steps and then collected $4\times10^6$ samples, which we thinned to $10^5$.
-The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was 18 (the value for water), using the approach of \cite{fortney13}.
+The MCMC burned in for $10^3$ steps and then collected $4\times10^6$ samples.
+The envelope metal mass fractions were converted to metallicities assuming the mean molecular weight of the metals was 18 (the value for water), using the approach of \cite{fortney13}.
 Figure\,\ref{fig:metal_prior} shows the results.
@@
 Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.
-For the older age, the upper limit is even lower (FIXME), because planets cool and contract as they age \citep{fortney08}.
-The inferred metallicity and core mass are inversely proportional, as expected (the more massive the core, the lighter the envelope must be to maintain the observed radius).
-The upper limit on the core mass is FIXME.
+For the older age, the upper limit is even lower ($20\times$ solar), because planets cool and contract as they age \citep{fortney08}.
+These are conservative upper limits, but realistically, the planet probably formed with a core.
+Assuming a $5\,M_\oplus$ core, the upper limit on metallicity is 20 (10) $M_\oplus$ for the young (old) age.
 
@@
 \caption{Estimated envelope metallicity for a given core mass based on interior structure modeling.
-The red shading corresponds to the $3\sigma$ confidence interval for a young planet ($0.2-1$ Gyr); the blue is for an older age ($1-14$ Gyr).}
+The red shading corresponds to the $1\sigma$ confidence interval for a young planet ($0.2-1$ Gyr); the blue is for an older age ($1-14$ Gyr).}
 \label{fig:metal_prior}
@@
 \label{sec:retrieval}
-We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval \citep{line13a, kreidberg15b}.
+We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval \citep{line13a}.
 Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.
@@
 %There is one $3.4\sigma$ outlier data point (at 1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.
-We tested the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.
-These models did not significantly improve the fit quality, and the main retrieval results were unchanged.
-We also tested the effect of varying the planet's surface gravity within the published FIXME $\sigma$ range, and found that our results were again minimally changed.
 
@@
 The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).
-The metallicity distribution spans the full range allowed by our priors, with some preference for larger values.
+The metallicity distribution spans the full range allowed by our priors, with preference for larger values.
 The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence.
@@
 
+We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.
+These models did not significantly improve the fit quality, and the main retrieval results were unchanged.
+We also explored cases without the $30\times$ solar metallicity prior upper limit.
+Without this prior, metallicities up to $100\times$ were permitted, but other parameters were not significantly changed.
+%We also tested the effect of varying the planet's surface gravity within the published FIXME $\sigma$ range, and found that our results were again minimally changed.
+
 \begin{figure}
@@
 \caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure in bars.
-Dashed vertical lines show the median and $\pm1\,\sigma$ confidence interval.
-Solar metallicity and C/O values are indicated with solid vertical lines.}  \label{fig:retrieval}
+Black vertical lines show the median and $\pm1\,\sigma$ confidence interval (solid and dashed, respectively).
+Solar metallicity and C/O values are indicated with red vertical lines.}  \label{fig:retrieval}
 \end{figure}
@@
 In one test, we assumed chemical equilibrium but excluded methane opacity.
-This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$), resolving the tension with the solar value.
+This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$).
 
 We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium.
-The $3\,\sigma$ upper limit on the methane abundance is $1.2\times10^{-3}$.
-We compared this to the expected methane abundance for the temperature and pressure sensed at the terminator.
-Assuming  solar C/O and $3\,\times$ solar metallicity (the median value for the chemical equilbrium retrieval), the methane abundance is predicted to be $1.3\times10^{-3}$, inconsistent with our estimated value.
+The retrieved methane abundance is low ($3\times10^{-11} - 3\times10^{-6}$ at 68\% confidence).
+We rule out at $3\,\sigma$ confidence the predicted methane abundance for a composition with solar C/O and $3\,\times$ solar metallicity (the median metallicity from the chemical equilibrium retrieval).
 Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.
+By contrast, the water abundance ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions (FIXME).
 
@@
 We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS; see \citealt{morley12}), for a range of metallicites from $1-50\times$ solar.
-We vary the cloud sedimentation efficiency from 0.1 to 3 (highly lofted, small particulate clouds).
+We vary the cloud sedimentation efficiency from 3 to 0.1 (normal to highly lofted, small particulate clouds).
 None of the models produce sufficiently low amplitude spectral features to match the observed spectrum.
-The results are sensitive to surface gravity, but even when we increase the gravity by FIXME $\sigma$ the clouds still form too low.
+This result suggests that if clouds are muting WASP-107b's transmission spectrum, the mechanism for cloud particle formation and vertical lofting is very efficient and not captured by the modeling used here.
+Futher studies of cloud formation in very low gravity environments are required.
 
 We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere.
-We predict the abundance of hydrocarbon haze precursors from previous photochemical models for GJ 436b \citep{line11, morley17}.
+We predict the abundance of hydrocarbon haze precursors from previously published photochemical models for GJ 436b \citep{line11, morley17}.
 With the most efficient haze production (100\% of precursors form haze) and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.
-However, these photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis, and we found in the previous section that methane is depleted in the atmosphere of WASP-107b.
-Even if the precursors were present, it is unclear whether the 100\% haze production efficiency is realistic.
-Further modeling of the condensates is needed for a satisfactory explanation of the data.
 
-%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
-%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.
+These photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis; however, we found in the previous section that methane is depleted.
+Furthermore, even if the precursors were present, it is unclear whether the 100\% haze production efficiency is realistic.
+However, organic hydrocarbons are not the only possible photolytic haze in cool H2-rich atmospheres.
+Sulfur chemistry may create additional haze material \citep{zahnle09, gao17}.
+Further modeling and laboratory work of the hazes that form in these conditions is needed for a satisfactory explanation of the data.
+
 \section{Discussion and Conclusions} \label{sec:discuss}
@@
 \begin{itemize}
-\item{\emph{The upper limit on atmospheric metallicity is $30\times$ solar.} This limit is at the edge of consistency with the Solar System metal enrichment pattern, which predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.
-If the star is old ($>1$ Gyr), the upper limit on metallicity would be even lower (FIXME), implying that WASP-107b has a different formation history from the Solar System planets.
-Further characterization of the stellar age is needed to explore this scenario.
-Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{moses13, morley17}.
-The difference between WASP-107b and GJ 436b suggests a greater diversity of compositions for exoplanets than is present in the Solar System.}
+\item{\emph{The upper limit on atmospheric metallicity is $30\times$ solar.} This limit is at the edge of consistency with the Solar System metal enrichment trend, which predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.
+Compared to results for other exoplanets of similar mass such as GJ 436b, which has a high metallicity ($>100\times$ solar) and HAT-P-26b, which is metal-poor compared to the Solar System trend, WASP-107b adds to the evidence that exoplanets exhibit a greater diversity of compositions than is present in the Solar System \citep{morley17, wakeford17}.}
 \item{\emph{The methane abundance is depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
@@
 \item{\emph{Optically thick condensates are present at high altitudes} ($0.1 - 3$ mbar).
-The amplitude of the water absorption feature in the transmissin spectrum is roughly half that expected for a clear atmosphere.
+The amplitude of the water absorption feature in the transmission spectrum is less than a third that expected for a clear atmosphere.
 Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the data.
-The atmospheric scale height is so large (800 km) that clouds form too deep in the atmosphere to loft to the needed height.
-Efficient hydrocarbon haze production at high altitudes can match the data, but these require methane photolysis as a precursor, and we found that methane is depleted.
-Other haze precursors such as H$_2$S are a possibility and would be worth exploring in future analyses \citep{zahnle09, gao17}.
-Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the coolest planets.}
+All the cloud models  we consider produce condensation too deep in the atmosphere.
+Efficient hydrocarbon haze production at high altitudes can match the data, but these require the presence of methane to serve as a haze precursor.
+Other precursors such as H$_2$S are a possibility and would be worth exploring in future analyses.
+Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the atmospheres of the coolest planets.}
 \end{itemize}
@@
 Kreidberg).
-In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations for the NIRISS and NIRSpec instruments\footnote{\url{https://jwst-docs.stsci.edu}}.
+In addition, WASP-107b is included in the \emph{JWST} Guaranteed Time Observations for the NIRISS, NIRCAM, and NIRSpec instruments\footnote{\url{https://jwst-docs.stsci.edu}}.
 This spate of observing programs is sure to add to the already rich and complex picture of WASP-107b's atmosphere presented here.
@@
 \acknowledgments
-We thank Fei Dai, Jessica Spake, and Ian Crossfield for productive conversations.
-L.R.K.
+We thank Fei Dai, Jessica Spake, Ian Crossfield, Hannah Diamond-Lowe, and Jonathan Fortney for productive conversations.
+L.K.
 acknowledges support from the Harvard Society of Fellows and the Harvard Astronomy Department Institute for Theory and Computation.
+C.V.M.
+acknowledges support from NASA through the Sagan Fellowship Program.
 

### 2017-08-17 Analysis/get_COratio.py

@@ -82,6 +82,8 @@ plt.subplot(131)
 plt.title("H2O")
 x = p[:,3]-6
 hist(x)
+q = [0.16, 0.84]
+print "68percent conf h2o", 10**quantiles(x, q)
 
 plt.subplot(132)
 plt.title("CH4")
@@ -94,6 +96,8 @@ print "CH4 sig", get_significance(alpha/2)
 print "alpha", alpha
 q = [0.997]
 print quantiles(x, q)
+q = [0.16, 0.84]
+print "68percent conf ch4", 10**quantiles(x, q)
 
 #print "test", get_significance(0.003/2)
 

### 2017-08-17 Figures/retrieval.py

@@ -34,15 +34,16 @@ plt.xlabel("[Fe/H]")
 plt.ylabel("Frequency")
 plt.xlim(-2, np.max(d[:,1]))
 
-plt.gca().text(0.08, 0.95, "a", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
+plt.gca().text(0.07, 0.95, "a", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
+plt.gca().set_xticks([-2, -1, 0, 1])
 
 lo, med, hi = quantiles(d[:,1], q)
 print "metal", lo, med, hi
 plt.axvline(lo, color = '0.1', linestyle = 'dashed')
-plt.axvline(med, color = '0.1', linestyle = 'dashed')
+plt.axvline(med, color = '0.1', linestyle = 'solid')
 plt.axvline(hi, color = '0.1', linestyle = 'dashed')
 plt.axvline(0., color = '#c44240', linestyle = 'solid')
-plt.gca().text(0.52, 0.7, "solar", transform=plt.gca().transAxes, fontsize=18, va='top', rotation = 90)
+plt.gca().text(0.49, 0.7, "solar", transform=plt.gca().transAxes, fontsize=18, va='top', rotation = 90)
 
 ##############################
 
@@ -51,15 +52,16 @@ hist(d[:,2], histtype ='stepfilled', color = 'k', facecolor='#b7c9e2', bins = nb
 
 plt.gca().get_yaxis().set_visible(False)		
 plt.xlabel("log(C/O)")
-plt.gca().text(0.08, 0.95, "b", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
+plt.gca().text(0.07, 0.95, "b", transform=plt.gca().transAxes, fontsize=18, fontweight='bold', va='top')
 
 lo, med, hi = quantiles(d[:,2], q)
 print "C/O", 10**lo, 10**med, 10**hi
 plt.axvline(lo, color = '0.1', linestyle = 'dashed')
-plt.axvline(med, color = '0.1', linestyle = 'dashed')
+plt.axvline(med, color = '0.1', linestyle = 'solid')
 plt.axvline(hi, color = '0.1', linestyle = 'dashed')
 plt.axvline(np.log10(0.54), color = '#c44240', linestyle = 'solid')
[…]

### 2017-08-18 ms.tex
edits

@@
 %% to the KEY in the \bibitem in the reference list below.
-
-%Figures and Tables:
-% 1.  transit light curves
-% 2.  best fit spectrum
-% 3.  retrieval results
-% 4.  table of transit depths
-% 5.  JWST predictions for cloud models
 
@@
 We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.
+We also varied the assumed planet mass by $3\,sigma$.
 These models did not significantly improve the fit quality, and the main retrieval results were unchanged.
@@
 We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium.
-The retrieved methane abundance is low ($3\times10^{-11} - 3\times10^{-6}$ at 68\% confidence).
-We rule out at $3\,\sigma$ confidence the predicted methane abundance for a composition with solar C/O and $3\,\times$ solar metallicity (the median metallicity from the chemical equilibrium retrieval).
+The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, ruling out the the predicted methane content for a composition with solar C/O and $3\,\times$ solar metallicity (the median metallicity from the chemical equilibrium retrieval).
 Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.
-By contrast, the water abundance ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions (FIXME).
+By contrast, the water abundance ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions for solar composition.
 
@@
 Furthermore, even if the precursors were present, it is unclear whether the 100\% haze production efficiency is realistic.
-However, organic hydrocarbons are not the only possible photolytic haze in cool H2-rich atmospheres.
-Sulfur chemistry may create additional haze material \citep{zahnle09, gao17}.
+However, organic hydrocarbons are not the only possible photolytic haze in cool H2-rich atmospheres, as sulfur chemistry may create additional haze material \citep{zahnle09, gao17}.
 Further modeling and laboratory work of the hazes that form in these conditions is needed for a satisfactory explanation of the data.

### 2017-08-21 ms.tex
edit
kevin's edits

@@
 The composition of a planet's atmosphere depends on where and how the planet formed.
-By measuring the inventory of atmospheric elemental abundances, we can shed light on important aspects of the formation process such as location within the disk and the relative accretion rates of gas versus solids \citep[][]{oberg11, fortney13, madhusudhan14, alidib16, mordasini16, espinoza17}.
+By measuring the inventory of atmospheric elemental abundances, we can shed light on important aspects of the formation process such as location within the disk and the relative accretion rates of gas versus solids \citep[][]{oberg11, fortney13, madhusudhan14,  mordasini16, espinoza17}.
 
@@
 For the remainder of the target visibility period (about 45 minutes), we obtained time series spectra with the G141 grism, which provides low-resolution spectroscopy over the wavelength range $1.1 - 1.7\,\mu$m.
-We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{\url{https://github.com/spacetelescope/PandExo\_HST}}.
+We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool \citep{batalha17}.
 As is standard for WFC3 observations of bright targets, we used the spatial scanning observing mode, which slews the telescope in the spatial direction over the course of an exposure.
@@
 For each exposure, we extracted the spectrum from each up-the-ramp sample (or ``stripe") separately using the optimal extraction algorithm of \cite{horne86}.
+We estimated the background from the median count level of pixels uncontaminated by the target spectrum (rows $5-250$, columns $5-515$).
 The stripe spectra were then summed to create the final spectrum.
 For each stripe, the extraction box was 80 pixels high and centered on the stripe's midpoint in the spatial direction.
-We corrected the spectra for the changing dispersion solution over the length of the spatial scan and small drifts over time ($<0.1$ pixel).
-To subtract the background, we took the median of sky pixels that were uncontaminated by the target spectrum (rows $5-250$, columns $5-15$).
+We corrected the spectra for the changing dispersion solution over this aperture and small drifts over time ($<0.1$ pixel).
 %dispersion solution: 46 Angstroms/pixel. max drift: 3 angstroms (from diagnostics.txt)
@@
 As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model.
-The final spectroscopic light curve fits had just three free parameters: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
+The final spectroscopic light curve fits had just three free parameters per channel: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
 
@@
 The features at $1.15$ and $1.4$ $\mu$m are due to water absorption.
-The right-hand axis indicates the normalized atmospheric scale height assuming a $10\times$ solar composition.
+The right-hand axis indicates the normalized atmospheric scale height assuming solar composition.
 The water feature crosses less than one scale height, indicating that condensates are present at high altitude.}
@@
 \begin{itemize}
-\item{\emph{The upper limit on atmospheric metallicity is $30\times$ solar.} This limit is at the edge of consistency with the Solar System metal enrichment trend, which predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.
+\item{\emph{The upper limit on atmospheric metallicity from interior structure modeling is $30\times$ solar.} This limit is at the edge of consistency with the Solar System metal enrichment trend, which predicts a metallicity of $30\times$ solar for WASP-107b \citep{kreidberg14b}.
 Compared to results for other exoplanets of similar mass such as GJ 436b, which has a high metallicity ($>100\times$ solar) and HAT-P-26b, which is metal-poor compared to the Solar System trend, WASP-107b adds to the evidence that exoplanets exhibit a greater diversity of compositions than is present in the Solar System \citep{morley17, wakeford17}.}

### 2017-08-21 Figures/combined_spectra.py

@@ -17,6 +17,9 @@ plt.figure(figsize=(8,4))
 d= np.genfromtxt("w107_spectrum_06-22-17.txt")
 plt.errorbar(d[:,0], d[:,1]*100., d[:,2]*100., linestyle = 'none', zorder = 100, marker = 'o', color = 'w', markeredgecolor = 'k', ecolor ='k', markeredgewidth = 1.3)
 
+d0 = np.genfromtxt("w107_g102_prelim_081817.txt")
+plt.errorbar(d0[:,0], 100*d0[:,3]**2, 100.*2.*d0[:,3]*d0[:,4], fmt ='.k')
+
 #d = np.genfromtxt("wasp107b_10xsolar_0.1mbarcloud.dat")
 #plt.plot(d[:,0]*1e6, convolve(d[:,1], g, boundary = 'extend')-0.08)
 
@@ -43,14 +46,12 @@ plt.ylim(2.03, 2.1)
 plt.xlabel("Wavelength (microns)")
 
 x2 = plt.gca().twinx()
-scaleheight = 520./1.e6*100.
-plt.plot(w,p[3]/scaleheight - np.min(p[3]/scaleheight)+0.06, linewidth=0.)
+scaleheight = 460./1.e6*100.
+plt.plot(w,p[3]/scaleheight - np.min(p[3]/scaleheight), linewidth=0.)
 
-plt.ylabel("Scale height")
-plt.xlim(1.05, 1.7)
-
+plt.ylabel("Scale heights")
+plt.xlim(0.8, 1.7)
 
 plt.tight_layout()
-plt.savefig("spectrum.pdf")
 plt.show()
 

### 2017-08-21 Figures/plot_lc.py

@@ -38,7 +38,7 @@ ax1.set_ylim(0.973, 1.005)
 ax1.yaxis.set_major_locator(FixedLocator(np.array([0.98, 0.99, 1.0])))
 ax1.set_yticklabels(["0.98", "0.99", "1.0"])
 
-ax1.text(2.1, 0.985,'$\lambda$ ($\mu$m):\n1.1 - 1.7', fontsize=10)
+ax1.text(2.1, 0.985,'$\lambda$ ($\mu$m):\n1.12 - 1.65', fontsize=10)
 ax2 = plt.subplot(gs[0,1])
 
 ind = model.phase > 0.5

### 2017-08-22 ms.tex
updates

@@
 In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \citep{dai17}.
-We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star with effective temperature 4300 K, calculated with the \texttt{limb-darkening} package from \cite{espinoza15}.
+We also fixed quadratic limb darkening parameters $u_1$ and $u_2$ on predictions from a PHOENIX model for a star with effective temperature 4300 K, calculated with the \texttt{limb-darkening} package from \cite{espinoza15}.
 We checked that the values we fixed were consistent with our estimates when we allowed them to vary freely.
@@
 Figure\,\ref{fig:metal_prior} shows the results.
-We find an upper limit on the metallicity of $30\times$ solar for the young stellar age range.
+We find a $3\,\sigma$ upper limit on the metallicity of $30\times$ solar for the young stellar age range.
 Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.

### 2017-08-22 Figures/format_table.py

@@ -3,8 +3,10 @@ import numpy as np
 d = np.genfromtxt("w107_spectrum_06-22-17.txt")
 
 n = len(d[:,0])
+bins = np.linspace(1.12, 1.65, 21)
 
 print np.mean(d[:,1])
 for i in range(n):
 #	print '{0:0.3f}'.format(d[i,0]), '&', '{0:0.0f}'.format((d[i,1] - np.mean(d[:,1]))*1.e6), '&', '{0:0.0f}'.format(d[i,2]*1e6), '\\\\' 
-	print '{0:0.3f}'.format(d[i,0]), '&', '{0:0.6f}'.format(d[i,1]), '&', '{0:0.1e}'.format(d[i,2]), '\\\\' 
+	#print '{0:0.3f}'.format(d[i,0]), '&', '{0:0.6f}'.format(d[i,1]), '&', '{0:0.1e}'.format(d[i,2]), '\\\\' 
+	print '{0:0.3f}'.format(bins[i]), '--', '{0:0.3f}'.format(bins[i+1]), '&', '{0:0.6f}'.format(d[i,1]), '&', '{0:0.1e}'.format(d[i,2]), '\\\\' 

### 2017-08-22 Figures/plot_lc.py

@@ -38,7 +38,8 @@ ax1.set_ylim(0.973, 1.005)
 ax1.yaxis.set_major_locator(FixedLocator(np.array([0.98, 0.99, 1.0])))
 ax1.set_yticklabels(["0.98", "0.99", "1.0"])
 
-ax1.text(2.1, 0.985,'$\lambda$ ($\mu$m):\n1.12 - 1.65', fontsize=10)
+#ax1.text(2.1, 0.985,'$\lambda$ ($\mu$m):\n1.12 - 1.65', fontsize=10)
+ax1.text(2., 0.992,'broadband', fontsize=10)
 ax2 = plt.subplot(gs[0,1])
 
 ind = model.phase > 0.5
@@ -58,8 +59,9 @@ ax2.set_xlabel("Time from central transit (hours)")
 #ax2 = ax1.twinx()
 ax3 = plt.subplot(gs[1,0])
 
-#ax3.text(2.5, 1.002,'$\lambda$ ($\mu$m)', fontsize=10)
+ax3.text(2.5, 1.0028,'$\lambda$ ($\mu$m):', fontsize=10)
 
+#ax1.text(2.1, 0.985,'$\lambda$ ($\mu$m):\n1.12 - 1.65', fontsize=10)
 for i, f in enumerate(files):
 	p = pickle.load(open(f, 'rb'))
 	data, model = p[0], p[1]

### 2017-08-22 ms.tex
final version?

@@
 We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius.
-We detect water at $6.5\,\sigma$ confidence and set an upper limit on the atmospheric metallicity of $30\times$ solar.
-We find that the methane abundance is depleted relative to expectations for a solar abundance pattern (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.
-The water features are smaller than expected for a cloudless atmosphere, crossing less than one scale height, and require a thick condensate layer at high altitudes (0.1 - 3 mbar).
-We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.
+The interior structure models set a $3\,\sigma$ upper limit on the atmospheric metallicity of $30\times$ solar.
+The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and we infer a water abundance consistent with expectations for a solar abundance pattern.
+On the other hand, methane is depleted relative to expectations (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.
+The water features are smaller than predicted for a cloudless atmosphere, crossing less than one scale height.
+A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations; however, we find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.
 Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
@@
 
-%% Keywords should appear after the \end{abstract} command. 
-%% See the online documentation for the full list of available subject
-%% keywords and the rules for their use.
 \keywords{planets and satellites: individual (WASP-107b), planets and satellites: atmospheres}
-
-%% From the front matter, we move on to the body of the paper.
-%% Sections are demarcated by \section and \subsection, respectively.
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-
-%% We recommend that authors also use the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below.
 
@@
 This direct image is used for wavelength calibration.
-For the remainder of the target visibility period (about 45 minutes), we obtained time series spectra with the G141 grism, which provides low-resolution spectroscopy over the wavelength range $1.1 - 1.7\,\mu$m.
+For the remainder of the target visibility period, we obtained time series spectra with the G141 grism, which provides low-resolution spectroscopy over the wavelength range $1.1 - 1.7\,\mu$m.
 We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool \citep{batalha17}.
@@
 \subsection{Spectroscopic Light Curve Fits}
-We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.65 $\mu$m, shown in Figure\,\ref{fig:lc}.
+We binned the spectrum into 20 spectrophotometric channels from 1.12 to 1.63 $\mu$m, shown in Figure\,\ref{fig:lc}.
 We fit the light curves with the \texttt{divide-white} technique, which assumes that the light curve systematics have the same morphology at all wavelengths \citep{stevenson14c, kreidberg14a}.
@@
 The amplitude of the feature has no detectable wavelength dependence at the level of precision of our data.
-As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model.
+As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model (coefficients listed in Table\,\ref{tab:transit_depths}.
 The final spectroscopic light curve fits had just three free parameters per channel: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
@@
 
-\begin{deluxetable}{c c c}
-\tablecaption{WASP-107b transmission spectrum \label{tab:transit_depths}}
-\tablehead{\colhead{Wavelength ($\mu$m)}  & \colhead{Transit depth} & \colhead{Uncertainty}}
+\begin{deluxetable}{c c c c c}
+\tablecaption{WASP-107b transmission spectrum and limb darkening coefficients\label{tab:transit_depths}}
+\tablehead{\colhead{Wavelength ($\mu$m)}  & \colhead{Transit depth} & \colhead{Uncertainty} & \colhead{$u_1$} & \colhead{$u_2$}}
 \startdata
-1.133 & 0.020641 & 5.9e-05 \\
-1.158 & 0.020733 & 5.5e-05 \\
-1.184 & 0.020505 & 5.6e-05 \\
-1.209 & 0.020498 & 5.4e-05 \\
-1.235 & 0.020455 & 5.9e-05 \\
-1.260 & 0.020492 & 5.0e-05 \\
-1.285 & 0.020620 & 6.2e-05 \\
-1.311 & 0.020739 & 5.0e-05 \\
-1.336 & 0.020660 & 5.7e-05 \\
-1.362 & 0.020858 & 4.8e-05 \\
-1.387 & 0.020794 & 4.8e-05 \\
-1.413 & 0.020888 & 5.2e-05 \\
-1.438 & 0.020821 & 6.2e-05 \\
-1.464 & 0.020691 & 5.1e-05 \\
-1.489 & 0.020682 & 6.9e-05 \\
-1.515 & 0.020679 & 6.7e-05 \\
-1.540 & 0.020509 & 6.0e-05 \\
-1.565 & 0.020480 & 6.4e-05 \\
-1.591 & 0.020500 & 5.6e-05 \\
-1.616 & 0.020514 & 6.5e-05 \\
+1.121 -- 1.145 & 0.020641 & 5.9e-05 & 0.38 & 0.13 \\
+1.145 -- 1.171 & 0.020733 & 5.5e-05 & 0.37 & 0.14 \\
+1.171 -- 1.196 & 0.020505 & 5.6e-05 & 0.36 & 0.14 \\
+1.196 -- 1.222 & 0.020498 & 5.4e-05 & 0.36 & 0.15 \\
+1.222 -- 1.248 & 0.020455 & 5.9e-05 & 0.36 & 0.15 \\
+1.248 -- 1.272 & 0.020492 & 5.0e-05 & 0.35 & 0.16 \\
+1.272 -- 1.298 & 0.020620 & 6.2e-05 & 0.34 & 0.17 \\
+1.298 -- 1.323 & 0.020739 & 5.0e-05 & 0.34 & 0.17 \\
+1.323 -- 1.349 & 0.020660 & 5.7e-05 & 0.33 & 0.18 \\
+1.349 -- 1.374 & 0.020858 & 4.8e-05 & 0.32 & 0.19 \\
+1.374 -- 1.401 & 0.020794 & 4.8e-05 & 0.31 & 0.20 \\
+1.401 -- 1.425 & 0.020888 & 5.2e-05 & 0.30 & 0.21 \\
+1.425 -- 1.452 & 0.020821 & 6.2e-05 & 0.29 & 0.21 \\
+1.452 -- 1.476 & 0.020691 & 5.1e-05 & 0.28 & 0.22 \\
+1.476 -- 1.502 & 0.020682 & 6.9e-05 & 0.26 & 0.23 \\
+1.502 -- 1.528 & 0.020679 & 6.7e-05 & 0.25 & 0.23 \\
+1.528 -- 1.552 & 0.020509 & 6.0e-05 & 0.23 & 0.25 \\
+1.552 -- 1.579 & 0.020480 & 6.4e-05 & 0.22 & 0.24 \\
+1.579 -- 1.603 & 0.020500 & 5.6e-05 & 0.20 & 0.24 \\
+1.603 -- 1.629 & 0.020514 & 6.5e-05 & 0.19 & 0.25 \\
 \enddata
@@
 
-We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium.
+We also performed a ``free" retrieval that varied the abundances of CH$_4$, H$_2$O and NH$_3$ with no assumption of chemical equilibrium.
 The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, ruling out the the predicted methane content for a composition with solar C/O and $3\,\times$ solar metallicity (the median metallicity from the chemical equilibrium retrieval).

### 2017-08-22 Figures/format_table.py

@@ -1,6 +1,7 @@
 import numpy as np
 
 d = np.genfromtxt("w107_spectrum_06-22-17.txt")
+l = np.genfromtxt("W107_22bins.dat", skip_header = 17)
 
 n = len(d[:,0])
 bins = np.linspace(1.12, 1.65, 21)
@@ -9,4 +10,5 @@ print np.mean(d[:,1])
 for i in range(n):
 #	print '{0:0.3f}'.format(d[i,0]), '&', '{0:0.0f}'.format((d[i,1] - np.mean(d[:,1]))*1.e6), '&', '{0:0.0f}'.format(d[i,2]*1e6), '\\\\' 
 	#print '{0:0.3f}'.format(d[i,0]), '&', '{0:0.6f}'.format(d[i,1]), '&', '{0:0.1e}'.format(d[i,2]), '\\\\' 
-	print '{0:0.3f}'.format(bins[i]), '--', '{0:0.3f}'.format(bins[i+1]), '&', '{0:0.6f}'.format(d[i,1]), '&', '{0:0.1e}'.format(d[i,2]), '\\\\' 
+#	print '{0:0.3f}'.format(bins[i]), '--', '{0:0.3f}'.format(bins[i+1]), '&', '{0:0.6f}'.format(d[i,1]), '&', '{0:0.1e}'.format(d[i,2]), '\\\\' 
+	print '{0:0.3f}'.format(l[i,0]), '--', '{0:0.3f}'.format(l[i+1,0]), '&', '{0:0.6f}'.format(d[i,1]), '&', '{0:0.1e}'.format(d[i,2]), '&', '{0:0.2f}'.format(l[i,3]), '&', '{0:0.2f}'.format(l[i,4]), '\\\\' 

### 2017-08-22 ApJL_submission/ms.tex
apjl version

--- 2017-08-22 ms.tex
@@
 
-\title{Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}}
+\title{Water, Methane Depletion, and High-Altitude Condensates in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}}
 
@@
 \begin{figure*}
-\includegraphics[width = \textwidth]{Figures/lc.pdf}
+\includegraphics[width = \textwidth]{fig1_lc.pdf}
 \caption{Left: Broadband and spectrophotometric transit light curves for WASP-107b compared to best fit models.
@@
 \begin{figure*}
-\includegraphics[width = \textwidth]{Figures/spectrum.pdf}
+\includegraphics[width = \textwidth]{fig2_spectrum.pdf}
 \caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals).
@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{Figures/metalByAge.pdf}
+\includegraphics[width = 0.5\textwidth]{fig3_metallicity.pdf}
 \caption{Estimated envelope metallicity for a given core mass based on interior structure modeling.
@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{Figures/retrieval.pdf}
+\includegraphics[width = 0.5\textwidth]{fig4_retrieval.pdf}
 \caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure in bars.
@@
 
-\bibliographystyle{aasjournal}
-\bibliography{ms.bib}
+\begin{thebibliography}{}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+\providecommand{\url}[1]{\href{#1}{#1}}
+
+\bibitem[{{Amundsen} {et~al.}(2016){Amundsen}, {Mayne}, {Baraffe}, {Manners},
+  {Tremblin}, {Drummond}, {Smith}, {Acreman}, \& {Homeier}}]{amundsen16}
+{Amundsen}, D.~S., {Mayne}, N.~J., {Baraffe}, I., {et~al.} 2016, \aap, 595, A36
+
+\bibitem[{{Anderson} {et~al.}(2017){Anderson}, {Collier Cameron}, {Delrez},
+  {Doyle}, {Gillon}, {Hellier}, {Jehin}, {Lendl}, {Maxted}, {Madhusudhan},
+  {Pepe}, {Pollacco}, {Queloz}, {S{\'e}gransan}, {Smalley}, {Smith}, {Triaud},
+  {Turner}, {Udry}, \& {West}}]{anderson17}
+{Anderson}, D.~R., {Collier Cameron}, A., {Delrez}, L., {et~al.} 2017, ArXiv
+  e-prints, arXiv:1701.03776
+
+\bibitem[{{Batalha} {et~al.}(2017){Batalha}, {Mandell}, {Pontoppidan},
+  {Stevenson}, {Lewis}, {Kalirai}, {Earl}, {Greene}, {Albert}, \&
+  {Nielsen}}]{batalha17}
+{Batalha}, N.~E., {Mandell}, A., {Pontoppidan}, K., {et~al.} 2017, \pasp, 129,
+  064501
+
+\bibitem[{{Benneke}(2015)}]{benneke15}
+{Benneke}, B. 2015, ArXiv e-prints, arXiv:1504.07655
+
+\bibitem[{{Brown}(2001)}]{brown01}
+{Brown}, T.~M. 2001, \apj, 553, 1006
+
+\bibitem[{{Buchner}(2016)}]{buchner16}
+{Buchner}, J. 2016, {PyMultiNest: Python interface for MultiNest}, Astrophysics
+  Source Code Library, , , ascl:1606.005
+
+\bibitem[{{Crossfield} \& {Kreidberg}(2017)}]{crossfield17}
+{Crossfield}, I.~J.~M., \& {Kreidberg}, L. 2017, ArXiv e-prints,
+  arXiv:1708.00016
+
+\bibitem[{{Dai} \& {Winn}(2017)}]{dai17}
+{Dai}, F., \& {Winn}, J.~N. 2017, \aj, 153, 205
+
+\bibitem[{{Dawson} {et~al.}(2016){Dawson}, {Lee}, \& {Chiang}}]{dawson16}
+{Dawson}, R.~I., {Lee}, E.~J., \& {Chiang}, E. 2016, \apj, 822, 54
+
+\bibitem[{{Espinoza} {et~al.}(2017){Espinoza}, {Fortney}, {Miguel},
+  {Thorngren}, \& {Murray-Clay}}]{espinoza17}
+{Espinoza}, N., {Fortney}, J.~J., {Miguel}, Y., {Thorngren}, D., \&
+  {Murray-Clay}, R. 2017, \apjl, 838, L9
+
+\bibitem[{{Espinoza} \& {Jord{\'a}n}(2015)}]{espinoza15}
+{Espinoza}, N., \& {Jord{\'a}n}, A. 2015, \mnras, 450, 1879
+
+\bibitem[{{Foreman-Mackey} {et~al.}(2013){Foreman-Mackey}, {Hogg}, {Lang}, \&
+  {Goodman}}]{foremanmackey13}
+{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2013, \pasp,
+  125, 306
+
+\bibitem[{{Fortney} {et~al.}(2008){Fortney}, {Lodders}, {Marley}, \&
+  {Freedman}}]{fortney08}
+{Fortney}, J.~J., {Lodders}, K., {Marley}, M.~S., \& {Freedman}, R.~S. 2008,
+  \apj, 678, 1419
+
+\bibitem[{{Fortney} {et~al.}(2007){Fortney}, {Marley}, \& {Barnes}}]{fortney07}
+{Fortney}, J.~J., {Marley}, M.~S., \& {Barnes}, J.~W. 2007, \apj, 659, 1661
+
+\bibitem[{{Fortney} {et~al.}(2013){Fortney}, {Mordasini}, {Nettelmann},
+  {Kempton}, {Greene}, \& {Zahnle}}]{fortney13}
+{Fortney}, J.~J., {Mordasini}, C., {Nettelmann}, N., {et~al.} 2013, \apj, 775,
+  80
+
+\bibitem[{{Freedman} {et~al.}(2014){Freedman}, {Lustig-Yaeger}, {Fortney},
+  {Lupu}, {Marley}, \& {Lodders}}]{freedman14}
+{Freedman}, R.~S., {Lustig-Yaeger}, J., {Fortney}, J.~J., {et~al.} 2014, \apjs,
+  214, 25
+
+\bibitem[{{Frelikh} \& {Murray-Clay}(2017)}]{frelikh17}
+{Frelikh}, R., \& {Murray-Clay}, R.~A. 2017, ArXiv e-prints, arXiv:1708.00862
+
+\bibitem[{{Gao} {et~al.}(2017){Gao}, {Marley}, {Zahnle}, {Robinson}, \&
+  {Lewis}}]{gao17}
+{Gao}, P., {Marley}, M.~S., {Zahnle}, K., {Robinson}, T.~D., \& {Lewis}, N.~K.
+  2017, \aj, 153, 139
+
+\bibitem[{Gordon \& McBride(1994)}]{gordon94}
+Gordon, S., \& McBride, B.~J. 1994, Computer program for calculation of complex
+  chemical equilibrium compositions and applications, Vol.~1 (National
+  Aeronautics and Space Administration, Office of Management, Scientific and
+  Technical Information Program)
+
+\bibitem[{{Guillot}(2010)}]{guillot10}
+{Guillot}, T. 2010, \aap, 520, A27
+
+\bibitem[{{Horne}(1986)}]{horne86}
+{Horne}, K. 1986, \pasp, 98, 609
+
+\bibitem[{{Kreidberg}(2015)}]{kreidberg15a}
+{Kreidberg}, L. 2015, \pasp, 127, 1161
+
+\bibitem[{{Kreidberg} {et~al.}(2014{\natexlab{a}}){Kreidberg}, {Bean},
+  {D{\'e}sert}, {Benneke}, {Deming}, {Stevenson}, {Seager}, {Berta-Thompson},
+  {Seifahrt}, \& {Homeier}}]{kreidberg14a}
+{Kreidberg}, L., {Bean}, J.~L., {D{\'e}sert}, J.-M., {et~al.}
+  2014{\natexlab{a}}, \nat, 505, 69
+
+\bibitem[{{Kreidberg} {et~al.}(2014{\natexlab{b}}){Kreidberg}, {Bean},
+  {D{\'e}sert}, {Line}, {Fortney}, {Madhusudhan}, {Stevenson}, {Showman},
+  {Charbonneau}, {McCullough}, {Seager}, {Burrows}, {Henry}, {Williamson},
+  {Kataria}, \& {Homeier}}]{kreidberg14b}
+---. 2014{\natexlab{b}}, \apjl, 793, L27
+
+\bibitem[{{Kreidberg} {et~al.}(2015){Kreidberg}, {Line}, {Bean}, {Stevenson},
+  {D{\'e}sert}, {Madhusudhan}, {Fortney}, {Barstow}, {Henry}, {Williamson}, \&
+  {Showman}}]{kreidberg15b}
+{Kreidberg}, L., {Line}, M.~R., {Bean}, J.~L., {et~al.} 2015, \apj, 814, 66
+
+\bibitem[{{Lacis} \& {Oinas}(1991)}]{lacis91}
+{Lacis}, A.~A., \& {Oinas}, V. 1991, \jgr, 96, 9027
+
+\bibitem[{{Line} {et~al.}(2014){Line}, {Knutson}, {Wolf}, \& {Yung}}]{line14}
+{Line}, M.~R., {Knutson}, H., {Wolf}, A.~S., \& {Yung}, Y.~L. 2014, \apj, 783,
+  70
+
+\bibitem[{{Line} {et~al.}(2011){Line}, {Vasisht}, {Chen}, {Angerhausen}, \&
+  {Yung}}]{line11}
+{Line}, M.~R., {Vasisht}, G., {Chen}, P., {Angerhausen}, D., \& {Yung}, Y.~L.
+  2011, \apj, 738, 32
+
+\bibitem[{{Line} {et~al.}(2013){Line}, {Wolf}, {Zhang}, {Knutson}, {Kammer},
+  {Ellison}, {Deroo}, {Crisp}, \& {Yung}}]{line13a}
+{Line}, M.~R., {Wolf}, A.~S., {Zhang}, X., {et~al.} 2013, \apj, 775, 137
+
+\bibitem[{{Line} {et~al.}(2016){Line}, {Stevenson}, {Bean}, {Desert},
+  {Fortney}, {Kreidberg}, {Madhusudhan}, {Showman}, \& {Diamond-Lowe}}]{line16}
+{Line}, M.~R., {Stevenson}, K.~B., {Bean}, J., {et~al.} 2016, \aj, 152, 203
+
+\bibitem[{{Madhusudhan} {et~al.}(2014){Madhusudhan}, {Amin}, \&
+  {Kennedy}}]{madhusudhan14}
+{Madhusudhan}, N., {Amin}, M.~A., \& {Kennedy}, G.~M. 2014, \apjl, 794, L12
+
+\bibitem[{{Madhusudhan} {et~al.}(2011){Madhusudhan}, {Harrington}, {Stevenson},
+  {Nymeyer}, {Campo}, {Wheatley}, {Deming}, {Blecic}, {Hardy}, {Lust},
+  {Anderson}, {Collier-Cameron}, {Britt}, {Bowman}, {Hebb}, {Hellier},
+  {Maxted}, {Pollacco}, \& {West}}]{madhusudhan11}
+{Madhusudhan}, N., {Harrington}, J., {Stevenson}, K.~B., {et~al.} 2011, \nat,
+  469, 64
+
+\bibitem[{{Molli{\`e}re} {et~al.}(2015){Molli{\`e}re}, {van Boekel},
+  {Dullemond}, {Henning}, \& {Mordasini}}]{molliere15}
+{Molli{\`e}re}, P., {van Boekel}, R., {Dullemond}, C., {Henning}, T., \&
+  {Mordasini}, C. 2015, \apj, 813, 47
+
+\bibitem[{{Mordasini} {et~al.}(2016){Mordasini}, {van Boekel}, {Molli{\`e}re},
+  {Henning}, \& {Benneke}}]{mordasini16}
+{Mordasini}, C., {van Boekel}, R., {Molli{\`e}re}, P., {Henning}, T., \&
+  {Benneke}, B. 2016, \apj, 832, 41
+
+\bibitem[{{Morley} {et~al.}(2012){Morley}, {Fortney}, {Marley}, {Visscher},
+  {Saumon}, \& {Leggett}}]{morley12}
+{Morley}, C.~V., {Fortney}, J.~J., {Marley}, M.~S., {et~al.} 2012, \apj, 756,
+  172
+
+\bibitem[{{Morley} {et~al.}(2015){Morley}, {Fortney}, {Marley}, {Zahnle},
+  {Line}, {Kempton}, {Lewis}, \& {Cahoy}}]{morley15}
+---. 2015, \apj, 815, 110
+
+\bibitem[{{Morley} {et~al.}(2017){Morley}, {Knutson}, {Line}, {Fortney},
+  {Thorngren}, {Marley}, {Teal}, \& {Lupu}}]{morley17}
+{Morley}, C.~V., {Knutson}, H., {Line}, M., {et~al.} 2017, \aj, 153, 86
+
+\bibitem[{{Moses} {et~al.}(2013){Moses}, {Madhusudhan}, {Visscher}, \&
+  {Freedman}}]{moses13}
+{Moses}, J.~I., {Madhusudhan}, N., {Visscher}, C., \& {Freedman}, R.~S. 2013,
+  \apj, 763, 25
+
+\bibitem[{{Mo{\v c}nik} {et~al.}(2017){Mo{\v c}nik}, {Hellier}, {Anderson},
+  {Clark}, \& {Southworth}}]{mocnik17}
+{Mo{\v c}nik}, T., {Hellier}, C., {Anderson}, D.~R., {Clark}, B.~J.~M., \&
+  {Southworth}, J. 2017, \mnras, 469, 1622
+
+\bibitem[{{{\"O}berg} {et~al.}(2011){{\"O}berg}, {Murray-Clay}, \&
+  {Bergin}}]{oberg11}
+{{\"O}berg}, K.~I., {Murray-Clay}, R., \& {Bergin}, E.~A. 2011, \apjl, 743, L16
+
+\bibitem[{{Pollack} {et~al.}(1996){Pollack}, {Hubickyj}, {Bodenheimer},
+  {Lissauer}, {Podolak}, \& {Greenzweig}}]{pollack96}
+{Pollack}, J.~B., {Hubickyj}, O., {Bodenheimer}, P., {et~al.} 1996, \icarus,
+  124, 62
+
+\bibitem[{{Saumon} {et~al.}(1995){Saumon}, {Chabrier}, \& {van
+  Horn}}]{saumon95}
+{Saumon}, D., {Chabrier}, G., \& {van Horn}, H.~M. 1995, \apjs, 99, 713
+
+\bibitem[{{Skemer} {et~al.}(2014){Skemer}, {Marley}, {Hinz}, {Morzinski},
+  {Skrutskie}, {Leisenring}, {Close}, {Saumon}, {Bailey}, {Briguglio},
+  {Defrere}, {Esposito}, {Follette}, {Hill}, {Males}, {Puglisi}, {Rodigas}, \&
+  {Xompero}}]{skemer14}
+{Skemer}, A.~J., {Marley}, M.~S., {Hinz}, P.~M., {et~al.} 2014, \apj, 792, 17
+
+\bibitem[{{Stevenson} {et~al.}(2014){Stevenson}, {Bean}, {Seifahrt},
+  {D{\'e}sert}, {Madhusudhan}, {Bergmann}, {Kreidberg}, \&
+  {Homeier}}]{stevenson14c}
+{Stevenson}, K.~B., {Bean}, J.~L., {Seifahrt}, A., {et~al.} 2014, \aj, 147, 161
+
+\bibitem[{{Thorngren} {et~al.}(2016){Thorngren}, {Fortney}, {Murray-Clay}, \&
+  {Lopez}}]{thorngren16}
+{Thorngren}, D.~P., {Fortney}, J.~J., {Murray-Clay}, R.~A., \& {Lopez}, E.~D.
+  2016, \apj, 831, 64
+
+\bibitem[{{Tinetti} {et~al.}(2012){Tinetti}, {Tennyson}, {Griffith}, \&
+  {Waldmann}}]{tinetti12}
+{Tinetti}, G., {Tennyson}, J., {Griffith}, C.~A., \& {Waldmann}, I. 2012,
+  Philosophical Transactions of the Royal Society of London Series A, 370, 2749
+
+\bibitem[{{Wakeford} {et~al.}(2017){Wakeford}, {Visscher}, {Lewis}, {Kataria},
+  {Marley}, {Fortney}, \& {Mandell}}]{wakeford17}
+{Wakeford}, H.~R., {Visscher}, C., {Lewis}, N.~K., {et~al.} 2017, \mnras, 464,
+  4247
+
+\bibitem[{{Zahnle} {et~al.}(2009){Zahnle}, {Marley}, {Freedman}, {Lodders}, \&
+  {Fortney}}]{zahnle09}
+{Zahnle}, K., {Marley}, M.~S., {Freedman}, R.~S., {Lodders}, K., \& {Fortney},
+  J.~J. 2009, \apjl, 701, L20
+
+\bibitem[{{Zahnle} \& {Marley}(2014)}]{zahnle14}
+{Zahnle}, K.~J., \& {Marley}, M.~S. 2014, \apj, 797, 41
+
+\bibitem[{{Zhou} {et~al.}(2017){Zhou}, {Apai}, {Lew}, \& {Schneider}}]{zhou17}
+{Zhou}, Y., {Apai}, D., {Lew}, B.~W.~P., \& {Schneider}, G. 2017, \aj, 153, 243
+
+\end{thebibliography}
 

### 2017-09-25 arxiv_v2/ms.tex
arxiv

--- 2017-08-22 ApJL_submission/ms.tex
@@
-\documentclass[twocolumn]{aastex61}
+\documentclass[twocolumn, trackchanges]{aastex61}
 
@@
 
-\title{Water, Methane Depletion, and High-Altitude Condensates in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}}
+\title{Water, Methane Depletion, and High-Altitude Condensates in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}}
 
@@
 \begin{abstract}
-The Neptune-mass exoplanet WASP-107b is an exciting target for atmosphere characterization.
+The super-Neptune exoplanet WASP-107b is an exciting target for atmosphere characterization.
 It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.
@@
 We checked that the values we fixed were consistent with our estimates when we allowed them to vary freely.
-We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.
+We also checked that the uncertainty in the stellar parameters does not affect the PHOENIX model predictions at the level of precision of our data: we varied the stellar effective temperature by $\pm100$ K from the published value and found that the transmission spectrum was not significantly changed.
 The remaining free parameters in the fit were $t_0$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
@@
 The amplitude of the feature has no detectable wavelength dependence at the level of precision of our data.
-As for the white light curve, we fixed some of the transit parameters on the estimates from \cite{dai17} and fixed the limb darkening on the PHOENIX model (coefficients listed in Table\,\ref{tab:transit_depths}.
+As for the white light curve, we fixed some of the transit parameters on the best fit values from \cite{dai17}.
+We did not put put priors on the transit parameters because varying them simply shifts the spectrum up or down by a constant value.
+Since the planet's 10-bar radius is a free parameter in our analysis of the spectrum, the absolute transit depth does not affect the retrieved atmospheric properties.
+In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions (coefficients listed in Table\,\ref{tab:transit_depths}.
+
 The final spectroscopic light curve fits had just three free parameters per channel: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
@@
 We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum.
-For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths.
-%(differing by just $0.3\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
-This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval (see \S\,\ref{sec:retrieval}).
+In one test, we fit the spectroscopic light curves with the same analytic model we used for the white light curve.
+We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
+The results from these tests are nearly identical to our nominal transmission spectrum (differing by $<0.5\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
+This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval with the 10-bar radius parameter (see \S\,\ref{sec:retrieval}).
+
 We also fit for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values.
@@
 For the older age, the upper limit is even lower ($20\times$ solar), because planets cool and contract as they age \citep{fortney08}.
-These are conservative upper limits, but realistically, the planet probably formed with a core.
+The largest sources of uncertainty in the metallicity estimate are the unknown core mass and stellar age, which are dominant over modeling uncertainties due to the equation of state and distribution of heavy elements in the envelope \citep{thorngren16}.
+By marginalizing over the unknown physical parameters, we obtain a conservative upper limit on the envelope metallicity.
+Realistically, the planet probably formed with a core.
 Assuming a $5\,M_\oplus$ core, the upper limit on metallicity is 20 (10) $M_\oplus$ for the young (old) age.
@@
 We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.
-We also varied the assumed planet mass by $3\,sigma$.
+We also varied the assumed planet mass by $3\,\sigma$.
 These models did not significantly improve the fit quality, and the main retrieval results were unchanged.
@@
 We also performed a ``free" retrieval that varied the abundances of CH$_4$, H$_2$O and NH$_3$ with no assumption of chemical equilibrium.
-The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, ruling out the the predicted methane content for a composition with solar C/O and $3\,\times$ solar metallicity (the median metallicity from the chemical equilibrium retrieval).
+The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, ruling out the the predicted equilibrium methane content at photospheric temperature and pressure (700 Kelvin, 10 mbar), for a composition with solar C/O and $3\,\times$ solar metallicity.
+We chose $3\times$ solar as a benchmark for comparison because this is the median metallicity from the chemical equilibrium retrieval.
 Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.
@@
 These results are a first look at the atmosphere of WASP-107b.
-The planet is already being observed at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (\emph{HST} Program GO 14916, PI J.
+It joins the ranks of two other relatively cool, relatively low mass exoplanets with detected water features \citep[HAT-P-11b and HAT-P-26b;][]{fraine14, wakeford17}.
+WASP-107b is already the target of additional observing programs at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (\emph{HST} Program GO 14916, PI J.
 Spake, \emph{Spitzer} Program 13052, PI M.
@@
   80
+
+\bibitem[{{Fraine} {et~al.}(2014){Fraine}, {Deming}, {Benneke}, {Knutson},
+  {Jord{\'a}n}, {Espinoza}, {Madhusudhan}, {Wilkins}, \& {Todorov}}]{fraine14}
+{Fraine}, J., {Deming}, D., {Benneke}, B., {et~al.} 2014, \nat, 513, 526
 

### 2017-09-25 ms.tex
arxiv

@@
 
-\title{Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}}
+\title{Water, Methane Depletion, and High-Altitude Condensates in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}}
 
@@
 \begin{figure*}
-\includegraphics[width = \textwidth]{Figures/lc.pdf}
+\includegraphics[width = \textwidth]{Figures/fig1_lc.pdf}
 \caption{Left: Broadband and spectrophotometric transit light curves for WASP-107b compared to best fit models.
@@
 \begin{figure*}
-\includegraphics[width = \textwidth]{Figures/spectrum.pdf}
+\includegraphics[width = \textwidth]{Figures/fig2_spectrum.pdf}
 \caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals).
@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{Figures/metalByAge.pdf}
+\includegraphics[width = 0.5\textwidth]{Figures/fig3_metallicity.pdf}
 \caption{Estimated envelope metallicity for a given core mass based on interior structure modeling.
@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{Figures/retrieval.pdf}
+\includegraphics[width = 0.5\textwidth]{Figures/fig4_retrieval.pdf}
 \caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure in bars.

### 2017-09-25 ms_revised.tex
arxiv

--- 2017-09-25 arxiv_v2/ms.tex
@@
 \begin{figure*}
-\includegraphics[width = \textwidth]{fig1_lc.pdf}
+\includegraphics[width = \textwidth]{Figures/fig1_lc.pdf}
 \caption{Left: Broadband and spectrophotometric transit light curves for WASP-107b compared to best fit models.
@@
 We checked that the values we fixed were consistent with our estimates when we allowed them to vary freely.
-We also checked that the uncertainty in the stellar parameters does not affect the PHOENIX model predictions at the level of precision of our data: we varied the stellar effective temperature by $\pm100$ K from the published value and found that the transmission spectrum was not significantly changed.
+We also checked that the uncertainty in the stellar parameters does not affect the PHOENIX model predictions at the level of precision of our data: \added{we varied the stellar effective temperature by $\pm100$ K from the published value and found that the transmission spectrum was not significantly changed}.
 The remaining free parameters in the fit were $t_0$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
@@
 The amplitude of the feature has no detectable wavelength dependence at the level of precision of our data.
-As for the white light curve, we fixed some of the transit parameters on the best fit values from \cite{dai17}.
-We did not put put priors on the transit parameters because varying them simply shifts the spectrum up or down by a constant value.
-Since the planet's 10-bar radius is a free parameter in our analysis of the spectrum, the absolute transit depth does not affect the retrieved atmospheric properties.
-In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions (coefficients listed in Table\,\ref{tab:transit_depths}.
+As for the white light curve, we fixed some of the transit parameters on the \replaced{estimates}{best fit values} from \cite{dai17}.
+\added{We did not put put priors on the transit parameters because varying them simply shifts the spectrum up or down by a constant value.
+Since the planet's 10-bar radius is a free parameter in our analysis of the spectrum, the absolute transit depth does not affect the retrieved atmospheric properties.} In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions (coefficients listed in Table\,\ref{tab:transit_depths}.
 
@@
 In one test, we fit the spectroscopic light curves with the same analytic model we used for the white light curve.
-We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
+\added{We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
 The results from these tests are nearly identical to our nominal transmission spectrum (differing by $<0.5\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
-This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval with the 10-bar radius parameter (see \S\,\ref{sec:retrieval}).
+This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval with the 10-bar radius parameter (see \S\,\ref{sec:retrieval}).}
 
@@
 \begin{figure*}
-\includegraphics[width = \textwidth]{fig2_spectrum.pdf}
+\includegraphics[width = \textwidth]{Figures/fig2_spectrum.pdf}
 \caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals).
@@
 For the older age, the upper limit is even lower ($20\times$ solar), because planets cool and contract as they age \citep{fortney08}.
-The largest sources of uncertainty in the metallicity estimate are the unknown core mass and stellar age, which are dominant over modeling uncertainties due to the equation of state and distribution of heavy elements in the envelope \citep{thorngren16}.
-By marginalizing over the unknown physical parameters, we obtain a conservative upper limit on the envelope metallicity.
-Realistically, the planet probably formed with a core.
+\added{The largest sources of uncertainty in the metallicity estimate are the unknown core mass and stellar age, which are dominant over modeling uncertainties due to the equation of state and distribution of heavy elements in the envelope \citep{thorngren16}.} By marginalizing over the unknown physical parameters, we obtain a conservative upper limit on the envelope metallicity.
+\deleted{These are conservative upper limits, but} Realistically, the planet probably formed with a core.
 Assuming a $5\,M_\oplus$ core, the upper limit on metallicity is 20 (10) $M_\oplus$ for the young (old) age.
@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{fig3_metallicity.pdf}
+\includegraphics[width = 0.5\textwidth]{Figures/fig3_metallicity.pdf}
 \caption{Estimated envelope metallicity for a given core mass based on interior structure modeling.
@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{fig4_retrieval.pdf}
+\includegraphics[width = 0.5\textwidth]{Figures/fig4_retrieval.pdf}
 \caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure in bars.
@@
 We also performed a ``free" retrieval that varied the abundances of CH$_4$, H$_2$O and NH$_3$ with no assumption of chemical equilibrium.
-The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, ruling out the the predicted equilibrium methane content at photospheric temperature and pressure (700 Kelvin, 10 mbar), for a composition with solar C/O and $3\,\times$ solar metallicity.
-We chose $3\times$ solar as a benchmark for comparison because this is the median metallicity from the chemical equilibrium retrieval.
+The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, ruling out the the predicted \added{equilibrium} methane content \added{at photospheric temperature and pressure (700 Kelvin, 10 mbar)}, for a composition with solar C/O and $3\,\times$ solar metallicity.
+\added{We chose $3\times$ solar as a benchmark for comparison because} this is the median metallicity from the chemical equilibrium retrieval.
 Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.
@@
 These results are a first look at the atmosphere of WASP-107b.
-It joins the ranks of two other relatively cool, relatively low mass exoplanets with detected water features \citep[HAT-P-11b and HAT-P-26b;][]{fraine14, wakeford17}.
-WASP-107b is already the target of additional observing programs at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (\emph{HST} Program GO 14916, PI J.
+\added{It joins the ranks of two other relatively cool, relatively low mass exoplanets with detected water features \citep[HAT-P-11b and HAT-P-26b;][]{fraine14, wakeford17}.} WASP-107b is already the target of additional observing programs at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (\emph{HST} Program GO 14916, PI J.
 Spake, \emph{Spitzer} Program 13052, PI M.

### 2018-02-21 Figures/fig2_spectrum.py
response

@@ -21,7 +21,10 @@ plt.errorbar(d[:,0], d[:,1]*100., d[:,2]*100., linestyle = 'none', zorder = 100,
 #plt.plot(d[:,0]*1e6, convolve(d[:,1], g, boundary = 'extend')-0.08)
 
 p = pickle.load(open("WASP107b_Global_Cloud_limit_Metallicity_spectra.pic", "rb"))
-#print "outliers", (p[9] - p[7])/p[8]
+print "outliers", (p[9] - p[7])/p[8]
+print "chi2, n", np.sum((p[9] - p[7])**2/p[8]**2), len(p[9])
+
+
 w = p[0]
 
 plt.fill_between(w, p[1], p[5], color = "navy", alpha = '0.2')
@@ -31,7 +34,7 @@ plt.fill_between(w, p[2], p[4], color = "navy", alpha = '0.3')
 # best fit
 pb = pickle.load(open("Model_best_fit.pic", "rb"))
 plt.plot(pb[0], pb[1]*100., color = 'navy', label = 'Best fit model')
-print 'outliers', (p[7] - pb[3]*100.)/p[8]
+#print 'outliers', (p[7] - pb[3]*100.)/p[8]
 
 """d = pickle.load(open("Model_Solar.pic", "rb"))
 x, y = d[0], d[1]*100. - 0.7
@@ -52,5 +55,6 @@ plt.xlim(1.05, 1.7)
 
 plt.tight_layout()
 plt.savefig("spectrum.pdf")
+plt.savefig("spectrum.png")
 plt.show()
 

### 2018-02-21 ApJL_submission/ms_v1.tex
ref response

--- 2017-09-25 ms_revised.tex
@@
 \begin{figure*}
-\includegraphics[width = \textwidth]{Figures/fig1_lc.pdf}
+\includegraphics[width = \textwidth]{fig1_lc.pdf}
 \caption{Left: Broadband and spectrophotometric transit light curves for WASP-107b compared to best fit models.
@@
 In one test, we fit the spectroscopic light curves with the same analytic model we used for the white light curve.
-\added{We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
+\added{We also tested a white light curve systematics vector $S_white$ from a fit that included the star-spot crossing orbit.
 The results from these tests are nearly identical to our nominal transmission spectrum (differing by $<0.5\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
@@
 \begin{figure*}
-\includegraphics[width = \textwidth]{Figures/fig2_spectrum.pdf}
+\includegraphics[width = \textwidth]{fig2_spectrum.pdf}
 \caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals).
@@
 For the older age, the upper limit is even lower ($20\times$ solar), because planets cool and contract as they age \citep{fortney08}.
-\added{The largest sources of uncertainty in the metallicity estimate are the unknown core mass and stellar age, which are dominant over modeling uncertainties due to the equation of state and distribution of heavy elements in the envelope \citep{thorngren16}.} By marginalizing over the unknown physical parameters, we obtain a conservative upper limit on the envelope metallicity.
+\added{The largest sources of uncertainty in the metallicity estimate are the unknown core mass and stellar age, which are dominant over modeling uncertainties due to the equation of state and distribution of heavy elements in the envelope \citep{thorngren16}.
+By marginalizing over the unknown physical parameters, we obtain a conservative upper limit on the envelope metallicity.
 \deleted{These are conservative upper limits, but} Realistically, the planet probably formed with a core.
@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{Figures/fig3_metallicity.pdf}
+\includegraphics[width = 0.5\textwidth]{fig3_metallicity.pdf}
 \caption{Estimated envelope metallicity for a given core mass based on interior structure modeling.
@@
 \begin{figure}
-\includegraphics[width = 0.5\textwidth]{Figures/fig4_retrieval.pdf}
+\includegraphics[width = 0.5\textwidth]{fig4_retrieval.pdf}
 \caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure in bars.
@@
 These results are a first look at the atmosphere of WASP-107b.
-\added{It joins the ranks of two other relatively cool, relatively low mass exoplanets with detected water features \citep[HAT-P-11b and HAT-P-26b;][]{fraine14, wakeford17}.} WASP-107b is already the target of additional observing programs at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (\emph{HST} Program GO 14916, PI J.
+The planet is already being observed at other wavelengths, including the WFC3/G102 grism and Spitzer 3.6 and 4.5 $\mu$m channels in transit and eclipse (\emph{HST} Program GO 14916, PI J.
 Spake, \emph{Spitzer} Program 13052, PI M.
@@
   80
-
-\bibitem[{{Fraine} {et~al.}(2014){Fraine}, {Deming}, {Benneke}, {Knutson},
-  {Jord{\'a}n}, {Espinoza}, {Madhusudhan}, {Wilkins}, \& {Todorov}}]{fraine14}
-{Fraine}, J., {Deming}, D., {Benneke}, B., {et~al.} 2014, \nat, 513, 526
 

### 2018-02-22 ApJL_submission/ms.tex
response2

@@
 This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$).
+\added{We infer from this test that the absence of methane is what drives the low retrieved C/O ratio.}
 
 We also performed a ``free" retrieval that varied the abundances of CH$_4$, H$_2$O and NH$_3$ with no assumption of chemical equilibrium.
-The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, ruling out the the predicted methane content for a composition with solar C/O and $3\,\times$ solar metallicity (the median metallicity from the chemical equilibrium retrieval).
-Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.
+The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, \added{ruling out the predicted equilibrium methane content at the base of the cloud deck ($1.4\time10^{-3}$ at 700 K, 10 mbar, for a solar C/O and $3\,\times$ solar metallicity composition).}  We chose $3\times$ solar as a benchmark for comparison because this is the median metallicity from the chemical equilibrium retrieval; \added{however, we note that the expected equilibrium abundance is sensitive to the true atmosphere metallicity.
+For a $1\times$ solar composition, the expected methane abundance drops by $1/3$.
+Similarly, the expected abundances are sensitive to the temperature and pressure in the photosphere.
+For example, at slightly higher temperatures and higher altitudes (1 mbar, 800 K), the expected methane abundance drops to $2.4\times10^{-5}$.}  Based on these tests, we conclude that the atmosphere of WASP-107b is \added{likely} depleted in methane relative to expectations for a solar abundance pattern.
 By contrast, the water abundance ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions for solar composition.
+FIXME
 

### 2018-02-24 ApJL_submission/ms.tex
edits

@@
 
-\title{Water, Methane Depletion, and High-Altitude Condensates in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}}
+\title{Water, High-Altitude Condensates, and \added{Possible} Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}}
 
@@
 The interior structure models set a $3\,\sigma$ upper limit on the atmospheric metallicity of $30\times$ solar.
-The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and we infer a water abundance consistent with expectations for a solar abundance pattern.
-On the other hand, methane is depleted relative to expectations (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.
-The water features are smaller than predicted for a cloudless atmosphere, crossing less than one scale height.
-A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations; however, we find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.
-Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
+The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern.
+\added{The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to methane depletion.} \deleted{and we infer a water abundance consistent with expectations for a solar abundance pattern.
+On the other hand, methane is depleted relative to expectations (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.}  The spectral features are smaller than predicted for a cloudless atmosphere, crossing less than one scale height.
+A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations.
+\added{We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) produce clouds at this pressure level.} \deleted{however, we find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.}  Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
 The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities.
@@
 WASP-107b also has a relatively low equilibrium temperature compared to most other exoplanets that are amenable to atmosphere characterization (780 K, assuming zero albedo).
-This results in a distinct atmospheric chemistry from other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.
-Spectral features from both water and methane are accessible with current observing facilities, enabling a spectroscopic estimate of the carbon-to-oxygen ratio (C/O).
-Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.
+This results in a distinct atmospheric chemistry compared to other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.
+Spectral features from both water and methane are accessible with current observing facilities -- \added{both of these molecules have strong absorption bands in the wavelength range covered by \emph{HST}/WFC3.
+Measuring the shape and amplitude of these spectral features enables a constraint on the abundance of water, the dominant oxygen-bearing molecule, and methane, the dominant carbon-bearing molecule, providing a bounded estimate of the carbon-to-oxygen ratio (C/O).} Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage, \added{and have mainly resulted in upper limits} \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.
 
@@
 Since the planet's 10-bar radius is a free parameter in our analysis of the spectrum, the absolute transit depth does not affect the retrieved atmospheric properties.
-In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions (coefficients listed in Table\,\ref{tab:transit_depths}.
-
+In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions (coefficients listed in Table\,\ref{tab:tspec}.
 The final spectroscopic light curve fits had just three free parameters per channel: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
+We refer to this fitting approach as ``Method A".
 
@@
 For each light curve we ran a fit with 50 walkers and 1000 steps per walker, and tested for convergence as we did for the white light curve.
-The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:transit_depths}.
+The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:tspec}.
 
@@
 In one test, we fit the spectroscopic light curves with the same analytic model we used for the white light curve.
-We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
-The results from these tests are nearly identical to our nominal transmission spectrum (differing by $<0.5\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature.
+\added{This fit (which we label ``Method B") has additional free parameters, so the transit depth uncertainties increase by 50\%, but the best fit transit depths are consistent to within $0.5\sigma$ on average over the bandpass.} We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
+The results from these tests are nearly identical to our nominal transmission spectrum, except with a small constant offset due to the uncorrected star-spot crossing feature.
 This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval with the 10-bar radius parameter (see \S\,\ref{sec:retrieval}).
+	\added{We also repeated the calculations from Method A, but fit} for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values.
+\added{This approach, which we dub ``Method C", also yields consistent results with Method A (the relative transit depths agree to better than 0.2 sigma on average).
+The transit depths and fitted limb darkening coefficients are provide in Table\,\ref{tab:tspec}.} The fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths.
+\added{We calculated model limb darkening coefficients for WASP-107 over the $1\,\sigma$ range of stellar effective temperatures, surface gravities, and metallicities published in the discovery paper ($T_\mathrm{eff} = 4430\pm130$ K, $\log g = 4.5 \pm 0.1$, and $[\mathrm{Fe/H}] = 0.02 \pm 0.1i$).
+The uncertainty in $T_\mathrm{eff}$ has the largest effect on the estimated limb darkening coefficients.
+We recalculated the transmission spectrum with limb darkening models that varied $T_\mathrm{eff}$ from 4310 to 4550 K.
+Over this temperature range, the change in limb darkening shifts the average transit depth by less than 20 ppm, which is below the level of noise in the data.
+The relative change in transit depth is smaller still ($< 1$ ppm).}
 
-We also fit for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values.
-We found that the fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths by about 10\%.
-We also performed an independent data reduction and fit using K.
-Stevenson's pipeline and found consistent results (well within $1\,\sigma$).
+In addition to these tests, we also performed an independent data reduction and fit using K. Stevenson's pipeline and found consistent results (well within $1\,\sigma$).
+
+\added{\subsubsection{Impact of Unocculted Star Spots}
+	Unocculted star spots may influence the measured transmission spectrum \citep{mccullough14, zellem17}.
+To estimate the impact of unocculted spots on our data,  we first estimated the star spot properties based on the observed photometric variability in the Kepler band pass, which is 0.3\% \citep{dai17,mocnik17}.
+This variability can be reproduced with a single spot with covering fraction of 1\% and temperature 300 K lower than the stellar photosphere (4130 vs 4430 K).
+We calculated the effect of unocculted spots with these properties using equation (1) from \cite{mccullough14}, and predict that they will introduce a positive slope in the transmission spectrum of 8 ppm over the WFC3 bandpass, well below our measurement uncertainties.
+We note that cool stars may have persistent spot coverage that does not produce any photometric variability \citep{rackham18}.
+However, even if there is persistent spot coverage of 5\%, the spots produce a slope in the WFC3 bandpass of 50 ppm, which is still below our uncertainties.}
 
@@
 
-\begin{deluxetable}{c c c c c}
-\tablecaption{WASP-107b transmission spectrum and limb darkening coefficients\label{tab:transit_depths}}
-\tablehead{\colhead{Wavelength ($\mu$m)}  & \colhead{Transit depth} & \colhead{Uncertainty} & \colhead{$u_1$} & \colhead{$u_2$}}
+\begin{deluxetable*}{LLLLLLL}
+	\tablecolumns{7}
+	\tablewidth{0pt}
+	\tablecaption{WASP-107b Transmission Spectrum and Limb Darkening Coefficients\label{tab:tspec}}
+	\tablehead{
+		\colhead{Wavelength} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{$u_1$} & \colhead{$u_2$} & \colhead{$u$} \\
+	\colhead{(micron)} & \colhead{(Method A)\tablenotemark{*}} & \colhead{(Method B)} & \colhead{(Method C)} & \colhead{(PHOENIX)} & \colhead{(PHOENIX)} & \colhead{(fitted)}}
 		\startdata
-1.121 -- 1.145 & 0.020641 & 5.9e-05 & 0.38 & 0.13 \\
-1.145 -- 1.171 & 0.020733 & 5.5e-05 & 0.37 & 0.14 \\
-1.171 -- 1.196 & 0.020505 & 5.6e-05 & 0.36 & 0.14 \\
-1.196 -- 1.222 & 0.020498 & 5.4e-05 & 0.36 & 0.15 \\
-1.222 -- 1.248 & 0.020455 & 5.9e-05 & 0.36 & 0.15 \\
-1.248 -- 1.272 & 0.020492 & 5.0e-05 & 0.35 & 0.16 \\
-1.272 -- 1.298 & 0.020620 & 6.2e-05 & 0.34 & 0.17 \\
-1.298 -- 1.323 & 0.020739 & 5.0e-05 & 0.34 & 0.17 \\
-1.323 -- 1.349 & 0.020660 & 5.7e-05 & 0.33 & 0.18 \\
-1.349 -- 1.374 & 0.020858 & 4.8e-05 & 0.32 & 0.19 \\
-1.374 -- 1.401 & 0.020794 & 4.8e-05 & 0.31 & 0.20 \\
-1.401 -- 1.425 & 0.020888 & 5.2e-05 & 0.30 & 0.21 \\
-1.425 -- 1.452 & 0.020821 & 6.2e-05 & 0.29 & 0.21 \\
-1.452 -- 1.476 & 0.020691 & 5.1e-05 & 0.28 & 0.22 \\
-1.476 -- 1.502 & 0.020682 & 6.9e-05 & 0.26 & 0.23 \\
-1.502 -- 1.528 & 0.020679 & 6.7e-05 & 0.25 & 0.23 \\
-1.528 -- 1.552 & 0.020509 & 6.0e-05 & 0.23 & 0.25 \\
-1.552 -- 1.579 & 0.020480 & 6.4e-05 & 0.22 & 0.24 \\
-1.579 -- 1.603 & 0.020500 & 5.6e-05 & 0.20 & 0.24 \\
-1.603 -- 1.629 & 0.020514 & 6.5e-05 & 0.19 & 0.25 \\
+		1.121 - 1.145 & 0.020641 \pm 5.9e-05 & 0.020674 \pm 7.8e-05 & 0.020521 \pm 7.4e-05 & 0.38 & 0.13 & 0.44 \pm 0.011 \\
+		1.145 - 1.171 & 0.020733 \pm 5.5e-05 & 0.020825 \pm 7.5e-05 & 0.020575 \pm 7.7e-05 & 0.37 & 0.14 & 0.41 \pm 0.012 \\
+		1.171 - 1.196 & 0.020505 \pm 5.6e-05 & 0.020655 \pm 7.4e-05 & 0.020388 \pm 8.2e-05 & 0.36 & 0.14 & 0.4 \pm 0.012 \\
+		1.196 - 1.222 & 0.020498 \pm 5.4e-05 & 0.020580 \pm 7.4e-05 & 0.020345 \pm 6.6e-05 & 0.36 & 0.15 & 0.41 \pm 0.012 \\
+		1.222 - 1.248 & 0.020455 \pm 5.9e-05 & 0.020620 \pm 7.8e-05 & 0.020327 \pm 7.7e-05 & 0.36 & 0.15 & 0.39 \pm 0.012 \\
+		1.248 - 1.272 & 0.020492 \pm 5.0e-05 & 0.020649 \pm 6.3e-05 & 0.020360 \pm 7.6e-05 & 0.35 & 0.16 & 0.39 \pm 0.01 \\
+		1.272 - 1.298 & 0.020620 \pm 6.2e-05 & 0.020702 \pm 9.0e-05 & 0.020475 \pm 8.3e-05 & 0.34 & 0.17 & 0.42 \pm 0.014 \\
+		1.298 - 1.323 & 0.020739 \pm 5.0e-05 & 0.020883 \pm 6.7e-05 & 0.020602 \pm 8.6e-05 & 0.34 & 0.17 & 0.39 \pm 0.011 \\
+		1.323 - 1.349 & 0.020660 \pm 5.7e-05 & 0.020789 \pm 8.7e-05 & 0.020489 \pm 8.3e-05 & 0.33 & 0.18 & 0.4 \pm 0.014 \\
+		1.349 - 1.374 & 0.020858 \pm 4.8e-05 & 0.020916 \pm 7.6e-05 & 0.020706 \pm 7.9e-05 & 0.32 & 0.19 & 0.42 \pm 0.012 \\
+		1.374 - 1.401 & 0.020794 \pm 4.8e-05 & 0.020864 \pm 8.1e-05 & 0.020653 \pm 7.9e-05 & 0.31 & 0.20 & 0.41 \pm 0.012 \\
+		1.401 - 1.425 & 0.020888 \pm 5.2e-05 & 0.020961 \pm 6.1e-05 & 0.020776 \pm 7.0e-05 & 0.30 & 0.21 & 0.41 \pm 0.01 \\
+		1.425 - 1.452 & 0.020821 \pm 6.2e-05 & 0.020933 \pm 7.8e-05 & 0.020708 \pm 8.5e-05 & 0.29 & 0.21 & 0.39 \pm 0.011 \\
+		1.452 - 1.476 & 0.020691 \pm 5.1e-05 & 0.020876 \pm 7.8e-05 & 0.020556 \pm 8.7e-05 & 0.28 & 0.22 & 0.36 \pm 0.012 \\
+		1.476 - 1.502 & 0.020682 \pm 6.9e-05 & 0.020890 \pm 8.7e-05 & 0.020549 \pm 1.0e-04 & 0.26 & 0.23 & 0.34 \pm 0.014 \\
+		1.502 - 1.528 & 0.020679 \pm 6.7e-05 & 0.020763 \pm 9.6e-05 & 0.020523 \pm 8.8e-05 & 0.25 & 0.23 & 0.37 \pm 0.015 \\
+		1.528 - 1.552 & 0.020509 \pm 6.0e-05 & 0.020667 \pm 9.3e-05 & 0.020359 \pm 1.0e-04 & 0.23 & 0.25 & 0.33 \pm 0.014 \\
+		1.552 - 1.579 & 0.020480 \pm 6.4e-05 & 0.020551 \pm 1.2e-04 & 0.020376 \pm 1.3e-04 & 0.22 & 0.24 & 0.38 \pm 0.019 \\
+		1.579 - 1.603 & 0.020500 \pm 5.6e-05 & 0.020542 \pm 1.1e-04 & 0.020321 \pm 1.1e-04 & 0.20 & 0.24 & 0.33 \pm 0.019 \\
+		1.603 - 1.629 & 0.020514 \pm 6.5e-05 & 0.020755 \pm 1.1e-04 & 0.020392 \pm 1.1e-04 & 0.19 & 0.25 & 0.29 \pm 0.019 \\
 		\enddata
-\end{deluxetable}
+		\tablenotetext{*}{Our retrieval analysis uses the transit depths from Method A.}
+		\tablecomments{comments}
+	\end{deluxetable*}
 
@@
 We also explored cases without the $30\times$ solar metallicity prior upper limit.
-Without this prior, metallicities up to $100\times$ were permitted, but other parameters were not significantly changed.
+Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57+{0.54}_{-0.31}$, and cloud top pressure $log(Pc) = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
+The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[M/H] = 1.5$).
+The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17}.}
+
 %We also tested the effect of varying the planet's surface gravity within the published FIXME $\sigma$ range, and found that our results were again minimally changed.
@@
 \subsubsection{Methane Abundance}
-We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O.
-Even though water is the dominant absorber in the WFC3 bandpass, there is enough methane opacity to change the shape of the spectral features at a detectable level for high precision data.
+\added{If methane is present in the atmosphere, it has a strong spectral feature centered at $1.4\,\mu$m.
+This feature overlaps the water feature, but has a wider shape.
+Our data are precise enough to distinguish this subtle difference in feature shape, so } to explore whether methane depletion is the underlying cause of the inferred low C/O, we ran two additional retrievals.
 In one test, we assumed chemical equilibrium but excluded methane opacity.
 This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$).
-\added{We infer from this test that the absence of methane is what drives the low retrieved C/O ratio.}
+We also performed a ``free" retrieval that varied the abundances of CH$_4$, H$_2$O and NH$_3$ with no assumption of chemical equilibrium.
+The $3\,\sigma$ upper limit on methane abundance is $1.4\times10^{-3}$.
+\added{This value is in tension with expectations for a solar abundance pattern: for $3.7\times$ solar metallicity, $650$ K, and $5\times10^{-4}$ bar (the median retrieved photosphere properties), the expected methane abundance is $1.2\times10^{-3}\times$ solar (for solar C/O).} Based on these tests, we conclude that the atmosphere of WASP-107b is \added{likely} depleted in methane relative to expectations for a solar abundance pattern.
+By contrast, the water abundance from the free retrieval ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions for solar composition.
 
-We also performed a ``free" retrieval that varied the abundances of CH$_4$, H$_2$O and NH$_3$ with no assumption of chemical equilibrium.
-The $3\,\sigma$ upper limit on methane abundance is $1.2\times10^{-3}$, \added{ruling out the predicted equilibrium methane content at the base of the cloud deck ($1.4\time10^{-3}$ at 700 K, 10 mbar, for a solar C/O and $3\,\times$ solar metallicity composition).}  We chose $3\times$ solar as a benchmark for comparison because this is the median metallicity from the chemical equilibrium retrieval; \added{however, we note that the expected equilibrium abundance is sensitive to the true atmosphere metallicity.
+%we calculate the thermochemical equilibrium methane abundance for ($1.2\time10^{-3}$ at 650 K, $5\times10^4$ bar, for a solar C/O and $3.7\,\times$ solar metallicity composition).}  We chose $3\times$ solar as a benchmark for comparison because this is the median metallicity from the chemical equilibrium retrieval; \added{however, we note that the expected equilibrium abundance is sensitive to the true atmosphere metallicity.
 For a $1\times$ solar composition, the expected methane abundance drops by $1/3$.
 Similarly, the expected abundances are sensitive to the temperature and pressure in the photosphere.
-For example, at slightly higher temperatures and higher altitudes (1 mbar, 800 K), the expected methane abundance drops to $2.4\times10^{-5}$.}  Based on these tests, we conclude that the atmosphere of WASP-107b is \added{likely} depleted in methane relative to expectations for a solar abundance pattern.
-By contrast, the water abundance ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions for solar composition.
-FIXME
+For example, at slightly higher temperatures and higher altitudes (1 mbar, 800 K), the expected methane abundance drops to $2.4\times10^{-5}$.}
 
@@
 We vary the cloud sedimentation efficiency from 3 to 0.1 (normal to highly lofted, small particulate clouds).
+\added{FIXME}.
+
 None of the models produce sufficiently low amplitude spectral features to match the observed spectrum.
@@
 Compared to results for other exoplanets of similar mass such as GJ 436b, which has a high metallicity ($>100\times$ solar) and HAT-P-26b, which is metal-poor compared to the Solar System trend, WASP-107b adds to the evidence that exoplanets exhibit a greater diversity of compositions than is present in the Solar System \citep{morley17, wakeford17}.}
-\item{\emph{The methane abundance is depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
+\item{\emph{The methane abundance is \added{likely} depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
 Another possibility is that the planet has a hot interior effective temperature ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}.
@@
 
+\bibitem[{{Fraine} {et~al.}(2014){Fraine}, {Deming}, {Benneke}, {Knutson},
+  {Jord{\'a}n}, {Espinoza}, {Madhusudhan}, {Wilkins}, \& {Todorov}}]{fraine14}
+{Fraine}, J., {Deming}, D., {Benneke}, B., {et~al.} 2014, \nat, 513, 526
+
 \bibitem[{{Freedman} {et~al.}(2014){Freedman}, {Lustig-Yaeger}, {Fortney},
@@
 
+\bibitem[{{Griffith}(2014)}]{griffith14}
+{Griffith}, C.~A. 2014, Philosophical Transactions of the Royal Society of
+  London Series A, 372, 20130086
+
 \bibitem[{{Guillot}(2010)}]{guillot10}
 {Guillot}, T. 2010, \aap, 520, A27
+
+\bibitem[{{Heng} \& {Kitzmann}(2017)}]{heng17}
+{Heng}, K., \& {Kitzmann}, D. 2017, \mnras, 470, 2972
 
@@
 
+\bibitem[{{McCullough} {et~al.}(2014){McCullough}, {Crouzet}, {Deming}, \&
+  {Madhusudhan}}]{mccullough14}
+{McCullough}, P.~R., {Crouzet}, N., {Deming}, D., \& {Madhusudhan}, N. 2014,
+  \apj, 791, 55
+
 \bibitem[{{Molli{\`e}re} {et~al.}(2015){Molli{\`e}re}, {van Boekel},
@@
 
+\bibitem[{{Rackham} {et~al.}(2018){Rackham}, {Apai}, \& {Giampapa}}]{rackham18}
+{Rackham}, B.~V., {Apai}, D., \& {Giampapa}, M.~S. 2018, \apj, 853, 122
+
 \bibitem[{{Saumon} {et~al.}(1995){Saumon}, {Chabrier}, \& {van
@@
 
+\bibitem[{{Zellem} {et~al.}(2017){Zellem}, {Swain}, {Roudier}, {Shkolnik},
+  {Creech-Eakman}, {Ciardi}, {Line}, {Iyer}, {Bryden}, {Llama}, \&
+  {Fahy}}]{zellem17}
+{Zellem}, R.~T., {Swain}, M.~R., {Roudier}, G., {et~al.} 2017, \apj, 844, 27
+
 \bibitem[{{Zhou} {et~al.}(2017){Zhou}, {Apai}, {Lew}, \& {Schneider}}]{zhou17}

### 2018-02-24 Analysis/get_COratio.py

@@ -94,10 +94,15 @@ CH4 = np.log10(0.00132)
 alpha = 1.0*sum(x>CH4)/len(x)*1.0
 print "CH4 sig", get_significance(alpha/2)
 print "alpha", alpha
-q = [0.997]
+q = [0.9974]
 print quantiles(x, q)
 q = [0.16, 0.84]
+
 print "68percent conf ch4", 10**quantiles(x, q)
+
+alpha = 1.0*sum(x>np.log10(1.228e-4))/len(x)*1.0
+sig = get_significance(alpha/2)
+print "rule out best expectations, (3.7xsolar, 650K, 0.0005 bar cloud top)", sig
 
 #print "test", get_significance(0.003/2)
 

### 2018-02-24 Figures/calc_unocculted_spot_effect.py

@@ -1,0 +1,39 @@
+import numpy as np
+
+h = 6.626e-34   #J/s
+c = 3.0e8       #m/s
+kb = 1.4e-23     #J/K
+
+def blackbody(l,T): 
+    return 2*h*c**2/(l**5*(np.exp(h*c/(l*kb*T))-1))
+
+#variability amplitude is 0.3% in the Kepler bandpass (Dai & Winn 2017)
+
+l = 0.8e-6
+depth = 0.021
+
+delta = 0.01
+Tstar =4430.
+Tspot = 4130.
+
+print "Kepler variability amplitude, high contrast", (delta*blackbody(l, Tspot) + (1 - delta)*blackbody(l, Tstar))/blackbody(l, Tstar)
+
+"""delta = 0.02
+Tstar =4430.
+Tspot = 4280.
+
+print "Kepler variability amplitude, low contrast", (delta*blackbody(l, Tspot) + (1 - delta)*blackbody(l, Tstar))/blackbody(l, Tstar)"""
+
+l = np.array([1.1, 1.7])*1.e-6
+l = np.array([1.1, 1.7])*1.e-6
+
+delta *= 5
+
+bbstar = blackbody(l, Tstar)
+bbspot = blackbody(l, Tspot)
+
+print (depth - depth*(1/(1 - delta*((1 - bbspot)/bbstar))))*1.e6
+
+
+
+

### 2018-02-24 Figures/compare_ld.py

@@ -1,0 +1,30 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import os, glob
+
+d = np.genfromtxt("tspec_LDfixT4310K.txt")
+plt.errorbar(d[:,0], d[:,1], d[:,2])
+e = np.genfromtxt("tspec_LDfixT4430K.txt")
+plt.errorbar(e[:,0], e[:,1], e[:,2])
+f = np.genfromtxt("tspec_LDfixT4550K.txt")
+plt.errorbar(f[:,0], f[:,1], f[:,2])
+
+print np.max((d[:,1] - np.mean(d[:,1]))**2 - (f[:,1] -np.mean(f[:,1]))**2)*1e6
+
+plt.show()
+
+path = "LD_coefficients"
+files = glob.glob(os.path.join(path, "*141*"))
+
+colors = ['red', 'blue', 'orange', 'yellow', 'green', 'purple']
+
+for i, f in enumerate(files):
+    d = np.genfromtxt(f)
+    print f
+    plt.plot(d[:,0], d[:,3], color = colors[i])
+    #plt.plot(d[:,0], d[:,4], color = colors[i])
+
+plt.legend()
+plt.show()
+
+    

### 2018-02-24 Figures/format_table.py

@@ -1,14 +1,69 @@
 import numpy as np
+import matplotlib.pyplot as plt
 
 d = np.genfromtxt("w107_spectrum_06-22-17.txt")
+d2 = np.genfromtxt("w107_spectrum_02-22-18-dw-freeld.txt")
+d3 = np.genfromtxt("w107_spectrum-02-22-18-analytic.txt")
 l = np.genfromtxt("W107_22bins.dat", skip_header = 17)
 
+ld = np.genfromtxt("w107_freeld_02-22-18.txt")
+
+print "increase in errors for mr", np.median(d3[:,2]/d[:,2])
+print "increase in errors for freeld", np.median(d2[:,2]/d[:,2])
+
+
 n = len(d[:,0])
-bins = np.linspace(1.12, 1.65, 21)
 
 print np.mean(d[:,1])
+
+
+#wavelength, transit depth 1, transit depth 2, transit depth 3, PHOENIX u1, PHOENIX u2, fitted LD
+
+print "\\begin{deluxetable*}{LLLLLLL}"
+print "\\tablecolumns{7}"
+print "\\tablewidth{0pt}"
+print "\\tablecaption{WASP-107b Transmission Spectrum and Limb Darkening Coefficients\label{table:tspec}}"
+print "\\tablehead{"
+print "\colhead{Wavelength} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{$u_1$} & \colhead{$u_2$} & \colhead{$u$} \\\\"
+print "\colhead{(micron)} & \colhead{(method 1)\\tablenotemark{*}} & \colhead{(method 2)} & \colhead{(method 3)} & \colhead{(PHOENIX)} & \colhead{(PHOENIX)} & \colhead{(fitted)}}"
+print "\\startdata"
+
 for i in range(n):
-#	print '{0:0.3f}'.format(d[i,0]), '&', '{0:0.0f}'.format((d[i,1] - np.mean(d[:,1]))*1.e6), '&', '{0:0.0f}'.format(d[i,2]*1e6), '\\\\' 
[…]

### 2018-02-25 ApJL_submission/ms.tex
clouds

@@
 Since the planet's 10-bar radius is a free parameter in our analysis of the spectrum, the absolute transit depth does not affect the retrieved atmospheric properties.
-In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions (coefficients listed in Table\,\ref{tab:tspec}.
+In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions \added{for a star with $T_\mathrm{eff} = 4300 K$, $\log g = 4.5$, and $[\mathrm{Fe/H}] = 0.$} (coefficients listed in Table\,\ref{tab:tspec}.
 The final spectroscopic light curve fits had just three free parameters per channel: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
@@
 In one test, we fit the spectroscopic light curves with the same analytic model we used for the white light curve.
-\added{This fit (which we label ``Method B") has additional free parameters, so the transit depth uncertainties increase by 50\%, but the best fit transit depths are consistent to within $0.5\sigma$ on average over the bandpass.} We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
+\added{This fit (which we label ``Method B") has additional free parameters, so the transit depth uncertainties increase by 50\%, but the best fit transit depths are consistent within $0.5\sigma$ on average with method A over the bandpass.} We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
 The results from these tests are nearly identical to our nominal transmission spectrum, except with a small constant offset due to the uncorrected star-spot crossing feature.
@@
 The transit depths and fitted limb darkening coefficients are provide in Table\,\ref{tab:tspec}.} The fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths.
-\added{We calculated model limb darkening coefficients for WASP-107 over the $1\,\sigma$ range of stellar effective temperatures, surface gravities, and metallicities published in the discovery paper ($T_\mathrm{eff} = 4430\pm130$ K, $\log g = 4.5 \pm 0.1$, and $[\mathrm{Fe/H}] = 0.02 \pm 0.1i$).
+	\added{To test that uncertainties on the stellar parameters do not significantly bias the limb darkening coefficients, we calculated PHOENIX model limb coefficients for WASP-107 over the $1\,\sigma$ range of stellar effective temperatures, surface gravities, and metallicities published in the discovery paper ($T_\mathrm{eff} = 4430\pm130$ K, $\log g = 4.5 \pm 0.1$, and $[\mathrm{Fe/H}] = 0.02 \pm 0.1$).
 The uncertainty in $T_\mathrm{eff}$ has the largest effect on the estimated limb darkening coefficients.
@@
 
-In addition to these tests, we also performed an independent data reduction and fit using K. Stevenson's pipeline and found consistent results (well within $1\,\sigma$).
+In addition to these tests, we also performed an independent data reduction and fit using K. Stevenson's pipeline and again found consistent results (well within $1\,\sigma$).
 
@@
 To estimate the impact of unocculted spots on our data,  we first estimated the star spot properties based on the observed photometric variability in the Kepler band pass, which is 0.3\% \citep{dai17,mocnik17}.
-This variability can be reproduced with a single spot with covering fraction of 1\% and temperature 300 K lower than the stellar photosphere (4130 vs 4430 K).
+This variability can be reproduced with a single spot with covering fraction of 1\% and temperature 300 K lower than the stellar photosphere.
 We calculated the effect of unocculted spots with these properties using equation (1) from \cite{mccullough14}, and predict that they will introduce a positive slope in the transmission spectrum of 8 ppm over the WFC3 bandpass, well below our measurement uncertainties.
@@
 We also explored cases without the $30\times$ solar metallicity prior upper limit.
-Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57+{0.54}_{-0.31}$, and cloud top pressure $log(Pc) = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
-The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[M/H] = 1.5$).
-The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17}.}
+Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
+The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$).
+The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit14}.}
 
@@
 In addition to the atmospheric retrieval, we also performed forward modeling of physically motivated, self-consistent clouds and hazes using the methods described in \cite{fortney08, morley15}.
-We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS; see \citealt{morley12}), for a range of metallicites from $1-50\times$ solar.
-We vary the cloud sedimentation efficiency from 3 to 0.1 (normal to highly lofted, small particulate clouds).
-\added{FIXME}.
+We modeled clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS; see \citealt{morley12}), for a solar composition atmosphere.
+We varied the cloud sedimentation efficiency from 3 to 0.3 (normal to highly lofted, small particulate clouds).
+\added{Only the model with the most lofted clouds ($f_\mathrm{sed} = 0.3$) produces sufficiently small amplitude features to match the observed spectrum.
+This result is in agreement with findings for other small planets, such as the low sedimentation efficiency inferred for GJ 1214b \citep[$f_\mathrm{sed} < 0.1$][]{morley15}.}
+%cite Peter Gao's paper?
 
-None of the models produce sufficiently low amplitude spectral features to match the observed spectrum.
-This result suggests that if clouds are muting WASP-107b's transmission spectrum, the mechanism for cloud particle formation and vertical lofting is very efficient and not captured by the modeling used here.
+%This result suggests that if clouds are muting WASP-107b's transmission spectrum, the mechanism for cloud particle formation and vertical lofting is very efficient and not captured by the modeling used here.
 Futher studies of cloud formation in very low gravity environments are required.
 
-We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere.
-We predict the abundance of hydrocarbon haze precursors from previously published photochemical models for GJ 436b \citep{line11, morley17}.
-With the most efficient haze production (100\% of precursors form haze) and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.
+We also modeled an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere.
+We predicted the abundance of hydrocarbon haze precursors from previously published photochemical models for GJ 436b \citep{line11, morley17}.
+\added{We ran haze models for two different particle sizes ($0.03$ and $0.3\,\mu$m), for haze production efficiencies $f_\mathrm{haze}$ of 3, 10, and 30\%.
+We found that the larger particles ($0.3\,\mu$m) with moderate haze production efficiency (3\%) match the shape and amplitude of the spectral features well.
+More efficient haze production produces features smaller than we measure, and smaller particles sizes produce a slope (an increase in transit depth towards bluer wavelengths) that is not seen.}
 
-These photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis; however, we found in the previous section that methane is depleted.
+%These photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis; however, we found in the previous section that methane is depleted.
 Furthermore, even if the precursors were present, it is unclear whether the 100\% haze production efficiency is realistic.
@@
 The amplitude of the water absorption feature in the transmission spectrum is less than a third that expected for a clear atmosphere.
-Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the data.
-All the cloud models we consider produce condensation too deep in the atmosphere.
-Efficient hydrocarbon haze production at high altitudes can match the data, but these require the presence of methane to serve as a haze precursor.
-Other precursors such as H$_2$S are a possibility and would be worth exploring in future analyses.
-Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the atmospheres of the coolest planets.}
+\added{We find that physically-motivated cloud and haze formation models can satisfactorily reproduce the observed feature amplitude.
+Either lofted clouds with low sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or $0.3\,\mu$m haze particles with moderate haze production efficiency (3\%) match the spectrum.} Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the atmospheres of the coolest planets.}
 \end{itemize}

### 2018-02-26 ApJL_submission/ms.tex
final

@@
 The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern.
-\added{The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to methane depletion.} \deleted{and we infer a water abundance consistent with expectations for a solar abundance pattern.
-On the other hand, methane is depleted relative to expectations (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.}  The spectral features are smaller than predicted for a cloudless atmosphere, crossing less than one scale height.
+\added{The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to possible methane depletion in the atmosphere.}  The spectral features are smaller than predicted for a cloud-free composition, crossing less than one scale height.
 A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations.
-\added{We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) produce clouds at this pressure level.} \deleted{however, we find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.}  Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
+\added{We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or hazes with a particle size of 0.3 $\mu$m reproduce the observed spectral feature amplitude.} Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
 The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities.
@@
 		\colhead{Wavelength} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{Transit Depth} & \colhead{$u_1$} & \colhead{$u_2$} & \colhead{$u$} \\
-	\colhead{(micron)} & \colhead{(Method A)\tablenotemark{*}} & \colhead{(Method B)} & \colhead{(Method C)} & \colhead{(PHOENIX)} & \colhead{(PHOENIX)} & \colhead{(fitted)}}
+	\colhead{(micron)} & \colhead{(Method A)} & \colhead{(Method B)} & \colhead{(Method C)} & \colhead{(PHOENIX)} & \colhead{(PHOENIX)} & \colhead{(fitted)}}
 		\startdata
@@
 		\enddata
-		\tablenotetext{*}{Our retrieval analysis uses the transit depths from Method A.}
-		\tablecomments{comments}
+		\tablecomments{Our retrieval analysis uses the transit depths from Method A.
+Transit depths from Methods A and B use quadratic limb darkening coefficients calculated from a PHOENIX stellar model (listed in columns 5 and 6).
+Method C fits a linear limb darkening coefficient (fit results listed in column 7).}
 	\end{deluxetable*}
@@
 The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$).
-The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit14}.}
+The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.}
 
@@
 We also modeled an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere.
-We predicted the abundance of hydrocarbon haze precursors from previously published photochemical models for GJ 436b \citep{line11, morley17}.
+We predicted the abundance of hydrocarbon haze precursors from previously published $50\times$ solar photochemical models for GJ 436b \citep{line11, morley17}.
 \added{We ran haze models for two different particle sizes ($0.03$ and $0.3\,\mu$m), for haze production efficiencies $f_\mathrm{haze}$ of 3, 10, and 30\%.
@@
 {Dawson}, R.~I., {Lee}, E.~J., \& {Chiang}, E. 2016, \apj, 822, 54
+
+\bibitem[{{de Wit} \& {Seager}(2013)}]{dewit13}
+{de Wit}, J., \& {Seager}, S. 2013, Science, 342, 1473
 

### 2018-02-26 arxiv_v3/ms.tex
arxiv v3

--- 2018-02-26 ApJL_submission/ms.tex
@@
 
-\title{Water, High-Altitude Condensates, and \added{Possible} Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}}
+\title{Water, High-Altitude Condensates, and Possible Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}}
 
@@
 The transmission spectrum shows strong evidence for water absorption ($6.5\,\sigma$ confidence), and the retrieved water abundance is consistent with expectations for a solar abundance pattern.
-\added{The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to possible methane depletion in the atmosphere.}  The spectral features are smaller than predicted for a cloud-free composition, crossing less than one scale height.
+The inferred carbon-to-oxygen ratio is subsolar at $2.7\,\sigma$ confidence, which we attribute to possible methane depletion in the atmosphere.
+The spectral features are smaller than predicted for a cloud-free composition, crossing less than one scale height.
 A thick condensate layer at high altitudes (0.1 - 3 mbar) is needed to match the observations.
-\added{We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or hazes with a particle size of 0.3 $\mu$m reproduce the observed spectral feature amplitude.} Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
+We find that physically motivated cloud models with moderate sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or hazes with a particle size of 0.3 $\mu$m reproduce the observed spectral feature amplitude.
+Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres.
 The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities.
@@
 This results in a distinct atmospheric chemistry compared to other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.
-Spectral features from both water and methane are accessible with current observing facilities -- \added{both of these molecules have strong absorption bands in the wavelength range covered by \emph{HST}/WFC3.
-Measuring the shape and amplitude of these spectral features enables a constraint on the abundance of water, the dominant oxygen-bearing molecule, and methane, the dominant carbon-bearing molecule, providing a bounded estimate of the carbon-to-oxygen ratio (C/O).} Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage, \added{and have mainly resulted in upper limits} \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.
+Spectral features from both water and methane are accessible with current observing facilities -- both of these molecules have strong absorption bands in the wavelength range covered by \emph{HST}/WFC3.
+Measuring the shape and amplitude of these spectral features enables a constraint on the abundance of water, the dominant oxygen-bearing molecule, and methane, the dominant carbon-bearing molecule, providing a bounded estimate of the carbon-to-oxygen ratio (C/O).
+Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage, and have mainly resulted in upper limits \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.
 
@@
 Since the planet's 10-bar radius is a free parameter in our analysis of the spectrum, the absolute transit depth does not affect the retrieved atmospheric properties.
-In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions \added{for a star with $T_\mathrm{eff} = 4300 K$, $\log g = 4.5$, and $[\mathrm{Fe/H}] = 0.$} (coefficients listed in Table\,\ref{tab:tspec}.
+In addition to fixing the transit parameters, we also fixed the limb darkening on the PHOENIX model predictions for a star with $T_\mathrm{eff} = 4300 K$, $\log g = 4.5$, and $[\mathrm{Fe/H}] = 0.$ (coefficients listed in Table\,\ref{tab:tspec}.
 The final spectroscopic light curve fits had just three free parameters per channel: $C_\lambda$, $V_\lambda$, and $r_p/r_s$.
@@
 In one test, we fit the spectroscopic light curves with the same analytic model we used for the white light curve.
-\added{This fit (which we label ``Method B") has additional free parameters, so the transit depth uncertainties increase by 50\%, but the best fit transit depths are consistent within $0.5\sigma$ on average with method A over the bandpass.} We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
+This fit (which we label ``Method B") has additional free parameters, so the transit depth uncertainties increase by 50\%, but the best fit transit depths are consistent within $0.5\sigma$ on average with method A over the bandpass.
+We also tested a white light curve systematics vector $S_\mathrm{white}$ from a fit that included the star-spot crossing orbit.
 The results from these tests are nearly identical to our nominal transmission spectrum, except with a small constant offset due to the uncorrected star-spot crossing feature.
 This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval with the 10-bar radius parameter (see \S\,\ref{sec:retrieval}).
-	\added{We also repeated the calculations from Method A, but fit} for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values.
-\added{This approach, which we dub ``Method C", also yields consistent results with Method A (the relative transit depths agree to better than 0.2 sigma on average).
-The transit depths and fitted limb darkening coefficients are provide in Table\,\ref{tab:tspec}.} The fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths.
-	\added{To test that uncertainties on the stellar parameters do not significantly bias the limb darkening coefficients, we calculated PHOENIX model limb coefficients for WASP-107 over the $1\,\sigma$ range of stellar effective temperatures, surface gravities, and metallicities published in the discovery paper ($T_\mathrm{eff} = 4430\pm130$ K, $\log g = 4.5 \pm 0.1$, and $[\mathrm{Fe/H}] = 0.02 \pm 0.1$).
+
+We also repeated the calculations from Method A, but fit for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values.
+This approach, which we dub ``Method C", also yields consistent results with Method A (the relative transit depths agree to better than 0.2 sigma on average).
+The transit depths and fitted limb darkening coefficients are provide in Table\,\ref{tab:tspec}.
+The fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths.
+
+To test that uncertainties on the stellar parameters do not significantly bias the limb darkening coefficients, we calculated PHOENIX model limb coefficients for WASP-107 over the $1\,\sigma$ range of stellar effective temperatures, surface gravities, and metallicities published in the discovery paper ($T_\mathrm{eff} = 4430\pm130$ K, $\log g = 4.5 \pm 0.1$, and $[\mathrm{Fe/H}] = 0.02 \pm 0.1$).
 The uncertainty in $T_\mathrm{eff}$ has the largest effect on the estimated limb darkening coefficients.
@@
 Over this temperature range, the change in limb darkening shifts the average transit depth by less than 20 ppm, which is below the level of noise in the data.
-The relative change in transit depth is smaller still ($< 1$ ppm).}
+The relative change in transit depth is smaller still ($< 1$ ppm).
 
@@
 
-\added{\subsubsection{Impact of Unocculted Star Spots}
+\subsubsection{Impact of Unocculted Star Spots}
 Unocculted star spots may influence the measured transmission spectrum \citep{mccullough14, zellem17}.
@@
 We note that cool stars may have persistent spot coverage that does not produce any photometric variability \citep{rackham18}.
-However, even if there is persistent spot coverage of 5\%, the spots produce a slope in the WFC3 bandpass of 50 ppm, which is still below our uncertainties.}
+However, even if there is persistent spot coverage of 5\%, the spots produce a slope in the WFC3 bandpass of 50 ppm, which is still below our uncertainties.
 
@@
 We also explored cases without the $30\times$ solar metallicity prior upper limit.
-Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
+Without this prior, the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
 The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$).
-The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.}
+The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.
 
@@
 \subsubsection{Methane Abundance}
-\added{If methane is present in the atmosphere, it has a strong spectral feature centered at $1.4\,\mu$m.
+If methane is present in the atmosphere, it has a strong spectral feature centered at $1.4\,\mu$m.
 This feature overlaps the water feature, but has a wider shape.
-Our data are precise enough to distinguish this subtle difference in feature shape, so } to explore whether methane depletion is the underlying cause of the inferred low C/O, we ran two additional retrievals.
+Our data are precise enough to distinguish this subtle difference in feature shape, so to explore whether methane depletion is the underlying cause of the inferred low C/O, we ran two additional retrievals.
 In one test, we assumed chemical equilibrium but excluded methane opacity.
@@
 The $3\,\sigma$ upper limit on methane abundance is $1.4\times10^{-3}$.
-\added{This value is in tension with expectations for a solar abundance pattern: for $3.7\times$ solar metallicity, $650$ K, and $5\times10^{-4}$ bar (the median retrieved photosphere properties), the expected methane abundance is $1.2\times10^{-3}\times$ solar (for solar C/O).} Based on these tests, we conclude that the atmosphere of WASP-107b is \added{likely} depleted in methane relative to expectations for a solar abundance pattern.
+This value is in tension with expectations for a solar abundance pattern: for $3.7\times$ solar metallicity, $650$ K, and $5\times10^{-4}$ bar (the median retrieved photosphere properties), the expected methane abundance is $1.2\times10^{-3}\times$ solar (for solar C/O).
+Based on these tests, we conclude that the atmosphere of WASP-107b is likely depleted in methane relative to expectations for a solar abundance pattern.
 By contrast, the water abundance from the free retrieval ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions for solar composition.
-
-%we calculate the thermochemical equilibrium methane abundance for ($1.2\time10^{-3}$ at 650 K, $5\times10^4$ bar, for a solar C/O and $3.7\,\times$ solar metallicity composition).}  We chose $3\times$ solar as a benchmark for comparison because this is the median metallicity from the chemical equilibrium retrieval; \added{however, we note that the expected equilibrium abundance is sensitive to the true atmosphere metallicity.
-For a $1\times$ solar composition, the expected methane abundance drops by $1/3$.
-Similarly, the expected abundances are sensitive to the temperature and pressure in the photosphere.
-For example, at slightly higher temperatures and higher altitudes (1 mbar, 800 K), the expected methane abundance drops to $2.4\times10^{-5}$.}
 
@@
 We varied the cloud sedimentation efficiency from 3 to 0.3 (normal to highly lofted, small particulate clouds).
-\added{Only the model with the most lofted clouds ($f_\mathrm{sed} = 0.3$) produces sufficiently small amplitude features to match the observed spectrum.
-This result is in agreement with findings for other small planets, such as the low sedimentation efficiency inferred for GJ 1214b \citep[$f_\mathrm{sed} < 0.1$][]{morley15}.}
+Only the model with the most lofted clouds ($f_\mathrm{sed} = 0.3$) produces sufficiently small amplitude features to match the observed spectrum.
+This result is in agreement with findings for other small planets, such as the low sedimentation efficiency inferred for GJ 1214b \citep[$f_\mathrm{sed} < 0.1$][]{morley15}.
 %cite Peter Gao's paper?
@@
 We predicted the abundance of hydrocarbon haze precursors from previously published $50\times$ solar photochemical models for GJ 436b \citep{line11, morley17}.
-\added{We ran haze models for two different particle sizes ($0.03$ and $0.3\,\mu$m), for haze production efficiencies $f_\mathrm{haze}$ of 3, 10, and 30\%.
+We ran haze models for two different particle sizes ($0.03$ and $0.3\,\mu$m), for haze production efficiencies $f_\mathrm{haze}$ of 3, 10, and 30\%.
 We found that the larger particles ($0.3\,\mu$m) with moderate haze production efficiency (3\%) match the shape and amplitude of the spectral features well.
-More efficient haze production produces features smaller than we measure, and smaller particles sizes produce a slope (an increase in transit depth towards bluer wavelengths) that is not seen.}
+More efficient haze production produces features smaller than we measure, and smaller particles sizes produce a slope (an increase in transit depth towards bluer wavelengths) that is not seen.
 
@@
 Compared to results for other exoplanets of similar mass such as GJ 436b, which has a high metallicity ($>100\times$ solar) and HAT-P-26b, which is metal-poor compared to the Solar System trend, WASP-107b adds to the evidence that exoplanets exhibit a greater diversity of compositions than is present in the Solar System \citep{morley17, wakeford17}.}
-\item{\emph{The methane abundance is \added{likely} depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
+\item{\emph{The methane abundance is likely depleted relative to expectations for a solar abundance pattern, whereas water is consistent with solar composition.} This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.
 Another possibility is that the planet has a hot interior effective temperature ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}.
@@
 The amplitude of the water absorption feature in the transmission spectrum is less than a third that expected for a clear atmosphere.
-\added{We find that physically-motivated cloud and haze formation models can satisfactorily reproduce the observed feature amplitude.
-Either lofted clouds with low sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or $0.3\,\mu$m haze particles with moderate haze production efficiency (3\%) match the spectrum.} Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the atmospheres of the coolest planets.}
+We find that physically-motivated cloud and haze formation models can satisfactorily reproduce the observed feature amplitude.
+Either lofted clouds with low sedimentation efficiency ($f_\mathrm{sed} = 0.3$) or $0.3\,\mu$m haze particles with moderate haze production efficiency (3\%) match the spectrum.
+Put in context with other systems, the muted water feature for WASP-107b agrees well with the trend in feature amplitude with temperature noted in \cite{crossfield17}, indicating that condensates may be common in the atmospheres of the coolest planets.}
 \end{itemize}

### 2018-04-16 arxiv_v3/ms.tex
draft response 3

@@
 \author{Laura Kreidberg}
-\affiliation{Harvard Society of Fellows\
-78 Mt. Auburn St.\\
-Cambridge, MA 02138, USA}
-\affiliation{Harvard-Smithsonian Center for Astrophysics\
-60 Garden St.\\
-Cambridge, MA 02138}
-\nocollaboration
+\affiliation{Harvard Society of Fellows, 78 Mt. Auburn St., Cambridge, MA 02138, USA}
+\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA 02138}
 
 \author{Michael R. Line}
-\affiliation{Arizona State University}
-\nocollaboration
+\affiliation{School of Earth \& Space Exploration, Arizona State University, Tempe AZ 85287, USA}
 
 \author{Daniel Thorngren}
-\affiliation{University of California Santa Cruz}
-\nocollaboration
+\affiliation{Department of Astronomy and Astrophysics, University of California, Santa Cruz, CA 95064}
 
@@
 \altaffiliation{Sagan Fellow}
-%\affiliation{Harvard-Smithsonian Center for Astrophysics}
-\affiliation{Harvard-Smithsonian Center for Astrophysics\
-60 Garden St.\\
-Cambridge, MA 02138}
-\nocollaboration
+\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA 02138}
 
 \author{Kevin B. Stevenson}
-\affiliation{Space Telescope Science Institute}
-\nocollaboration
+\affiliation{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA}
 

### 2018-04-16 ApJL_submission/Revision3_figs/summary_plot.py

@@ -31,11 +31,11 @@ solarctoo = np.log10(0.54)
 #files = ["CC_FULL_TP_samples.pic", "CC_locked_TP_shape_samples.pic", "CC_locked_TP_shape_Metallicity_Prior_samples.pic"]
 #ind = [4,2,2]
 
-#files = ["CC_FULL_TP_samples.pic", "CC_locked_TP_shape_Metallicity_Prior_samples.pic"]
-#ind = [4,2]
+files = ["CC_FULL_TP_samples.pic", "CC_locked_TP_shape_Metallicity_Prior_samples.pic"]
+ind = [4,2]
 
-files = ["CC_FULL_TP_samples.pic", "CC_locked_TP_shape_Metallicity_Prior_samples.pic", "CC_FULL_TP_Quench_limit_Met_samples.pic"]
-ind = [4,2, 4]
+#files = ["CC_FULL_TP_samples.pic", "CC_locked_TP_shape_Metallicity_Prior_samples.pic", "CC_FULL_TP_Quench_limit_Met_samples.pic"]
+#ind = [4,2, 4]
 
 labels = ['Parameterized T/P', 'Scaled T/P', 'Parameterized T/P + Quench']
 colors = ['red', 'blue', 'orange']
@@ -57,7 +57,8 @@ plt.ylabel("Probability Density")
 
 
 plt.subplot(122)
-files = ["CC_FULL_TP_TP.pic", "CC_locked_TP_shape_TP.pic", "CC_FULL_TP_Quench_limit_Met_TP.pic"]
+files = ["CC_FULL_TP_TP.pic", "CC_locked_TP_shape_TP.pic"]
+#files = ["CC_FULL_TP_TP.pic", "CC_locked_TP_shape_TP.pic", "CC_FULL_TP_Quench_limit_Met_TP.pic"]
 
 for i, f in enumerate(files):
     P, T_array = pickle.load(open(f, "rb"))
@@ -69,6 +70,7 @@ for i, f in enumerate(files):
         Tlo.append(q[0])
         Tmid.append(q[1])
         Thi.append(q[2])
+        if (P[j] >= 0.000099)&(P[j] <= 0.00011): print "0.0001",   q[1], q[2] - q[1], q[1] - q[0]
 
     plt.plot(np.array(Tmid), P, color = colors[i], label = labels[i])
     plt.fill_betweenx(P, np.array(Tlo), np.array(Thi), color = colors[i], linestyle = 'dashed', alpha = 0.2)

### 2018-04-16 Figures/fig2_spectrum.py

@@ -20,8 +20,8 @@ plt.errorbar(d[:,0], d[:,1]*100., d[:,2]*100., linestyle = 'none', zorder = 100,
 #d = np.genfromtxt("wasp107b_10xsolar_0.1mbarcloud.dat")
 #plt.plot(d[:,0]*1e6, convolve(d[:,1], g, boundary = 'extend')-0.08)
 
-d = np.genfromtxt("EQ_Solar_full_trans.txt", skip_header = 1)
-plt.plot(d[:,0], d[:,1], color = 'r', zorder = 100)
+#d = np.genfromtxt("EQ_Solar_full_trans.txt", skip_header = 1)
+#plt.plot(d[:,0], d[:,1], color = 'r', zorder = 100)
 
 p = pickle.load(open("WASP107b_Global_Cloud_limit_Metallicity_spectra.pic", "rb"))
 print "outliers", (p[9] - p[7])/p[8]
@@ -34,14 +34,21 @@ plt.fill_between(w, p[1], p[5], color = "navy", alpha = '0.2')
 plt.fill_between(w, p[2], p[4], color = "navy", alpha = '0.3')
 #plt.plot(w, p[3], color = 'navy', label = "Best fit model")
 
+
+p2 = pickle.load(open("compare_spectra.pic", "rb"))
+plt.plot(p2[0], 100.*p2[2], label = "best fit, low C/O", zorder = 100, color = '0.2', linestyle = 'dashed')
+plt.plot(p2[0], 100.*p2[1], label = "best fit 2", zorder = 100, color = 'y', linestyle = 'dashed')
+
+
 # best fit
 pb = pickle.load(open("Model_best_fit.pic", "rb"))
-plt.plot(pb[0], pb[1]*100., color = 'navy', label = 'Best fit model')
+plt.plot(pb[0], pb[1]*100., color = 'navy', label = 'best fit model')
 #print 'outliers', (p[7] - pb[3]*100.)/p[8]
 
-d = pickle.load(open("Model_Solar.pic", "rb"))
-x, y = d[0], d[1]*100. - 0.7
-plt.plot(x, y, color = '0.5', alpha = 0.5, label = "1X solar, cloud-free", zorder = -20) 
+#d = pickle.load(open("Model_Solar.pic", "rb"))
+#x, y = d[0], d[1]*100. - 0.7
[…]

### 2018-04-17 ApJL_submission/ms.tex
final response 3

@@
 The right-hand axis indicates the normalized atmospheric scale height assuming solar composition.
-The water feature crosses less than one scale height, indicating that condensates are present at high altitude.}
+The water feature crosses less than one scale height, indicating that condensates are present at high altitude.
+\added{We also show the best fit model from an analysis where the C/O is fixed at 0.54, the solar value (dashed line).
+There are subtle differences in feature shape due to methane absorption in the solar C/O case, which cause the retrieval to favor sub-solar C/O values at 2.7$\sigma$ confidence.}}
 \label{fig:spectrum}
@@
 
-We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.
+We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, a power law haze opacity, \added{no prior on atmospheric metallicity, a more flexible T/P profile (with added free parameters for infrared opacity and visible-to-infrared opacity), and the more flexible T/P profile with quenching}.
 We also varied the assumed planet mass by $3\,\sigma$.
-These models did not significantly improve the fit quality, and the main retrieval results were unchanged.
-We also explored cases without the $30\times$ solar metallicity prior upper limit.
-Without this prior, \added{the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
+These models did not significantly improve the fit quality, and the \added{retrieved cloud-top pressure and chemical composition were consistent to within $0.1\,\sigma$}.
+
+\added{For example, when we removed the $30\times$ solar metallicity prior upper limit,} the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
 The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$).
-The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.}
+The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.
+
+\added{Similarly, we find that the transmission spectrum does not provide a strong constraint on the shape of the temperature pressure profile.
+The primary constraint on temperature structure is the effective scale height temperature, which sets the amplitude of features in the spectrum.
+To confirm this, we explored using a more flexible T/P profile with free parameters for infrared and visible-to-infrared opacity, in addition to irradiation temperature (which we label case 2).
+Compared to our nominal model (case 1), which just varies irradiation temperature but keeps the T/P profile shape fixed, we found that the retrieved temperature at the photosphere is nearly identical.
+For example, at $P = 10^{-4}$ bar (the most probable cloud-top pressure), we find $T = 618^{+150}_{-120}$ K for case 1, and $T = 615^{+178}_{-123}$ K for case 2.
+The constraints on chemical composition agree at the 0.1-$\sigma$ level for the two cases.}
 

### 2018-04-17 ApJL_submission/Revision3_figs/summary_plot.py

@@ -37,7 +37,7 @@ ind = [4,2]
 #files = ["CC_FULL_TP_samples.pic", "CC_locked_TP_shape_Metallicity_Prior_samples.pic", "CC_FULL_TP_Quench_limit_Met_samples.pic"]
 #ind = [4,2, 4]
 
-labels = ['Parameterized T/P', 'Scaled T/P', 'Parameterized T/P + Quench']
+labels = ['Case 2', 'Case 1', 'Case 3']
 colors = ['red', 'blue', 'orange']
 
 for i, f in enumerate(files):
@@ -50,7 +50,7 @@ for i, f in enumerate(files):
     print "0.16, 0.5, 0.84 quantiles", q[1], "+", q[2] - q[1], "-", q[1] - q[0]
 
 plt.axvline(solarctoo, color = '0.5')
-
+plt.gca().text(np.log10(0.54) - 0.1, 0.7, 'solar C/O', rotation = 90) 
 plt.legend(loc = 'upper right', frameon = True)
 plt.xlabel("log10(C/O)")
 plt.ylabel("Probability Density")
@@ -76,7 +76,7 @@ for i, f in enumerate(files):
     plt.fill_betweenx(P, np.array(Tlo), np.array(Thi), color = colors[i], linestyle = 'dashed', alpha = 0.2)
 plt.gca().set_yscale('log')
 
-#plt.legend(loc = 'upper left')
+plt.legend(loc = 'upper left')
 plt.ylim(1.e-3, 1e-6)
 plt.xlim(300, 1000)
 plt.xlabel("Temperature (K)")

### 2018-04-17 Figures/fig2_spectrum.py

@@ -35,9 +35,6 @@ plt.fill_between(w, p[2], p[4], color = "navy", alpha = '0.3')
 #plt.plot(w, p[3], color = 'navy', label = "Best fit model")
 
 
-p2 = pickle.load(open("compare_spectra.pic", "rb"))
-plt.plot(p2[0], 100.*p2[2], label = "best fit, low C/O", zorder = 100, color = '0.2', linestyle = 'dashed')
-plt.plot(p2[0], 100.*p2[1], label = "best fit 2", zorder = 100, color = 'y', linestyle = 'dashed')
 
 
 # best fit
@@ -48,6 +45,11 @@ plt.plot(pb[0], pb[1]*100., color = 'navy', label = 'best fit model')
 #d = pickle.load(open("Model_Solar.pic", "rb"))
 #x, y = d[0], d[1]*100. - 0.7
 #plt.plot(x, y, color = '0.5', alpha = 0.5, label = "1X solar, cloud-free", zorder = -20) 
+
+#plots best fit with solar C/O
+p2 = pickle.load(open("compare_spectra.pic", "rb"))
+plt.plot(p2[0], 100.*p2[2], label = "best fit, solar C/O", zorder = 100, color = '0.2', linestyle = 'dashed')
+#plt.plot(p2[0], 100.*p2[1], label = "best fit 2", zorder = 100, color = 'y', linestyle = 'dashed')
 
 plt.legend(loc= 'upper right')
 

### 2018-04-30 arxiv_v4_accepted/ms.tex
accepted version with proof corrections

--- 2018-04-17 ApJL_submission/ms.tex
@@
 
-The best fit light curves have a median $\chi^2_\nu$ value of 1.16.
-We conservatively rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2_\nu$ values are unity.
-We performed an MCMC fit to the light curves with \texttt{emcee}.
-For each light curve we ran a fit with 50 walkers and 1000 steps per walker, and tested for convergence as we did for the white light curve.
-The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:tspec}.
+The best fit light curves have a median $\chi^2_\nu$ value of 1.16.  We
+conservatively rescaled the photometric uncertainties for all spectroscopic
+channels such that the $\chi^2_\nu$ values are unity. We performed an MCMC fit
+to the light curves with \texttt{emcee}.  For each light curve we ran a fit with
+50 walkers and 1000 steps per walker, and tested for convergence as we did for
+the white light curve. The median transit depths and $1\,\sigma$ uncertainties
+are given in Table\,\ref{tab:tspec}, and the transmission spectrum is shown in
+Figure\,\ref{fig:spectrum}.
 
@@
 The water feature crosses less than one scale height, indicating that condensates are present at high altitude.
-\added{We also show the best fit model from an analysis where the C/O is fixed at 0.54, the solar value (dashed line).
-There are subtle differences in feature shape due to methane absorption in the solar C/O case, which cause the retrieval to favor sub-solar C/O values at 2.7$\sigma$ confidence.}}
+We also show the best fit model from an analysis where the C/O is fixed at 0.54, the solar value (dashed line).
+There are subtle differences in feature shape due to methane absorption in the solar C/O case, which cause the retrieval to favor sub-solar C/O values at 2.7$\sigma$ confidence.}
 \label{fig:spectrum}
@@
 
-The nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.
+The best-fit spectrum is shown in Figure\,\ref{fig:spectrum}, and the nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.
 The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).
@@
 
-We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, a power law haze opacity, \added{no prior on atmospheric metallicity, a more flexible T/P profile (with added free parameters for infrared opacity and visible-to-infrared opacity), and the more flexible T/P profile with quenching}.
+We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, a power law haze opacity, no prior on atmospheric metallicity, a more flexible T/P profile (with added free parameters for infrared opacity and visible-to-infrared opacity), and the more flexible T/P profile with quenching.
 We also varied the assumed planet mass by $3\,\sigma$.
-These models did not significantly improve the fit quality, and the \added{retrieved cloud-top pressure and chemical composition were consistent to within $0.1\,\sigma$}.
+These models did not significantly improve the fit quality, and the retrieved cloud-top pressure and chemical composition were consistent to within $0.1\,\sigma$.
 
-\added{For example, when we removed the $30\times$ solar metallicity prior upper limit,} the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
+For example, when we removed the $30\times$ solar metallicity prior upper limit, the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.
 The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$).
@@
 
-\added{Similarly, we find that the transmission spectrum does not provide a strong constraint on the shape of the temperature pressure profile.
+Similarly, we find that the transmission spectrum does not provide a strong constraint on the shape of the temperature pressure profile.
 The primary constraint on temperature structure is the effective scale height temperature, which sets the amplitude of features in the spectrum.
@@
 For example, at $P = 10^{-4}$ bar (the most probable cloud-top pressure), we find $T = 618^{+150}_{-120}$ K for case 1, and $T = 615^{+178}_{-123}$ K for case 2.
-The constraints on chemical composition agree at the 0.1-$\sigma$ level for the two cases.}
+The constraints on chemical composition agree at the 0.1-$\sigma$ level for the two cases.
 
@@
 acknowledges support from NASA through the Sagan Fellowship Program.
+M.R.L.
+acknowledges NASA XRP grant NNX17AB56G for partial support of the theoretical interpretation of the data, as well as the ASU Research Computing staff for support with the Saguaro and Agave compute clusters.
 
@@
 
-\bibitem[{{Fraine} {et~al.}(2014){Fraine}, {Deming}, {Benneke}, {Knutson},
-  {Jord{\'a}n}, {Espinoza}, {Madhusudhan}, {Wilkins}, \& {Todorov}}]{fraine14}
-{Fraine}, J., {Deming}, D., {Benneke}, B., {et~al.} 2014, \nat, 513, 526
-
 \bibitem[{{Freedman} {et~al.}(2014){Freedman}, {Lustig-Yaeger}, {Fortney},
@@
 {Frelikh}, R., \& {Murray-Clay}, R.~A. 2017, ArXiv e-prints, arXiv:1708.00862
-
-\bibitem[{{Gao} {et~al.}(2017){Gao}, {Marley}, {Zahnle}, {Robinson}, \&
-  {Lewis}}]{gao17}
-{Gao}, P., {Marley}, M.~S., {Zahnle}, K., {Robinson}, T.~D., \& {Lewis}, N.~K.
-  2017, \aj, 153, 139
 
@@
 
-\bibitem[{{Zahnle} {et~al.}(2009){Zahnle}, {Marley}, {Freedman}, {Lodders}, \&
-  {Fortney}}]{zahnle09}
-{Zahnle}, K., {Marley}, M.~S., {Freedman}, R.~S., {Lodders}, K., \& {Fortney},
-  J.~J. 2009, \apjl, 701, L20
-
 \bibitem[{{Zahnle} \& {Marley}(2014)}]{zahnle14}
